### 2024-10-20 write/main.tex
initial commit

\documentclass[doc, floatsintext]{apa7}
\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
\DeclareLanguageMapping{american}{american-apa}
\addbibresource{sample.bib}
\setlength\bibhang{.15in}
\usepackage{amsmath}
\usepackage{float}
\usepackage{graphicx}
\usepackage{mathrsfs}
\usepackage{setspace}
\setstretch{1.0}
\usepackage{caption}
\usepackage{gensymb}

\title{Criterial Learning and Feedback Delay: Insights from
Computational Models and Behavioral Experiments}
\shorttitle{Computational Criterion Learning}

\authorsnames[{1, 2}, {3}, {4}]{
    Matthew J. Crossley, 
    Benjamin O. Pelzer,
    F. Gregory Ashby
}

\authorsaffiliations{
    {School of Psychological Sciences, Macquarie University, Sydney, Australia}, 
    {Macquarie University Performance and Expertise Research Centre, Macquarie University, Sydney, Australia},
    {Nanyang Technological University},
    {Department of Psychological \& Brain Sciences, University of California, Santa Barbara}
    }
\abstract{The notion of a response criterion is ubiquitous
    in psychology, yet its cognitive and neural
    underpinnings remain poorly understood. To address this
    shortcoming, three computational models that capture
    different hypotheses about criterial learning were
    developed and tested. The time-dependent drift model
    assumes the criterion is stored in working memory and
    that its value drifts over time. The delay-sensitive
    learning model assumes that the magnitude of criterial
    learning is temporally discounted by feedback delay. The
    reinforcement-learning model assumes that criterial
    learning emerges from stimulus-response association
    learning without an explicit representation of the
    criterion, with learning rate also temporally discounted
    by feedback delay. The performance of these models was
    investigated under varying feedback delay and intertrial
    interval (ITI) durations. The time-dependent drift model
    predicted that long ITIs and feedback delays both impair
    criterial learning. In contrast, the delay-sensitive and
    reinforcement-learning models predicted impairments only
    with feedback delays. Two behavioral experiments, which
    tested these predictions, showed that human criterial
    learning is impaired by delayed feedback but not by long
    ITIs.  These results support the delay-sensitive and
    reinforcement-learning models, and suggest that even in
    tasks that appear to rely on explicit, rule-based
    reasoning, criterial learning may have strong
    associative underpinnings.
}

\authornote{Correspondence: Matthew J. Crossley, PhD,
  School of Psychological Sciences, Macquarie University,
  Australian Hearing Hub, 16 University Ave, Macquarie
  University, NSW 2109, Australia. Email:
  matthew.crossley@mq.edu.au 
}

\keywords{response criterion; criterial learning;
associative learning; categorization; procedural learning}

\begin{document}
\maketitle

\section{Introduction}
The notion of a response criterion is ubiquitous in
psychology. It is a key component of almost all decision
models. For example, the hypothesis that even YES-NO
detection decisions are determined by comparing the sensory
magnitude to a response criterion that is under the
observer's control, rather than to a fixed absolute
threshold, allowed signal detection theory to supplant
classical threshold theory as the dominant model in
psychophysics \parencite{GreenSwets1966}. All models that
include a response criterion assume its value is learned and
can shift if changes are made to instructions or payoffs. So
criterial learning is a fundamental component of almost all
decision-making models. Despite its importance, however, the
cognitive and neural mechanisms that underlie criterial
learning remain poorly understood.

This article addresses this shortcoming through a
combination of computational modeling and empirical data
collection. Specifically, we develop and test three
different computational models that make qualitatively
different assumptions about how the criterion is learned.
The models differ in the role they assign to working memory
and in whether they treat the response criterion as a
fundamental psychological construct, or instead assume that
behavior is driven purely by stimulus-response associations
without any criterion guiding responses. These models are
then tested in two behavioral experiments. The modeling and
empirical focus are on how feedback delays and the length of
the intertrial interval (ITI) affect criterial learning. All
criterial-learning models assume that updating (i.e.,
learning) of the criterion occurs during the time interval
between feedback presentation and the stimulus presentation
that defines the onset of the next trial. So feedback delay
and the length of the ITI are the independent variables that
most clearly differentiate the conflicting predictions of
criterial-learning models. Furthermore, feedback delays are
known to impair some forms of learning (i.e., procedural)
much more than others (i.e., learning that relies on
declarative memory), so feedback-delay manipulations offer a
powerful method of disambiguating the nature of criterial
learning \parencite{ell2009critrial, MaddoxAshbyBohil2003,
MaddoxIng2005}.

Although the models could be tested in any task that depends
on criterial learning, the two experiments we describe used
a one-dimensional category-learning task. In such tasks,
stimuli vary across trials on two or more dimensions -- one
relevant to categorization and one or more that are
irrelevant. The observer's goal is to identify the relevant
dimension and learn the response criterion that maximizes
accuracy. This task has been used in hundreds of studies,
and all current models of performance in this task emphasize
the role of criterial learning.

In the first behavioral experiment, participants were
explicitly instructed about the relevant dimension, thereby
isolating criterial learning from rule selection and
switching processes. The results showed that short feedback
delays impaired learning compared to immediate feedback.
The second behavioural experiment used stimuli composed of
binary features -- where no criterial learning is needed --
and did not instruct participants about the relevant
dimension. Thus, this experiment isolated rule selection and
switching from criterial learning. The results found that
increasing feedback delay or ITI did not affect performance.
These findings suggest that feedback delays impact criterial
learning but not the discovery of the relevant stimulus
dimension.  More broadly, they indicate that criterial
learning may recruit procedural learning, even in tasks that
seem to rely on explicit, rule-based processes.

\section{Experiment 1: The Models}
We developed three computational models, each representing
different architectures of criterial learning, and examined
their sensitivity to the duration of ITIs and feedback
delay. Two of the models assume that the criterion is stored
in memory and that response decisions are made by comparing
the current stimulus-driven percept to this stored referent.
Both of these models further posit that the internal
representation of the criterion is updated following errors
using a simple gradient-descent rule. The third model
assumes that no criterion is stored nor used to generate
responses. Instead, reinforcement learning is used to form
stimulus-response associations and these associations drive
responding without any appeal to the notion of a criterion.

For each of the three models, denote the time of stimulus
presentation on the current trial by T$_\text{S}$, the
response time by T$_\text{R}$, the time when feedback is
displayed by T$_\text{F}$, and the time when the stimulus
that begins the next trial is displayed by T$_{\text{S}^+}$.
Then note that the feedback delay equals
$\text{t}_\text{FD} = \text{T}_\text{F} - \text{T}_\text{R}$
and the duration of the ITI equals $\text{t}_\text{ITI} =
\text{T}_{\text{S}^+} - \text{T}_\text{F}$.

\subsection{Time-Dependent Drift Model}
The time-dependent drift model assumes that the observer
constructs a criterion, holds this value in working memory,
and then makes response decisions by comparing the percept
to the criterion. The model also assumes that both the
memory representation of the criterion and the perceived
value of the stimulus gradually drift over time. The model
further assumes that the extent of this drift increases with
time, both within a trial and between consecutive trials.

Let $x_n(t)$ and $c_n(t)$ denote the values of the percept
and the criterion, respectively, on trial $n$ at time $t$.
Then the decision rule on trial $n$ is:
\begin{equation}
  \text{Respond } R_n =
  \begin{cases}
    \text{A}, & \text{if ~} x_n(\text{T}_\text{S}) \leq c_n(\text{T}_\text{S})  \\
    \text{B}, & \text{if ~} x_n(\text{T}_\text{S}) > c_n(\text{T}_\text{S}).
  \end{cases}
  \label{eq:DR}
\end{equation}
If positive feedback is received, then the value of the
criterion remains unchanged for the next trial (except for
drift -- see Equation 3). If negative feedback is received,
then the criterion is modified according to the standard
model \parencite{SuttonBarto1998}:
\begin{equation}
c_n(\text{T}_\text{F}+ \Delta_t) = c_n(\text{T}_\text{F}) + \alpha [x_n(\text{T}_\text{F}) - c_n(\text{T}_\text{F})],
\label{RL}
\end{equation}
where $\alpha$ is a learning-rate parameter, and $\Delta_t$
is the time it takes to complete the updating. It is
straightforward to show that the iterative Equation \ref{RL}
is equivalent to computing a weighted mean (weighted by
recency) of the values of all percepts that occur on error
trials \parencite[e.g.,][]{Ashby2017}. This updating rule
will gradually converge on the optimal criterion value.
Since this model assumes that criterial learning relies on
working memory -- and the available evidence suggests that
logical reasoning and working memory are unaffected by
feedback delays of several seconds \parencite[e.g., in
one-dimensional rule-based category learning
tasks;][]{ell2009critrial, MaddoxAshbyBohil2003,
MaddoxIng2005} -- we assume that the learning process
described in Equation \ref{RL} is likewise unaffected by
feedback delay.

Although the criterial-learning process described by
Equation \ref{RL} is not affected by feedback delay, we
assume that both the stimulus and the criterion
representations drift randomly throughout the duration of
time they are maintained in working memory. The
representation of the criterion must always be maintained in
working memory, whereas drift in the stimulus representation
affects performance up until the feedback is presented, but
not afterwards.

We modeled the drift in both the criterion and the percept
by adding white noise to their initial values. Specifically,
we assumed that for all $t>\text{T}_\text{S}$
\begin{equation}
  c_n(t) = c_n(\text{T}_\text{S}) + \eta_c \epsilon(t),
  \label{eq:criterion}
\end{equation}
and
\begin{equation}
  x_n(t) = x_n(\text{T}_\text{S}) + \eta_x \epsilon(t),
  \label{eq:percept}
\end{equation}
where $\epsilon(t)$ is white noise and $\eta_c$ and $\eta_x$
are parameters that determine the amount of drift over time.
This model predicts that at the time of feedback, the
response criterion $c_n(\text{T}_\text{F})$ will be normally
distributed with mean $c_n(\text{T}_\text{S})$ and variance
$\text{t}_\text{FD} \eta_c^2$. Similarly, at the time when
the stimulus that defines the next trial is presented, the
criterion $c_n(\text{T}_{\text{S}^+}) =
c_{n+1}(\text{T}_\text{S})$ is normally distributed with
mean $c_n(\text{T}_\text{S})$ and variance
$(\text{t}_\text{FD}+\text{t}_\text{ITI}) \eta_c^2$. The
predictions for the percept are similar. Specifically, at
the time when feedback is presented, the percept
$x_n(\text{T}_\text{F})$ is normally distributed with mean
$x_n(\text{T}_\text{S})$ and variance
$(\text{T}_\text{F}-\text{T}_\text{S}) \eta_x^2$.

\subsection{Delay-Sensitive Learning Model}
Like the time-dependent drift model, the delay-sensitive
learning model assumes that decisions are based on comparing
the current stimulus to a stored referent. However, unlike
the time-dependent drift model, the criterion remains stable
over time and does not drift. For this reason, the memory
system used to store the criterion may be different from
working memory. In other words, $c_n(t)=c_n$ for all
$t>\text{T}_\text{S}$. This model also assumes that the
magnitude of error-driven updates to the criterion decreases
in proportion to the length of the feedback delay.  As a
result, when feedback is delayed, the system becomes less
responsive to errors, leading to slower learning compared to
immediate feedback conditions.

This model also assumes that the percept does not drift over
time. Instead, it models perceptual noise as a
time-invariant perturbation. Specifically, the
delay-sensitive learning model assumes the observer uses the
following decision rule:
\begin{equation}
  \text{Respond } R_n =
  \begin{cases}
    \text{A}, & \text{if ~} x_n(\text{T}_\text{S}) + \epsilon_p \leq c_n  \\
    \text{B}, & \text{if ~} x_n(\text{T}_\text{S}) + \epsilon_p > c_n,
  \end{cases}
  \label{eq:DR}
\end{equation}
where the time-invariant perceptual noise term $\epsilon_p$
is normally distributed with mean 0 and variances
$\sigma_p^2$. This same value of perceptual noise corrupts
all future values of the percept. Therefore, $x_n(t) =
x_n(\text{T}_\text{S}) + \epsilon_p$ for all $t >
\text{T}_\text{S}$.

The delay-sensitive learning model assumes that longer
feedback delays slow criteral learning. Specifically, the
model assumes that if negative feedback is received, the
criterion is updated as follows:
\begin{equation}
  c_{n+1} = c_n + \frac{\alpha}{t_\text{FD}} [x_n(\text{T}_\text{F}) - c_n].
  \label{eq:DSL_learning}
\end{equation}
The scale factor $(1/t_\text{FD})$ captures the notion that
the length of the feedback delay slows the \textit{rate} of
procedural learning.

\subsection{Reinforcement-Learning Model}
The reinforcement-learning model gradually associates
responses with stimuli, and therefore does not include an
explicit representation of the response criterion. The model
is based on an actor-critic architecture and uses
reinforcement learning to form stimulus-response
associations \parencite{SuttonBarto1998}.

This model assumes that the perceptual representation of
each stimulus is defined by the pattern of activation across
25 sensory units that are characterized by overlapping
tuning curves. Each unit is maximally excited by one
specific stimulus, which we call the unit's preferred
stimulus. Specifically, the activation of the $i^\text{th}$
sensory unit on trial $n$ is given by
\begin{equation}
  A_i(n) = \text{exp} \left( \frac{-\left[d_{i,S_n}\right]^2}{\sigma^2} \right)
\end{equation}
where $d_{i,S_n}$ is the (Euclidean) distance between the preferred
stimulus value of the $i^\text{th}$ sensory unit and the
value of the stimulus that was presented on trial $n$, and
$\sigma$ is a constant that increases with perceptual noise.

The model includes two decision or actor units -- one
associated with each of the two possible responses.
Initially, each sensory unit is connected to both actor
units with some random connection strength. Let
$\omega_{iJ}(n)$ denote the strength of the connection
between sensory unit $i$ and actor unit $J$ (for $J$ = A or
B) on trial $n$. Then the activation in actor unit $J$ on
trial $n$, denoted by $V_J(n)$, equals
\begin{equation}
  V_J(n) = \sum_{i} \omega_{iJ}(n) A_i(n)
\end{equation}
Responses are generated by the decision rule:
\begin{equation}
 \text{Respond } R_n =
  \begin{cases}
    A, & \text{if $V_A(n) > V_B(n)$}\\
    B, & \text{if $V_A(n) \leq V_B(n).$}
  \end{cases}
\end{equation}
The connection strengths $\omega_{iA}(n)$ and
$\omega_{iB}(n)$ are updated after feedback is received on
each trial according to standard reinforcement learning
rules \parencite{SuttonBarto1998}:
\begin{equation}
  \omega_{iA}(n) = \omega_{iA}(n-1) + \frac{\alpha_{actor}}{t_\text{FD}} \delta(n-1)
\end{equation}
\begin{equation}
  \omega_{iB}(n) = \omega_{iB}(n-1) + \frac{\alpha_{actor}}{t_\text{FD}} \delta(n-1),
\end{equation}
where $\alpha_{actor}$ is the learning-rate of the actor, and
$\delta(n-1)$ is the reward prediction error on trial $n-1$.
Note that as in the delay-sensitive learning model, the
learning rate is scaled by the inverse of the feedback
delay. Much evidence suggests that this type of
stimulus-response learning is mediated largely within the
striatum, and is facilitated by a dopamine (DA) mediated
reinforcement learning signal that is time dependent
\parencite[e.g.,][]{ValentinMaddoxAshby2014}. Specifically,
the dopamine signal generated by positive feedback appears
to peak at around 500 ms after feedback and then decay back
to baseline levels within 2 or 3 s
\parencite{YagishitaEtAl2014}. As a result, synaptic
plasticity at cortical-striatal synapses is attenuated with
increasing feedback delays \parencite{YagishitaEtAl2014}.
The scaling of $\alpha_{actor}$ by $1/t_\text{FD}$ models this
phenomenon.

The reward prediction error on trial $n-1$ is defined as the
value of the obtained reward, denoted by $R(n-1)$ minus the
value of the predicted reward, denoted by $P(n-1)$:
\begin{equation}
  \delta(n-1) = R(n-1) - P(n-1).
\end{equation}
The predicted reward on trial $n$ is determined by the
critic via
\begin{equation}
  P(n+1) = P(n) + \frac{\alpha_{critic}}{t_\text{FD}} \delta(n),
\end{equation}
where $\alpha_{critic}$ is the learning rate of the critic.
This learning rate is again scaled by the inverse of the
feedback delay.

\subsection{Simulation Results}
We investigated how changing the duration of the feedback
delay and ITI affect criterial learning for each of these
three models. More specifically, we simulated performance of
each model in a categorization task that included two
categories of stimuli that varied on a single stimulus
dimension and that could be categorized perfectly by
comparing each stimulus to an appropriate value of the
response criterion (We used the same experimental design as
in Experiment 1, so see those methods for more details). Our
simulations included three experimental conditions: Delayed
Feedback, Long ITI, and Control. The Delayed-Feedback
condition included a long feedback delay (3.5 s) and a short
ITI (0.5 s). The Long-ITI condition matched the total trial
duration of the Delayed-Feedback condition but with a short
feedback delay (0.5 s) and a long ITI (3.5 s). The Control
condition included both a short feedback delay (0.5 s) and a
short ITI (0.5 s). Each simulation continued for 200 trials
or until the model responded correctly for 12 trials in a
row. For each set of parameter values, we simulated the
model 100 times and averaged the results, yielding one
observed performance metric measured in trials-to-criterion
for each condition. All scores were then normalized by the
largest observed value, and as a result, all axes in the
figures that describe the simulation results range from zero
to one.

Our approach was to generate predictions for each of the
three experimental conditions across a wide range of
parameter values. We then used parameter-space partitioning
(PSP) to evaluate the performance of each model
\parencite[PSP;][]{pitt2006global}. PSP calculates the
proportion of the parameter space where a model makes
specific qualitative predictions. We focused on four such
predictions. The first was that learning under feedback
delay would be slower than in the other two conditions. The
second was that learning in the Long ITI condition would be
slower than in the other two. The third was that learning in
the control condition would be slower than in either of the
other two conditions. The fourth was a catch-all category
encompassing any other possible patterns of results. For
each model, the PSP analysis quantified the proportions of
the explored parameter space where the model made these
qualitative predictions.

\subsubsection{Time-dependent drift model}
We investigated predictions of this model across a wide
range of values for the parameters $\eta_{x}$, $\eta_{c}$,
and $\alpha$. Specifically, in the case of $\alpha$,  we
stepped through every value in the interval $[0, 1]$, with a
step size of .01. In the case of both $\eta_{x}$, and
$\eta_{c}$, we searched over the interval $[0.1, 5]$, with a
step size of 0.5. The results are shown in Figure
\ref{fig:TDD_results}.

\begin{figure} \centering
  \includegraphics[width=1\textwidth]{../figures/model_working_memory.pdf}
  \caption{ 
      \textbf{A:} The proportion of parameter space where
      the time-dependent drift model predicted each of four
      qualitative patterns: (1) slower learning under
      feedback delay, (2) slower learning in the Long-ITI
      condition, (3) slower learning in the control
      condition, and (4) any other pattern. 
      % 
      \textbf{B:} Simulated trials-to-criterion for each
      condition. All scores were normalized by the largest
      trials-to-criterion value within a given parameter
      set, so all axes range from zero to one. 
      % 
      \textbf{C:} Boxplot of the parameter ranges leading to
      each PSP pattern. All parameter values were normalized
      by the largest value in the search range, so the
      ordinate ranges from zero to one for all parameters.
      % 
      \textbf{D:} Scatter plot of the parameter ranges
      associated with each PSP pattern. 
      %
      \textit{Note:} In all panels, color indicates the PSP
      pattern.
}
  \label{fig:TDD_results}
\end{figure}

Figure \ref{fig:TDD_results}A shows that the model
essentially always predicts that increasing the feedback
delay or the ITI have similar detrimental effects on
learning. The model makes this prediction because it assumes
that the criterion drifts during both the feedback delay and
during the ITI. So the critical variable for the model is
the time between the response on trial $n$ and the
presentation of the stimulus on trial $n+1$. The model
assumes that the criterion drifts during this entire time,
and therefore how this interval is divided between feedback
delay and ITI is relatively unimportant to the model
predictions.

\subsubsection{Delay-sensitive learning model}
We simulated the performance of the delay-sensitive learning
across a wide range of parameter values for $\sigma_p$ and
$\alpha$. Specifically, in the case of $\alpha$ we stepped
through every value in the interval $[0, 1]$, with a step
size of .1. In the case of $\sigma_p$, we searched over the
interval $[0.1, 5]$, with a step size of 0.1.

The results are shown in Figure \ref{fig:DSL_results}. Note
that virtually all combinations of the parameter values
predict that feedback delay impairs criterial learning more
than increasing the ITI. This makes sense because Equation
\ref{eq:DSL_learning} shows that the delay-sensitive
learning model predicts that increasing the feedback delay
(i.e., increasing $t_{FD}$) will impair learning -- for any
value of $\alpha>0$. Large values of $\sigma_p$ will also
impair performance because of distortion to the percept, but
this interference will be the same in all conditions.

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{../figures/model_procedural_supervised.pdf}
  \caption{ 
      \textbf{A:} The proportion of parameter space where
      the delay-sensitive learning model predicted each of
      four qualitative patterns: (1) slower learning under
      feedback delay, (2) slower learning in the Long-ITI
      condition, (3) slower learning in the control
      condition, and (4) any other pattern. 
      % 
      \textbf{B:} Simulated trials-to-criterion for each
      condition. All scores were normalized by the largest
      trials-to-criterion value within a given parameter
      set, so all axes range from zero to one. 
      % 
      \textbf{C:} Boxplot of the parameter ranges leading to
      each PSP pattern. All parameter values were normalized
      by the largest value in the search range, so the
      ordinate ranges from zero to one for all parameters.
      % 
      \textbf{D:} Scatter plot of the parameter ranges
      associated with each PSP pattern. 
      %
      \textit{Note:} In all panels, color indicates the PSP
      pattern.
}
  \label{fig:DSL_results}
\end{figure}

\subsubsection{Reinforcement-learning model}
We investigated the effects on performance predicted by the
reinforcement-learning model of three parameters -- the
perceptual-noise variance $\sigma^2$ and the actor and
critic learning rates (i.e., $\alpha_{actor}$ and
$\alpha_{critic}$, respectively). In the case of both
$\alpha_{actor}$ and $\alpha_{critic}$, we stepped through
every value in the interval $[0, 0.2]$, with a step size of
.01. We constrained our search over these parameters to this
interval because reinforcement-learning models are prone to
instability at very high learning rates
\parencite{SuttonBarto1998}. As evidence of this, at higher
learning rates, the model failed to learn with any
consistency -- that is, in most cases, it failed to reach
the learning criterion (12 correct responses in a row)
within the allowable 200 trials. In the case of $\sigma$, we
searched over the interval $[1, 10]$, with a step size of 1.
As with the other models, all scores were normalized by the
largest observed value.

The results are described in Figure \ref{fig:RL_results}.
Note this model predicts that delayed feedback impairs
criterial learning more than a long ITI across a wide volume
of parameters settings. As can be seen in Figure
\ref{fig:RL_results}C, the only exceptions tend to occur
when the value of any of the three parameters approaches the
maximum end of the range explored. The logic here is similar
to the delay-sensitive learning model -- that is, both
models predict that increasing the feedback delay $t_{FD}$
will impair learning -- for all values of the learning rate.

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{../figures/model_procedural_reinforcement.pdf}
  \caption{ 
      \textbf{A:} The proportion of parameter space where
      the reinforcement-learning model predicted each of
      four qualitative patterns: (1) slower learning under
      feedback delay, (2) slower learning in the Long-ITI
      condition, (3) slower learning in the control
      condition, and (4) any other pattern. 
      % 
      \textbf{B:} Simulated trials-to-criterion for each
      condition. All scores were normalized by the largest
      trials-to-criterion value within a given parameter
      set, so all axes range from zero to one. 
      % 
      \textbf{C:} Boxplot of the parameter ranges leading to
      each PSP pattern. All parameter values were normalized
      by the largest value in the search range, so the
      ordinate ranges from zero to one for all parameters.
      % 
      \textbf{D:} Scatter plot of the parameter ranges
      associated with each PSP pattern. 
      %
      \textit{Note:} In all panels, color indicates the PSP
      pattern.
    }
  \label{fig:RL_results}
\end{figure}

\subsubsection{Summary of modeling results}
We investigated three criterial-learning models that make
different predictions about how increases in the feedback
delay and ITI affect learning. The time-dependent drift
model predicts that the criterion always drifts, so the
critical variable is the sum of the feedback delay and ITI.
How this sum is divided into separate feedback delay and ITI
time intervals has little effect on the model's predictions.
In contrast, the delay-sensitive learning model predicts
that feedback delays are necessarily more detrimental to
performance than increases in the ITI.  Finally, the
reinforcement-learning model also predicts that, in general,
feedback delays should impair learning more than long ITIs,
although this model is somewhat more flexible than the
delay-sensitive learning model and can account for a small
or null effect of feedback delay within a restricted region
of its parameter space. The obvious next question is how
human criterial learning is affected by these independent
variables. Experiments 2 and 3 were designed to address this
question, and therefore also to test the predictions of the
three models.

\section{Experiment 2}
Experiment 2 investigated how feedback delay and the length
of the ITI affect criterial learning in humans. As a result,
it also provides rigorous tests of the predictions of the
time-dependent drift model, the delay-sensitive learning
model, and the reinforcement-learning model.

The stimuli in Experiment 2 were circular sine-wave gratings
that varied across trials in bar width and bar orientation.
These stimuli were divided into two categories according to
their value on one of the two dimensions. In other words,
the optimal strategy was to set a response criterion on the
single relevant dimension, and then choose a categorization
response based on whether the value of the presented stimuli
on this dimension was larger or smaller than the criterion
value.

The experiment isolated criterial learning by (1) explicitly
instructing participants on the relevant stimulus dimension
and rule structure (e.g., thick bars are ``A'', thin bars
are ``B''), and (2) eliminating variability along the
irrelevant stimulus dimension. In other words, the
instructions identified the relevant stimulus dimension, and
as a result, the only learning required was criterial
learning.

\subsection{Method}

\subsubsection{Apparatus}
All experiments were performed in a dimly lit room.
Participants sat approximately 24'' from a 17'' $\times$
11'' monitor running at a resolution of 1680 $\times$ 1050
pixels. Participants made category judgments by pressing the
`d' or `k' keys on a standard computer keyboard for `A' or
`B' choices, respectively. Stickers with bold print `A' or
`B' were placed on the appropriate keys.

\subsubsection{Stimuli and Categories}
Stimuli were circular sine-wave gratings that varied in bar
width and bar orientation, drawn from various 1-dimensional
uniform distributions specific to the current category
problem. We first defined an arbitrary 2-dimensional
$[0-100,0-100]$ stimulus space, and then split each
dimension of this space into 7 bins of width 14 units each.
Each $(x,y)$ pair from this arbitrary stimulus space was
converted to a grating according to the nonlinear
transformations defined by
\textcite{treutwein1989perceptual}, which roughly equate the
salience of each dimension (for details, see also
\textcite{CrossleyAshby2015}).

The structure of the various criterial-learning tasks is
illustrated in Figure \ref{fig_design_exp_1_space}. Each
criterial-learning problem was created by first randomly
selecting a relevant dimension, and then randomly selecting
one of the 7 bins defined on that dimension. Each bin was
also associated with a corresponding unique value on the
irrelevant dimension. We buffered the to-be-learned response
criterion by 10\% of total bin width on either side with a
no-stimulus region. Random uniform samples from the
remaining eligible region of each bin were then selected and
presented to the participant until 9 correct responses out
of any 10 responses in a row advanced the participant to the
next problem. Note that every category problem was a simple
one-dimensional rule in which optimal accuracy was 100\%.
Note also that the relative location of the optimal response
criterion varied across problems.

\begin{figure}
  \centering
  \includegraphics[width=.5\textwidth]{../figures/fig_design_exp_1_space.png}
  \caption{
      Category sample space. Different colors represent
      different category problems. Dashed lines are category
      boundaries (criterion) and the surrounding solid lines
      mark the no-stimulus region in which no stimuli were
      sampled.
}
  \label{fig_design_exp_1_space}
\end{figure}

\subsubsection{Procedure}
There were three conditions (described in detail in Table
\ref{conditions_exp_1}). In the Delayed-Feedback condition,
feedback was delayed 3.5 s after the response and the ITI
was 0.5 s. In the Long-ITI condition, feedback was delayed
0.5 s after the response and the ITI was 3.5 s. Finally, in
the Control condition, feedback was delayed 0.5 s after the
response and the ITI was 0.5 s.

Each participant completed a series of one-dimensional
category-learning tasks or problems, which are described in
Figure \ref{fig_design_exp_1_space}. Each problem included
stimuli in two distinct clusters that varied over a
restricted range of the relevant dimension. Critically, the
optimal criterion value varied from problem-to-problem with
respect to its position within this range. For example, for
some problems the optimal criterion was below the midpoint
of the range and for other problems it was above the
midpoint.

Participants were explicitly told the relevant dimension for
each problem, as well as the generic response mapping (e.g.,
thick bars = ``A'', thin bars = ``B''). Figure
\ref{fig:trial} shows the structure of an example trial,
along with an example of a typical category structure. All
trials in every condition included a 500 ms fixation cross,
a response-terminated stimulus, a circular white-noise mask,
corrective feedback, and an inter-trial interval (ITI) that
varied according to condition. The text `Correct' was
displayed in centered, large green font after correct
responses, and the text `Incorrect' was displayed in
centered, large red font after incorrect responses.

Participants practiced each problem until they responded
correctly on 9 of the previous 10 trials. At this point, the
problem changed. Each participant completed as many problems
as possible in 512 trials or until they had been in the lab
for 60 minutes (including time to acquire consent and give
instructions), at which point the session was terminated.

\begin{figure}
  \centering
  \includegraphics[width=.5\textwidth]{../figures/fig_design_exp_1.png}
  \caption{
      Example trial and category problem. A) Events that
      occurred on each trial. B) An example of a typical
      category structure.
}
  \label{fig:trial}
\end{figure}

\begin{table}
    \caption{
        Durations (in s) of Trial Events in each Condition.
    }
    \label{conditions_exp_1}
    \begin{tabular}{c|cccc}
        Conditions & Stim & Mask & FB & ITI \\[0.5ex] \hline Control & RT & 0.5 &
        1.0 & 0.5 \\[0.5ex]
        \\[-1.5ex] Delayed Feedback &
        RT & 3.5 & 1.0 & 0.5 \\[0.5ex]   Long ITI & RT & 0.5 & 1.0 & 3.5
        \\[0.5ex]
    \end{tabular}
\end{table}

\subsubsection{Participants}
Fifty-nine participants participated in Experiment 2. All
were UCSB undergraduates and received course credit for
their participation. All had normal or corrected to normal
vision. We randomly assigned each participant to one of
three conditions (target $N>16$ per condition based on
similar previous research): Delayed Feedback ($N = 20$);
Long ITI ($N = 21$), Control ($ N = 17$).

\subsection{Results}
Figure \ref{fig_gmm_hist_1} shows a relative frequency
histogram of the trials-to-criterion observed across all
three conditions. The histogram shows that the majority of
participants were able to learn each problem on average in
less than 100 trials. However, this histogram also shows
that a subset of participants required many more trials to
learn each problem. Given that each problem is very simple
and that the relevant dimension of each problem is
explicitly instructed to participants, it is likely that the
participants in the tails of this distribution did not pay
attention to the instructions, were not motivated to learn
the task, were distracted by other factors, or should be
considered outliers for some other reason.

\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{../figures/fig_exp_1_gmm.png}
  \caption{
      Relative frequency histogram of the
      trials-to-criterion observed across all three
      conditions in Experiment 2. The black line are the
      predictions of the best-fitting one-component Gaussian
      Mixture Model to these data, whereas the red line
      represents the best fit of the two-component Gaussian
      Mixture Model. The two-component model provided a
      significantly better fit according to AIC, BIC, and a
      likelihood ratio test.
}
  \label{fig_gmm_hist_1}
\end{figure}

% AIC for 1 component: 735.7652217198266
% AIC for 2 components: 662.8095009936025
% 
% BIC for 1 component: 739.8861077409194
% BIC for 2 components: 673.1117160463345
%
% LRT statistic: 78.95572072622417
% Degrees of freedom: 3
% P-value: 5.140608761896481e-17

We modeled the distribution shown in Figure
\ref{fig_gmm_hist_1} using a Gaussian mixture model with two
components. The first component captured participants with a
relatively low mean number of trials-to-criterion, while the
second component identified outliers according to the above
rationale. We compared the fit of this two-component model
to a single-component model using the Akaike Information
Criterion (AIC) and Bayesian Information Criterion (BIC),
both of which indicated a better fit for the two-component
model (AIC: 662.81 vs. 735.77; BIC: 673.11 vs. 739.89).
Additionally, a likelihood ratio test confirmed that the
two-component model provided a significantly better fit than
the one-component model ($\chi^2(3) = 78.96$, $p < .001$).
Outlier participants were then excluded from further
analysis.  After making these exlcusions, we were left with
the following sample sizes: Control condition ($ N = 11$);
Delayed-Feedback condition ($N = 14$);  Long-ITI condition
($N = 17$).

Figure \ref{fig_exp_1_t2c}A shows the mean
trials-to-criterion in each condition of Experiment 2.  A
one-way ANOVA revealed a significant effect of condition
($F(2, 39) = 6.17$, $p < .01$, $\eta^2 = .24$) and planned
comparisons revealed that performance in the
Delayed-Feedback condition was significantly worse than in
either the Control condition ($t(23.00) = 2.70$, $p < .05$)
or the Long-ITI condition ($t(20.96) = 2.94$, $p < .01$).
Performance in the Control and Long-ITI conditions did not
differ significantly from each other ($t(17.99) = 0.22$, $p
= .83$).

%   Source  ddof1  ddof2         F     p-unc       np2
% 0    cnd      2     39  6.171331  0.004693  0.240398
%
%           A          B         T        dof     p-unc
% 0     Delay   Long ITI  2.938217  20.959026  0.007864
% 1     Delay  Short ITI  2.701083  22.997274  0.012748
% 2  Long ITI  Short ITI  0.219862  17.987044  0.828454

Figure \ref{fig_exp_1_t2c}B shows the mean number of
problems solved in each condition of Experiment 2. A one-way
ANOVA revealed a significant effect of condition ($F(2, 39)
= 5.77$, $p < .01$, $\eta^2 = .23$). Planned comparisons
revealed that performance in the Delayed-Feedback condition
was significantly worse than in the Control condition
($t(18.54) = -3.28$, $p < .01$) and worse than in the Long
ITI condition ($t(21.60) = -2.14$, $p < .05$).  Performance
in the Control and Long-ITI conditions did not differ
significantly from each other ($t(25.97) = -1.50$, $p=.15$).

%   Source  ddof1  ddof2         F     p-unc       np2
% 0    cnd      2     39  5.766358  0.006399  0.228223
% 
%           A          B         T        dof     p-unc
% 0     Delay   Long ITI -2.143777  21.590392  0.043579
% 1     Delay  Short ITI -3.277500  18.536603  0.004061
% 2  Long ITI  Short ITI -1.501591  25.965920  0.145267

\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{../figures/fig_exp_1_t2c.png}
    \caption{
        \textbf{A}: Mean trials to criterion in each
        condition of Experiment 2. Error bars are standard
        errors of the mean.
        %
        \textbf{B}: Mean number of problems solved in each
        condition of Experiment 2. Error bars are standard
        errors of the mean.
}
  \label{fig_exp_1_t2c}
\end{figure}

\section{Experiment 3}
Experiment 3 was designed to reinforce the findings of
Experiment 2 by using the same design but with stimuli that
have binary-valued dimensions, thereby eliminating criterial
learning. Unlike Experiment 2, we did not explicitly
instruct participants on the relevant stimulus dimension in
Experiment 3.  While Experiment 2 isolated criterial
learning from rule selection and rule switching, this
experiment isolates rule selection and switching from
criterial learning. Therefore, Experiment 3 was designed to
test whether the impaired learning that we observed in
Experiment 2 when the feedback was delayed can be attributed
principally to criterial learning or whether it could be due
to some more general category-learning process.

\subsection{Method}

\subsubsection{Apparatus}
The apparatus was the same as in Experiment 2.

\subsubsection{Stimuli and Categories}
The stimuli consisted of colored geometric figures presented
on a colored background. These varied across six binary
dimensions: the number of items (either one or two), the
size of the items (small or large), the color of the items
(yellow or blue), the shape of the items (circle or square),
the texture of the background (smooth or rough), and the
orientation of the background (horizontal or tilted by 20
degrees). This combination resulted in a total of 64 unique
stimuli ($2^6$). An example trial and stimuli are shown in
Figure \ref{fig_design_exp_2}. In all conditions, the order
of stimulus presentation was fully randomized for each
participant, and the relevant dimension for each problem was
selected randomly without replacement from the set of six
dimensions.

\begin{figure}
  \centering
  \includegraphics[width=.5\textwidth]{../figures/fig_design_exp_2.png}
  \caption{
      \textbf{A:} An example trial from Experiment 3.
      \textbf{B:} Eight example stimuli (out of 64 total
      stimuli) from a possible one-dimensional rule problem.
      In this case the correct rule is `A' if the shape
      color is yellow and `B' if the shape color is blue.
      The solid line represents the category boundary.
}
  \label{fig_design_exp_2}
\end{figure}

\subsubsection{Procedure}
Participants received detailed instructions about the task.
However, unlike in Experiment 2, they were not informed of
the relevant dimension or the correct rule. The trial timing
for the Delayed-Feedback and Long-ITI conditions in
Experiment 3 were the same as in Experiment 2. Experiment 3
did not include a condition that was analogous to the
Control condition of Experiment 2. Participants practiced
each problem until they responded correctly on 12
consecutive trials. At this point, the problem changed. Each
participant completed as many problems as possible in 600
trials or until they had been in the lab for 30 minutes
(including time to acquire consent and give instructions),
at which point the session was terminated.

\subsubsection{Participants}
Thirty-four participants participated in Experiment 3. All
were Macquarie University undergraduates and received course
credit for their participation. All had normal or corrected
to normal vision. We randomly assigned each participant to
one of two conditions: Delayed Feedback ($N = 17$) or Long
ITI ($N = 17$).

\subsection{Results}
Figure \ref{fig_gmm_hist_2} shows a relative frequency
histogram of the trials-to-criterion observed in both
conditions. The histogram shows that the majority of
participants were able to learn each problem on average in
less than 100 trials.  Unlike Experiment 2, there were no
highly suspicious outliers in this data set. Nevertheless,
for symmetry with Experiment 2, we performed the same
Gaussian mixture model analysis as used there. The
two-component model provided a slightly smaller AIC value
(317.40 vs. 317.78), but a slightly larger BIC value (325.03
vs. 320.83). The likelihood-ratio test was not significant
($\chi^2(3) = 6.38$, $p = .09$). We therefore did not
exclude any participants from further analysis in Experiment
2.

\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{../figures/fig_exp_2_gmm.png}
  \caption{
      Relative frequency histogram of the
      trials-to-criterion observed across all three
      conditions in Experiment 3. The black line represents
      the best fit of the one-component Gaussian Mixture
      Model, and the red line represents the best fit of the
      two-component Gaussian Mixture Model.
}
  \label{fig_gmm_hist_2}
\end{figure}

% AIC for 1 component: 317.77535652645616
% AIC for 2 components: 317.3995958509652
% 
% BIC for 1 component: 320.8280775756885
% BIC for 2 components: 325.03139847404606
% 
% LRT statistic: 6.3757606754909375
% Degrees of freedom: 3
% P-value: 0.09469309595624516

Figure \ref{fig_exp_2_t2c}A shows the mean
trials-to-criterion in each condition of Experiment 3. An
independent-samples t-test revealed no significant effect of
condition ($t(32.0)=0.69$, $p = .50$, $\eta^2 = .01$).  The
mean number of problems solved by each participant is shown
in Figure \ref{fig_exp_2_t2c}B. An independent-samples
t-test revealed no significant effect of condition ($t(32) =
-0.49$, $p = .63$, $\eta^2 = .01$).

%   Source  ddof1  ddof2        F     p-unc       np2
% 0    cnd      1     32  0.47046  0.497715  0.014489
% 
%        A          B         T   dof     p-unc
% 0  delay  immediate  0.685901  32.0  0.497715

%   Source  ddof1  ddof2         F    p-unc       np2
% 0    cnd      1     32  0.235294  0.63093  0.007299
% 
%        A          B         T   dof    p-unc
% 0  delay  immediate -0.485071  32.0  0.63093

\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{../figures/fig_exp_2_t2c.png}
    \caption{
        \textbf{A}: Mean trials to criterion in each
        condition of Experiment 3. Error bars are standard
        errors of the mean.
        %
        \textbf{B}: Mean number of problems solved in each
        condition of Experiment 3. Error bars are standard
        errors of the mean.
}
  \label{fig_exp_2_t2c}
\end{figure}

\subsection{Discussion of Experiments 2 and 3}
Experiment 2 clearly showed that a short feedback delay of
3.5 s slowed criterial learning. In contrast, increasing the
ITI to this same value had no effect on learning. The task
we used was one-dimensional category learning, but we
isolated criterial learning by instructing participants
about the optimal strategy. Specifically, we told them which
stimulus dimension was relevant and that there was no
trial-by-trial variability on the irrelevant dimension.
Although these results strongly suggest that the feedback
delay had its interfering effects on criterial learning,
Experiment 3 was designed to confirm this inference. The
goal here was to examine the effects of the same feedback
delays and ITIs on performance in a one-dimensional
category-learning task that did not require any criterial
learning, but did require category-learning processes that
are thought to mediate rule discovery (e.g., rule selection
and switching). If the feedback delay effects observed in
Experiment 2 were acting on some general category-learning
skill then we should have seen the same interfering effects
of feedback delay in Experiment 3. However, if the
feedback-delay interference of Experiment 2 was operating
selectively on criterial learning, then it should disappear
in Experiment 3, since no criterial learning was required.
Our results strongly supported this latter prediction.
Therefore, Experiments 2 and 3 together strongly suggest
that criterial learning is impaired by feedback delays and
is relatively unaffected by the length of the ITI.

Two prior studies investigated similar issues. First, Ell
and colleagues reported that feedback delay and also a
concurrent memory-scanning task each impaired rule-based
category learning \parencite{ell2009critrial}. Based on
these results, they hypothesized that criterial learning may
be tied to working-memory capacity and therefore to explicit
cognitive mechanisms. However, criterial learning was
confounded with rule selection and task difficulty in their
design.  Furthermore, \textcite{ell2009critrial} found that
feedback delay only impaired learning when working memory
demand was high -- that is, when participants had to learn
more than one response criterion for optimal performance. In
contrast, we found that feedback delay impairs learning even
when working memory demands are trivial (participants never
have to keep in mind more than one criterion).

Second, \textcite{bohil2014implicit} reported that the
effects of unequal base rates on criterion placement in
rule-based category learning were diminished under delayed
feedback and also under an observational training protocol
that is also thought to selectively impair basal
ganglia-dependent associative-learning mechanisms. These
results are consistent with our findings. However, because
they were obtained by manipulating base rates, it is
difficult to conclude that the delay affected criterial
learning, \emph{per se}. For example, consider a simple
two-stage model in which the first stage learns the base
rates and the second stage uses what the first stage learned
to adjust the response criterion appropriately. The
\textcite{bohil2014implicit} results are also consistent
with the hypothesis that the feedback delay impaired the
first of these stages but not the second. Our results
strongly suggest that feedback delays impair criterial
learning (and therefore the second of these hypothetical
stages).

\section{General Discussion}
We developed three novel models of criterial learning that
were designed to explore how feedback delay and ITI duration
affect criterial learning. Two of these models assumed that
decisions are made by comparing the current percept to a
stored referent, or criterion, and that the remembered value
of this criterion is updated following error feedback via a
gradient-descent learning rule. The first of these models,
the time-dependent drift model, posits that both the
criterion and the percept drift over time, leading to
similar impairments when feedback is delayed or the ITI is
increased. The second model, the delay-sensitive learning
model, assumes that although neither the percept nor the
criterion drift over time, optimal criterial learning
requires immediate feedback. As a result, this model
predicts that increasing the feedback delay will slow the
rate of criterial learning. The third model, the
reinforcement-learning model, assumes that criterial
learning arises through the gradual formation of
stimulus-response associations, rather than via the
construction of a response criterion. This model also
assumes that the rate at which these associations are
learned is slowed by feedback delays.

Simulations of these models in a task that was structurally
identical to Experiment 2 revealed that the time-dependent
drift model typically predicts that the most important
variable for criterial learning is the total amount of time
between the response and the stimulus presentation that
defines the next trial. Where the feedback is presented
within this interval is relatively unimportant. In contrast,
both the delay-sensitive learning model and the
reinforcement-learning model consistently predict that
feedback delays should impair performance more than a long
ITI.

The experiments were designed to test these predictions. Our
results strongly suggested that human criterial learning is
sensitive to feedback delay but not to the ITI duration.
This suggests that the delay-sensitive learning model and
the reinforcement-learning model provide a better account of
human behavior across wide ranges of their respective
parameter spaces than the time-dependent drift model.

The time-dependent drift model was based on the idea that
criterial learning might be entirely supported by working
memory. The assumption of drift in this model stems from the
understanding that maintaining items in working memory is
inherently challenging.  The longer an item is held, the
more likely it is to deteriorate. Therefore, the model
predicts that performance should deteriorate as the time
between the response on trial $n$ and stimulus presentation
on trial $n+1$ increases, regardless of how this interval is
divided between feedback delay and ITI. The results of
Experiments 1 and 2 strongly disconfirm this prediction.

In contrast, the delay-sensitive learning model was inspired
by the idea that a response criterion might be encoded in a
more stable memory system. One appealing candidate for this
function is the cerebellum, where synaptic plasticity has
been shown to follow a gradient descent learning rule and
also to be sensitive to feedback timing
\parencite{brudner2016delayed, held_adaptation_1966,
honda_adaptation_2012, kitazawa_effects_1995,
kitazawa_prism_2002}.  Finally, the reinforcement-learning
model draws on the idea that criterial learning could emerge
through stimulus-response association learning driven by
dopamine-dependent synaptic plasticity in the striatum. This
process aligns with established theories of category
learning that posit a key role to this learning process in
the basal ganglia \cite{AshbyCOVIS1998}.

Our results do not strongly favor the delay-sensitive
learning model over the reinforcement learning model or vice
versa. The critical disagreement between these accounts is
whether the criterion has any real psychological meaning. In
the delay-sensitive learning model it does, and thus this
model is psychologically similar to the time-dependent drift
model -- the main disagreement being about whether the
representation of the criterion is stable and if the
updating of the criterion is sensitive to feedback delay.

The reinforcement-learning model makes very different
psychological assumptions. This model assumes that behaviors
described as ``criterial learning'' are actually mediated by
the learning of stimulus-response associations. According to
this account, the criterion has no internal representation
and therefore no psychological meaning. Testing between
these two very different accounts of criterial learning
should be a focus of future research.

We are not aware of any data that directly addresses the
question of whether a response criterion is a fundamental
psychological construct, as the delay-sensitive learning
model suggests, or whether it is unnecessary, as assumed by
the reinforcement learning model. Even so, there are some
results in the category-learning literature that support the
interpretation of the reinforcement-learning model. In
information-integration (II) category-learning tasks, the
optimal strategy is similarity-based, and difficult or
impossible to describe verbally
\parencite[e.g.,][]{AshbyValentin2018}. When the stimuli
vary on two dimensions, the stimuli from contrasting
categories can be partitioned by a decision bound that is
conceptually similar to a response criterion. In both cases,
all stimuli on one side are associated with one response,
and all stimuli on the other side are associated with the
contrasting response. Furthermore, in agreement with
Experiment 2, II category learning is impaired by short
feedback delays \parencite{MaddoxAshbyBohil2003,
MaddoxIng2005}. The analogous question in II category
learning is whether the decision bound is learned directly
or whether it is simply the set of points that divide the
perceptual space into contrasting response regions.  In
fact, the evidence strongly supports this latter
interpretation \parencite{AshbyWaldron1999, CasaleEtAl2012}.
For example, if the decision bound is learned, then it
should be possible to apply this bound to novel stimuli.
With rule-based categories, this is easy for participants,
but with II categories there is no evidence that the
response strategy that participants learn can be generalized
to novel stimuli \parencite{CasaleEtAl2012} -- a result that
strongly supports the hypothesis that the decision bound has
no psychological meaning.

Our results clearly demonstrate that criterial learning is
impaired by delayed feedback, and not by extending the
intertrial interval. These results are consistent with the
hypothesis that criterial learning is a form of
basal-ganglia mediated associative learning, and are
inconsistent with hypotheses that criterial learning is a
working-memory-based process. Thus, our results provide a
critical constraint on future models of rule-based
classification and decision making, and possibly also on
more general accounts of criterion setting, such as in
signal detection theory.

Previous studies have failed to find that feedback delays
impair rule-based category learning, and on the face of it,
this seems to contradict our finding that feedback delays
impair criterial learning. However, all earlier RB studies
that looked for effects of a feedback delay, either used
binary-valued stimulus dimension and so no criterial
learning was needed, or else set the response criterion
exactly midway between the category prototypes, which makes
criterial learning trivial. For example, under these
conditions, criterial learning might not even require
feedback. The unsupervised category-learning experiments
reported by \textcite{ashby1999dominance} provide strong
support for this because all of their rule-based
participants learned the correct criterion (which was midway
between the category means), even though the task was
completely unsupervised. Moreover, all previous studies
failed to isolate criterial learning, so even if there was
an effect of feedback delay on criterial learning, it could
have been masked by larger effects caused by other
rule-learning processes.

Criterial learning is among the most classic and ubiquitous
of all cognitive skills. For example, signal-detection
theory teaches that it is the central form of learning in an
enormous range of decision-making tasks -- everything from
simple YES-NO detection of a weak signal to assessing the
guilt or innocence of a defendant in a jury trial. Our
results suggest that even in simple rule-based tasks,
criterial learning seems to be subserved, at least in part,
by associative mechanisms. Most current theories tend to
classify tasks as either executive function (e.g.,
rule-based category learning) or procedural (e.g., mirror
tracing). Our results suggest that such classification
schemes might oversimplify how humans perform these tasks,
and therefore that much more work is needed to understand
how different learning and memory systems interact.

% TODO
\section{Transparency and Openness}
All data have been made publicly available at the [trusted
repository name] and can be accessed at [persistent URL or
DOI].

\section{Author Notes}
Preparation of this article was supported by Public Health
Service Grant MH2R01-063760.

\printbibliography

\end{document}

### 2024-10-20 code/imports.py

@@ -1,0 +1,10 @@
+import os
+import glob
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+import pingouin as pg
+from sklearn.mixture import GaussianMixture
+from scipy.stats import chi2
+import string

### 2024-10-20 code/inspect_results_behave.py

@@ -1,0 +1,15 @@
+from util_func import *
+from imports import *
+
+d = load_data()
+
+inspect_duration_vs_trials(d)
+
+dd = d.groupby(['exp', 'cnd', 'sub'])[['t2c', 'nps']].mean().reset_index()
+print(dd.groupby(['exp', 'cnd'])['sub'].nunique())
+
+dd1 = dd.loc[dd['exp'] == 'Exp 1'].copy()
+dd2 = dd.loc[dd['exp'] == 'Exp 2'].copy()
+
+report_exp_1(dd1)
+report_exp_2(dd2)

### 2024-10-20 code/inspect_results_models.py

@@ -1,0 +1,23 @@
+from experiment_imports import *
+import model_working_memory
+import model_procedural_supervised
+import model_procedural_reinforcement
+
+# model_working_memory.fit('../data/delay/', *(3.0, 0.5))
+# model_working_memory.fit_validate_ppt()
+
+# model_working_memory.fit()
+# model_procedural_supervised.fit()
+# model_procedural_reinforcement.fit()
+
+# model_working_memory.psp()
+# model_procedural_supervised.psp()
+# model_procedural_reinforcement.psp()
+
+model_working_memory.plot_psp()
+model_procedural_supervised.plot_psp()
+model_procedural_reinforcement.plot_psp()
+
+# model_working_memory.fit_validate()
+# model_procedural_supervised.fit_validate()
+# model_procedural_reinforcement.fit_validate()

### 2024-10-20 code/inspect_validate.py

@@ -1,0 +1,9 @@
+import numpy as np
+import scipy as sp
+import pandas as pd
+from ggplot import *
+
+d = pd.read_csv('fit_validate.csv', header=0)
+
+ggplot(d, aes(x=problem, y=alpha_fit)) +\
+    geom_point()

### 2024-10-20 code/model_procedural_supervised.py

@@ -1,0 +1,297 @@
+from experiment_imports import *
+
+
+def simulate(xc_init, xc_true, alpha, noise, delay, iti, stim):
+
+    xc = xc_init
+    num_consecutive_correct = 0
+    t2c = 0
+
+    for x in stim:
+
+        # percept is corrupted by perceptual noise
+        noise_err = np.random.normal(0, noise, 1)
+        x += noise_err
+
+        if x < xc_true:
+            cat = 1
+        else:
+            cat = 2
+
+        if x - xc < 0:
+            resp = 1
+        else:
+            resp = 2
+
+        if cat != resp:
+            xc += (1 / delay) * (alpha) * (x - xc)
+            num_consecutive_correct = 0
+        else:
+            num_consecutive_correct += 1
+
+        t2c += 1
+
+        if num_consecutive_correct >= 12:
+            break
+
+    return (t2c)
+
+
+def fit(fit_params, delay, iti, human_data):
+
+    xc_init = fit_params[0]
+    alpha = fit_params[1]
+    noise = fit_params[2]
+
+    # human_data must be a dict with keys: 'cat', 'x', 'y', 'resp'
+    num_trials = len(human_data['x'])
+    stim_range_x = max(human_data['x']) - min(human_data['x'])
+    stim_range_y = max(human_data['y']) - min(human_data['y'])
+    stim_range = max([stim_range_x, stim_range_y])
+
+    xc = xc_init
+    L = []
+
+    for trial in range(num_trials):
+
+        # get current stimulus
+        if stim_range_x > stim_range_y:
+            x = human_data['x'][trial]
+        else:
+            x = human_data['y'][trial]
+
+        # percept is corrupted by perceptual noise
+        noise_err = np.random.normal(0, noise, 1)
+        x += noise_err
+
+        # compute the liklihood of observing the human response given the current model
+        L.extend(abs(x - xc) / float(stim_range))
+
+        if human_data['cat'][trial] != human_data['resp'][trial]:
+            xc += (1 / delay) * (alpha) * (x - xc)
+
+    return (np.prod(L))
+
+
+def fit_validate():
+    # generate fake data from the model using simulate()
+    # find whether or not fit() recovers parameters
+    # find constraints on when fit() performs well and not
+    pass
+
+
+def fit_humans():
+    # load human data
+    # iterate through per human per problem
[…]

### 2024-10-20 code/experiment_imports.py
initial commit

@@ -1,0 +1,37 @@
+import csv
+import time
+import os
+import sys
+import random
+import math
+import itertools
+import numpy as np
+import pandas as pd
+from scipy.optimize import minimize, brute, fmin, differential_evolution
+import scipy.stats as sps
+import multiprocessing as mp
+import pandas as pd
+from itertools import product
+import matplotlib.pyplot as plt
+from matplotlib import colors, cm
+from mpl_toolkits.mplot3d import Axes3D
+from matplotlib.lines import Line2D
+from matplotlib import rcParams
+from matplotlib import rc
+import matplotlib as mpl
+import matplotlib.colors as mcolors
+import seaborn as sns
+
+# mpl.use('pgf')
+# plt.rcParams.update({
+#     "pgf.texsystem": "lualatex",
+#     "font.family": "serif",
+#     "text.usetex": True,
+#     "pgf.rcfonts": False,
+#     "pgf.preamble": [
+#         "\\usepackage{fontspec}",
+#         "\\usepackage{amsmath,amsfonts,amssymb}",
+#         "\\usepackage{gensymb}",
+#         r"\setmainfont{Arial}"
+#          ]
+# })

### 2024-10-20 code/model_procedural_reinforcement.py

@@ -1,0 +1,533 @@
+from experiment_imports import *
+
+
+def simulate(xc_true, vis_width, alpha_critic, alpha_actor, delay, iti, stim):
+
+    data_record = {'cat': [], 'x': [], 'y': [], 'resp': []}
+
+    num_consecutive_correct = 0
+    t2c = 0
+
+    # static params:
+    dim = 25
+    w_base = 0.5
+    w_noise = 0.01
+    vis_amp = 1.0
+
+    # buffers
+    vis_act = np.zeros(dim)
+    w_A = np.zeros(dim)
+    w_B = np.zeros(dim)
+
+    # init
+    v = 0.5
+
+    # init weights: Note that we assume fresh weights every problem
+    w_A = np.random.normal(w_base, w_noise, dim)
+    w_B = np.random.normal(w_base, w_noise, dim)
+
+    for x in stim:
+
+        # Determine true category label
+        if x < xc_true:
+            cat = 1
+        else:
+            cat = 2
+
+        # compute input activation via radial basis functions
+        vis_dist_x = 0.0
+        for i in range(0, dim):
+            vis_dist_x = x - i
+            vis_act[i] = vis_amp * math.exp(-(vis_dist_x**2) / vis_width)
+
+        # Compute A and B unit activations via dot product
+        act_A = 0.0
+        act_B = 0.0
+        act_A = np.inner(vis_act, w_A)
+        act_B = np.inner(vis_act, w_B)
+
+        # Compute resp via max
+        discrim = act_A - act_B
+        resp = 1 if discrim > 0 else 2
+
+        # Implement strong lateral inhibition
+        act_A = act_A if resp == 1.0 else 0.0
+        act_B = act_B if resp == 2.0 else 0.0
+
+        # compute outcome
+        r = 1.0 if cat == resp else 0.0
+
+        # compute prediction error
+        delta = (r - v)
+
+        # update critic
+        v += (1 / delay) * alpha_critic * delta
+
+        # update actor
+        for i in range(0, dim):
+
+            if delta < 0:
+                w_A[i] += (
+                    1 /
+                    delay) * alpha_actor * vis_act[i] * act_A * delta * w_A[i]
+                w_B[i] += (
+                    1 /
+                    delay) * alpha_actor * vis_act[i] * act_B * delta * w_B[i]
+            else:
+                w_A[i] += (
+                    1 / delay) * alpha_actor * vis_act[i] * act_A * delta * (
+                        1 - w_A[i])
[…]

### 2024-10-20 code/model_working_memory.py

@@ -1,0 +1,624 @@
+from experiment_imports import *
+
+
+def simulate(x, *args):
+    alpha = x[0]
+    eta_stim = x[1]
+    eta_crit = x[2]
+
+    delay = args[0]
+    iti = args[1]
+
+    num_consecutive_correct = 0
+    t2c = 0
+    max_trials = 200
+
+    simulated_data = {
+        'cat': [],
+        'x': [],
+        'y': [],
+        'rsp': [],
+        'pe': [],
+        'prob': []
+    }
+
+    num_bins = 7
+    bin_width = 14
+    bin_bounds = np.arange(0, 100, bin_width)
+    bins = np.zeros((num_bins, 2))
+    for i in range(bin_bounds.shape[0] - 1):
+        bins[i, 0] = bin_bounds[i]
+        bins[i, 1] = bin_bounds[i + 1]
+
+    bin_dim = np.arange(0, num_bins, 1)
+    np.random.shuffle(bin_dim)
+
+    for prob in range(num_bins):
+        x_range = bins[bin_dim[prob]]
+        xc_true = np.random.uniform(x_range[0], x_range[1])
+        lb1 = x_range[0]
+        ub1 = xc_true - 0.1 * bin_width
+        lb2 = xc_true + 0.1 * bin_width
+        ub2 = x_range[1]
+        xc = np.random.uniform(x_range[0], x_range[1])
+
+        for i in range(max_trials):
+
+            # sample stimulus for current trial
+            x1 = np.random.uniform(lb1, ub1)
+            x2 = np.random.uniform(lb2, ub2)
+            x = np.random.choice([x1, x2])
+
+            # memory for the criterion drifts during the ITI
+            crit_noise = np.random.normal(0, 1)
+            xc += eta_crit * iti * crit_noise
+
+            if x < xc_true:
+                cat = 1
+            else:
+                cat = 2
+
+            if x - xc < 0:
+                rsp = 1
+            else:
+                rsp = 2
+
+            # the criterion and the stimulus drift over the feedback delay
+            stim_drift = np.random.normal(0, 1)
+            crit_drift = np.random.normal(0, 1)
+            x += eta_stim * delay * stim_drift
+            xc += eta_crit * delay * crit_drift
+
+            if cat != rsp:
+                xc += alpha * (x - xc)
+                num_consecutive_correct = 0
+            else:
+                num_consecutive_correct += 1
+
+            t2c += 1
+
[…]

### 2024-10-20 code/util_func.py

@@ -1,0 +1,323 @@
+from imports import *
+
+
+def load_data():
+    d1 = load_data_exp_1()
+    d2 = load_data_exp_2()
+
+    d1['exp'] = 'Exp 1'
+    d2['exp'] = 'Exp 2'
+
+    d1 = d1[['exp', 'sub', 'cnd', 't', 't_prob', 'rt']]
+    d2 = d2[['exp', 'sub', 'cnd', 't', 't_prob', 'rt']]
+
+    d = pd.concat([d1, d2], ignore_index=True)
+
+    d['experiment_duration'] = d.groupby(['cnd', 'sub'])['rt'].transform('sum')
+    d['experiment_duration'] = d['experiment_duration'] * 5 / 60
+    d['trials_completed'] = d.groupby(['cnd', 'sub'])['t'].transform('max')
+
+    d['prob_num'] = d.groupby(
+        ['cnd',
+         'sub'])['t_prob'].transform(lambda x: (x < x.shift(1)).cumsum() + 1)
+    d['t2c'] = d.groupby(['cnd', 'sub',
+                          'prob_num'])['t_prob'].transform('count')
+    d['nps'] = d.groupby(['cnd', 'sub'])['prob_num'].transform('max')
+
+    return d
+
+
+def load_data_exp_1():
+
+    dir_delay = '../data/delay'
+    dir_immed = '../data/immed'
+    dir_maip = '../data/maip'
+
+    filenames_delay = [
+        os.path.join(dir_delay, f) for f in os.listdir(dir_delay)
+    ]
+    filenames_immed = [
+        os.path.join(dir_immed, f) for f in os.listdir(dir_immed)
+    ]
+    filenames_maip = [os.path.join(dir_maip, f) for f in os.listdir(dir_maip)]
+
+    col_names = ['t', 't_prob', 'bnd', 'cat', 'x', 'y', 'rsp', 'rt']
+
+    d_delay = [
+        pd.read_csv(f, sep='\t', header=None, names=col_names)
+        for f in filenames_delay
+    ]
+    d_immed = [
+        pd.read_csv(f, sep='\t', header=None, names=col_names)
+        for f in filenames_immed
+    ]
+    d_maip = [
+        pd.read_csv(f, sep='\t', header=None, names=col_names)
+        for f in filenames_maip
+    ]
+
+    for idx, df in enumerate(d_delay):
+        df['sub'] = idx + 101
[…]

### 2025-06-11 write/response_to_reviews.tex
generic

\documentclass[12pt]{article}
% Language and font
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
% Bibliography with APA style
\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
\DeclareLanguageMapping{american}{american-apa}
\addbibresource{sample.bib}
% Layout and formatting
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\setstretch{1.0}
\usepackage{parskip} % disables indentation and adds vertical space between paragraphs
\setlength{\parindent}{0pt}
% Math and symbols
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{gensymb}
% Graphics and floats
\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
% Optional: nicer font
\usepackage{lmodern}
% Turn off section numbers
\setcounter{secnumdepth}{0}

\title{Response to Reviews: \\ \vspace{1cm}
    Criterial Learning and Feedback Delay: Insights from
Computational Models and Behavioral Experiments}
\author{
    Matthew J. Crossley, 
    Benjamin O. Pelzer,
    F. Gregory Ashby
}
\date{}

\begin{document}

\maketitle

\section{Reviewer 1}

\subsection{Reviewer comment}
The research question of the study is to investigate the
effect of delay in criterion learning in category learning.
Experiment 1 shows that in a continuous criterion learning
task with known relevant dimension delaying the feedback but
not increasing the ITI decreases performance. In other
words, delaying feedback makes it more difficult to learn
the position of a continuous criterion. Experiment 2 uses
many binary dimensions in the category learning task and
participants' goal is to learn the binary dimension. In this
task, delaying feedback does not decrease performance. The
researchers simulate the performance of three different
models and show that a model that assumes the criterion is
maintained in working memory with information drifting
across time cannot predict the observed effect. Two models
that make different assumption can predict the observed
pattern.

\subsection{Author response}
...

\subsection{Reviewer comment}
I am no expert in category learning, so cannot adequately
judge the novelty of the observed results. Given this
caveat, I really enjoyed reading the manuscript and felt it
made a convincing argument using a rigorous and
state-of-the-art methodology. Using parameter-space
partitioning is an excellent and sadly underutilised way of
comparing models. It goes much beyond the typical approach
of comparing models in terms of their model fit and instead
allows deriving qualitative differences between the
candidate models. In addition, the two experiments are well
designed and executed and provide the necessary empirical
basis on which the models are compared. The paper is also
overall clearly written. Taken together, I highly recommend
this paper to be published after addressing a number of
mostly minor concerns detailed below. My main issues with
the current version relate to the presentation of the models
and modelling results and whether the space of possible
models is sufficiently explored.

\subsection{Author response}
...

\subsection{Reviewer comment}
1. Exploration of model space: When reading the description
of the models it was not entirely clear to me whether the
three presented models explore the space of possible models
to a sufficient degree or whether there might be some models
missing from consideration that could change the overall
message. For example, from a theoretical point of view the
main difference between the time-dependent drift model and
the delay-sensitive learning model is that for the latter
"the criterion remains stable over time and does not drift".
However, when looking at the model descriptions more aspects
seem to differ than solely the presence of drift, namely the
presence of the feedback delay in the updating rule (Eq. 2
vs. 6). Is the change in the updating rule necessary for the
models to just differ in the noise in the criterion or is it
part of it? I am not sure if I just do not understand the
difference or if a hybrid model with drift and feedback
delay in the updating rule is possible. Likewise, are any of
the assumption in the reinforcement-learning model, which
seems to differ quite a bit from the previous models,
somewhat arbitrary or are these always the only possible
choices?

I am not asking for an extensive exploration of the model
space using additional simulations, but some descriptions of
the extent to which the presented models explore the
possible model space or if there are other possible models
that could potentially lead to different results is needed.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
2. Simulation results figures: There were some aspects of
the results figures for the simulation that I found
difficult to understand:

(a) Even after reading the full paper and thinking about it
for a while, I do not understand what panels B are supposed
to show. Does this only show the subset of the parameter
space in which one of the two relevant patterns was
observed? If not, shouldn't there be data points everywhere?

(b) I do not understand why in Figure 1, for example, panel
C shows some cases in which the predicted qualitative
pattern is either "control impaired" or "other", but there
is no corresponding bar in panel A.

(c) The bars in Figure 2A are smaller than in Figures 1A and
3A. Why?

\subsection{Author response}
...

\subsubsection{Reviewer comment}
3. Calling the simulation Experiment 1: I found it extremely
confusing that the simulation using parameter-space
partitioning was called "Experiment 1". If I read
"Experiment" I expect to see some empirical data obtained
from participants and not solely a simulation. I suggest
renaming this section to something clearer. For example
"Simulation Study" or "Parameter-Space Partitioning Across
Models".

\subsection{Author response}
...

\subsubsection{Reviewer comment}
4. Outliers in Exp. 2:

- While I think the argument for removing outliers in Exp. 2
is fine, the actual criterion defining an outlier is not
actually given (which feels a bit ironic in a paper about
criteria).

- How do the results look with these outliers included? Does
the difference between condition vanish? Adding a footnote
or so should be enough.

- The fact that some participants did not even learn a
single criterion feels very surprising to me. Is this common
in such experiments? Please add a sentence embedding this
finding within your general experience running such
experiments.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
Minor points:

- Does Eq. 13 show results for trial n or n + 1? The text
and equation are inconsistent.

- The description regarding the randomisation of Experiment
2 on page 12 (paragraph before "Procedure") seems to be
inconsistent with Figure 4. Is the dimension picked at
random for each trial or always the same for certain problem
numbers (as implied by Figure 4)?

- In figure 4, the top says Problems "1 thru 13", but only
shows 7 problems.

\subsection{Author response}
...

\section{Reviewer 2}

\subsubsection{Reviewer comment}
Your action editor asked me if I could provide a review, as
they have only been able to secure one reviewer after a
considerable delay. Although I am editor in chief for the
journal, I am acting in the role of a reviewer in this case.
The action editor retains full authority to determine the
decision outcome for this manuscript.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
This manuscript reports a sophisticated approach to an
important problem. I especially like the use of mathematical
models.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
I was concerned that the models seem to have multiple
varying features. It might be more useful to isolate
specific processes while holding other model elements
constant. Based on the introduction, I was not clear on the
theoretical motivation for comparing the time-dependent
drift and delay sensitive learning models, as the latter did
not seem to be just a version of the former but with delay
as a relevant factor. Some additional justification and/or
adjustments to the tested set of models is needed.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
Despite these concerns about which models are the most
relevant to compare, I was impressed by the simulation
strategy and the focus on finding “signature” predictions
that distinguish different pairs of models.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
I think it makes more sense to have a prediction simulation
section and Experiments 1 and 2 rather than to consider the
simulation the first experiment.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
The data in Figure 6 can likely be fit much more closely by
an ExGaussian distribution than either the 1 or 2 component
Gaussian mixtures that are shown.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
Replication the criterial learning results from the
currently-labeled Experiment 2 seems like a good idea given
the fairly low sample sizes per condition.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
Did stimuli remain on the screen across the feedback delay?
I don’t think so. If not, what happens if the stimulus
reappears at the time of feedback? I could imagine this
would eliminate or dramatically reduce the negative effect
of feedback delay. If so, is that consistent with the
conclusions? That is, would the models that predict a
deleterious effect of feedback delay also predict that it
should be eliminated with “reminder” stimuli at the time of
feedback?

\subsection{Author response}
...

\subsubsection{Reviewer comment}
For the currently-labeled Experiment 3, reporting
fail-to-reject results for t-tests is not a valid method for
establishing that feedback delay has no effect. This is
especially a concern given the low participant count per
condition and the fact that empirical effects go in the
“right” direction (lower performance for delayed feedback).
A different inferential approach is needed, and one
possibility is reporting confidence intervals over effect
size. Your conclusions rely on demonstrating that you can
rule out meaningfully large effects (i.e., only negligibly
small effect sizes are included in the confidence
intervals). I suspect that the intervals will be wide and
include a range of fairly big effects, of the same magnitude
as the effects reported in Experiment 1. As such, this
experiment does not establish that the effects of feedback
delay are specific to a criterion learning task.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
I recommend running a bigger replication study that combines
both tasks to demonstrate the claimed interaction (feedback
delay impairs performance only for the criterion learning
task). The replication study should be preregistered with a
sample size justification based on the interaction effect
defined by the across-experiment comparison in the current
data. Adding a “stimulus reminder at feedback” condition to
this new study might be helpful in clarifying the mechanisms
producing the delay effect.

Signed,
Jeff Starns

\subsection{Author response}
...

\section*{Action Editor}

\subsubsection{Reviewer comment}
The topic is made very clear in the first paragraph and
notes the need for investigating how criterions are learned
in decision models.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
The final paragraph of the introduction summarizes the
experiments. This seems more appropriate in the abstract,
given that no information about the experiments are
presented in the abstract.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
Equation 1 needs some explanation, as this implies that the
decision Rn is made immediately on stimulus presentation. In
other words, Rn is not an outcome of some time-extended
cognitive process. Am I misunderstanding equation 1?

\subsection{Author response}
...

\subsubsection{Reviewer comment}
What was the reason behind the choice to have equation 2
only be sensitive to negative feedback? How is "optimal"
defined in the current case, given that response time does
not factor in the update rule?

\subsection{Author response}
...

\subsubsection{Reviewer comment}
The use of parameter space partitioning is sensible given
the many differences among the three models. However, PSP
operates on non-stochastic models. In addition, the way you
have approached PSP is not how the original PSP is
implemented (i.e., using MCMC). This should be mentioned
explicitly to avoid confusion.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
The three-dimensional plots in figures 1B, 2B, and 3B do not
come out right on 2D. You could improve these by having 2D
projections on the planes. Better still, three boxplots
would be more appropriate as there is no additional
information by using 3D plots.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
Although not a focus in this manuscript, I wondered whether
the models are able to account for finer-grained data
patterns such as complete RT distributions and
speed-accuracy tradeoff functions.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
The experiments are numbered 2 and 3, instead of 1 and 2.
The models were following the methods of experiment 1, which
is missing. Experiment 1 is also mentioned in the discussion
separately with experiment 2. It looks like experiment 1 has
been deleted from this manuscript, but the model simulations
and discussion still refer to it.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
In figure 6, it is clear that the participants form two
subcohorts. However, the use of a Gaussian seems odd given
that there can not be any negative values for
trials-to-criterion. Given that the authors model this
frequency histogram to exclude participants, more
appropriate distributions, such as Poisson should be used.

\subsection{Author response}
...

\subsubsection{Reviewer comment}
Given the experimental results, what is the added benefit of
the three models? Experiment 2 is the critical experiment
showing that 3.5-0.5 leads to more trials-to-criterion than
0.5-3.5 or 0.5-0.5. The models do not provide any detailed
understanding of mechanisms and they are not fitted to
actual data. The general message from the experiments is
that any updating rule should be sensitive to feedback
delay. The models do not go beyond this.

\subsection{Author response}
...

\end{document}

### 2025-06-11 write/cover_letter.tex
generic

\documentclass[10pt]{article}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{geometry}
\geometry{
  top=20mm,
}
\usepackage{fancyhdr}
\lhead{\includegraphics[height=2.0cm]{../figures/MQ_MAS_HOR_RGB_POS.png}\vspace{0.0cm}}
\rhead{\includegraphics[width=6.0cm]{../figures/UC_Santa_Barbara_Wordmark_Navy_RGB.png}\vspace{0.5cm}}
\setlength\headheight{2cm}

\usepackage[hang, symbol]{footmisc}
\setlength\footnotemargin{10pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\usepackage{hyperref}

\pagenumbering{gobble}

\renewcommand{\thefootnote}{\dag}

\begin{document}
\thispagestyle{fancy}
Dear Editors,

I am pleased to submit our manuscript titled ``Criterial
Learning and Feedback Delay: Insights from Computational
Models and Behavioral Experiments'' for consideration in the
\textit{Journal of Experimental Psychology: Learning,
Memory, and Cognition}.

In this work, we develop and evaluate three novel
computational models designed to explore the mechanisms
underlying criterial learning. We test the predictions
of these models in two behavioral experiments, focusing on
the effects of feedback delay and intertrial intervals (ITI)
on criterial learning in both models and humans.

Our results indicate that human criterial learning is
significantly impaired by delayed feedback, but not by
extended ITIs. The computational models suggest that these
results align most naturally with mechanisms based on
procedural or associative learning. This indicates that
even in simple rule-based tasks, criterial learning may
be driven by associative processes.

Given the importance of criterial learning across various
decision models and the current gaps in understanding its
cognitive and neural mechanisms, we believe this work will
be of broad interest to the journal's readership.

For expert evaluation, we recommend the following reviewers
who possess expertise in criterial learning, category
learning, and computational modeling:

\begin{itemize}
    \item Dr. Carol A. Seger; \href{mailto:Carol.Seger@colostate.edu}{Carol.Seger@colostate.edu}
    \item Dr. S\'ebastien H\'elie; \href{mailto:shelie@purdue.edu}{shelie@purdue.edu}
    \item Dr. Corey Bohil \href{mailto:cbohil@ltu.edu}{cbohil@ltu.edu}
\end{itemize}

We affirm that this manuscript is original, has not been
previously published, and is not under concurrent
consideration elsewhere. All data and materials necessary
for reproducing the reported results are available in our
GitHub repository:
\url{https://github.com/crossley/crit_learn_delay}.

We appreciate your consideration of our manuscript for
publication in \textit{JEP: Learning, Memory, and Cognition}
and look forward to your feedback.

Sincerely,

Matthew J. Crossley\footnote{Corresponding author: matthew.crossley@mq.edu.au} \\
matthew.crossley@mq.edu.au \\
School of Psychological Sciences \\
Macquarie University Performance and Expertise Centre \\
Australian Hearing Hub \\
16 University Ave \\
Macquarie University, NSW 2109, Australia \\

Benjamin O. Pelzer \\
neurobot01@gmail.com \\
Independent Researcher \\
J\"ulich, North Rhine-Westphalia, Germany \\

F. Gregory Ashby \\
ashby@psych.ucsb.edu \\
Department of Psychological and Brain Sciences \\
251 University of California Santa Barbara \\
Santa Barbara, CA 93106, USA \\

\end{document}

### 2025-06-11 write/main.tex
generic

@@
     {Macquarie University Performance and Expertise Research Centre, Macquarie University, Sydney, Australia},
-    {Nanyang Technological University},
+    {Independent Researcher},
     {Department of Psychological \& Brain Sciences, University of California, Santa Barbara}
@@
 
-\keywords{response criterion; criterial learning;
-associative learning; categorization; procedural learning}
+\keywords{response criterion; criterial learning; associative learning; categorization; procedural learning}
 
@@
 similar previous research): Delayed Feedback ($N = 20$);
-Long ITI ($N = 21$), Control ($ N = 17$).
+Long ITI ($N = 21$), Control ($ N = 17$). All participants
+gave written informed consent before participating in the
+study. All experimental protocols were approved by the
+University of California at Santa Barbara Human Subjects
+Committee in the Office of Research Development and
+Administration.
 
@@
 one of two conditions: Delayed Feedback ($N = 17$) or Long
-ITI ($N = 17$).
+ITI ($N = 17$).  All participants gave written informed
+consent before participating in the study. All experimental
+protocols were approved by the Macquarie University Human
+Research Ethics Committee (protocol number: 52020339922086).
 
@@
 
-% TODO
 \section{Transparency and Openness}
-All data have been made publicly available at the [trusted
-repository name] and can be accessed at [persistent URL or
-DOI].
+All data have been made publicly available in the following
+GitHub repository:
+\url{https://github.com/crossley/crit_learn_delay}
 

### 2025-06-11 code/experiment_imports.py

@@ -1,4 +1,3 @@
-import csv
 import time
 import os
 import sys
@@ -10,7 +9,6 @@ import pandas as pd
 from scipy.optimize import minimize, brute, fmin, differential_evolution
 import scipy.stats as sps
 import multiprocessing as mp
-import pandas as pd
 from itertools import product
 import matplotlib.pyplot as plt
 from matplotlib import colors, cm

### 2025-06-11 code/model_new.py

@@ -1,0 +1,370 @@
+from experiment_imports import *
+
+"""
+Model class I: No criterion is learned
+- perceptual drift (y/n)
+- criterial drift (NA)
+- delay-sensitive updating (y/n)
+
+Model class II: Criterion is learned
+- perceptual drift (y/n)
+- criterial drift (y/n)
+- delay-sensitive updating (y/n)
+"""
+
+
+def simulate_class_I(params, args):
+
+    # Unpack params into sigma_perceptual_noise, alpha_actor, and alpha_critic
+    sigma_perceptual_noise = params[0]
+    alpha_actor = params[1]
+    alpha_critic = params[2]
+    eta_perceptual_drift = params[3]
+    delay_sensitive_update = params[4]
+
+    # Unpack args into xc_true, t_delay, t_iti, and stim
+    xc_true = args[0]
+    t_delay = args[1]
+    t_iti = args[2]
+    stim = args[3]
+
+    if delay_sensitive_update:
+        t_delay_actor = t_delay
+        t_delay_critic = t_delay
+    else:
+        t_delay_actor = 1
+        t_delay_critic = 1
+
+    num_consecutive_correct = 0
+    t2c = 0
+
+    # static params
+    dim = 25
+    w_base = 0.5
+    w_noise = 0.01
+    vis_amp = 1.0
+
+    # buffers
+    vis_act = np.zeros(dim)
+    w_A = np.zeros(dim)
+    w_B = np.zeros(dim)
+
+    # init
+    v = 0.5
+
+    # init weights: Note that we assume fresh weights every problem
+    w_A = np.random.normal(w_base, w_noise, dim)
+    w_B = np.random.normal(w_base, w_noise, dim)
+
+    for x in stim:
+
+        # Determine true category label
+        if x < xc_true:
+            cat = 1
+        else:
+            cat = 2
+
+        # compute input activation via radial basis functions
+        vis_dist_x = 0.0
+        for i in range(0, dim):
+            vis_dist_x = x - i
+            vis_act[i] = vis_amp * math.exp(-(vis_dist_x**2) / sigma_perceptual_noise)
+
+        # Compute A and B unit activations via dot product
+        act_A = 0.0
+        act_B = 0.0
+        act_A = np.inner(vis_act, w_A)
+        act_B = np.inner(vis_act, w_B)
+
+        # Compute resp via max
+        discrim = act_A - act_B
+        resp = 1 if discrim > 0 else 2
+
[…]

### 2025-09-11 write/response_to_reviews.tex
new figures etc

@@
 \subsection{Author response}
-...
+We thank the reviewer for their careful reading of our
+manuscript and their thoughtful comments. We address each of
+the reviewer's points in turn below.
 
@@
 \subsection{Author response}
-...
+We are grateful for the reviewer's positive assessment of
+our work and for their constructive suggestions.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for raising this important point. We
+have used this as an opportunity more systematically explore
+the classes of possible models.  In particular, we explore
+two classes of models: one class that assumes a criterion is
+learned and one class that assumes SR learning with no
+criterion.  Within the criterial-learning class we explore
+three different binary assumptions: perceptual drift (yes or
+no), criterial drift (yes or no), and delay-sensitive
+updating (yes or no).  Within the SR learning class there is
+no criterion to drift, but we do explore perceptual drift
+and delay-sensitive feedback (yes or no on each).
 
@@
 \subsection{Author response}
-...
+With the new approach to model exploration, we have heavily 
+revised the simulation results figures.  We hope that the
+new figures are clearer and address the reviewer's concerns.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for this suggestion. We have made this
+change in the revision.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for the opportunity to clarify our
+method. Outliers were identified on the basis of the Gaussian
+mixture model fit to the trials-to-criterion data. We now
+say ``Participants that were deemed to most likely belong in
+the low-performance Gaussian by this analysis were excluded
+from further analysis.''
 
@@
 - The description regarding the randomisation of Experiment
-2 on page 12 (paragraph before "Procedure") seems to be
+2 on page 12 (paragraph before ``Procedure'') seems to be
 inconsistent with Figure 4. Is the dimension picked at
@@
 \subsection{Author response}
-...
+We thank the reviewer for catchign these points. All are
+corrected or revised for greater clarity in the revision.
 
@@
 \subsection{Author response}
-...
+We thank the editor in chief for acting as a reviewer and for
+their careful reading of our manuscript and their thoughtful
+comments. We address each of the editor's points in turn
+below.
 
@@
 \subsection{Author response}
-...
+We are glad the reviewer viewed our approach favourably.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for raising this important point. In
+the revision we more systematically explore the classes of
+possible models.  In particular, we explore two classes of
+models: one class that assumes a criterion is learned and
+one class that assumes SR learning with no criterion.
+Within the criterial-learning class we explore three
+different binary assumptions: perceptual drift (yes or no),
+criterial drift (yes or no), and delay-sensitive updating
+(yes or no).  Within the SR learning class there is no
+criterion to drift, but we do explore perceptual drift and
+delay-sensitive feedback (yes or no on each). The motivation
+for these models and parameter choices is now more fully
+explained in the revision.
 
@@
 relevant to compare, I was impressed by the simulation
-strategy and the focus on finding “signature” predictions
+strategy and the focus on finding ``signature'' predictions
 that distinguish different pairs of models.
@@
 \subsection{Author response}
-...
+We are glad the reviewer viewed our simulation strategy
+favourably.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for this suggestion. We now estimate
+both a single-component truncated ExGaussian and a
+two-component truncated ExGaussian mixture via MLE with
+proper normalization on the tasks bounded support $[9,
+512]$, and compared them via AIC, BIC, and a parametric
+bootstrap likelihood-ratio test (LRT).  The results from
+moving from the  mixture of Gaussians used in our orginal
+submission to the truncated ExGaussian mixture used in the
+revision does not change the qualitative pattern of our
+results in any way. However, we agree that the ExGaussian
+and ExGaussian mixture better capture the skewed and bounded
+distribution of our trials-to-criterion data so we have
+adopted it in the revision.
 
@@
 \subsection{Author response}
-...
+We agree with the reviewer that a replication would be
+valuable but feel the data and models currently presented
+represent a suitable contribution to the field as is. For
+this reason, paired with the time and resource costs of
+conducting a replication, we have not pursued this in the
+revision.
 
@@
 deleterious effect of feedback delay also predict that it
-should be eliminated with “reminder” stimuli at the time of
+should be eliminated with ``reminder'' stimuli at the time of
 feedback?
@@
 \subsection{Author response}
-...
+We thank the reviewer for the opportunity to clarify the
+design of our experiment. As shown in Fig. 5 and in Fig. 8
+the stimulus was presented up until a response was made. At
+this time the stimulus was removed and replaced by a noise
+mask until feedback was delivered. We also thank the
+reviewer for raising the interesting question of whether the
+negative effect of feedback delay would be eliminated if the
+stimulus were to reappear at the time of feedback. We agree
+that this would be an interesting manipulation to explore in
+future work. However, we believe that their are many
+interesting followup questions that could be explored in
+future work and that the current manuscript represents a
+suitable contribution to the field as is.
 
@@
 condition and the fact that empirical effects go in the
-“right” direction (lower performance for delayed feedback).
+``right'' direction (lower performance for delayed feedback).
 A different inferential approach is needed, and one
@@
 \subsection{Author response}
-...
+TODO: Greg, I'd appreciate your take on this comment.
+
+I ran an equivalence test, and as the reviewer suspected,
+the data are inconclusive. I also have Experiment 2 data
+from an earlier semester. It shows the same pattern (though
+with fewer participants and more noise), but I left it out
+to avoid mixing cohorts. Just out of curiosity, I combined
+it with the current data and reran the analysis, but the
+results remain inconclusive. So I don’t think we can make a
+strong claim that Exp 2 shows no effect of delay --- at
+least not with this method.
+
+I also tried testing whether the delay effect is larger in
+Exp 1 than Exp 2 (a difference-of-differences approach), but
+that too was inconclusive.
+
+One possible way to address this is to add data from another
+archived experiment, where we examined criterion learning
+with a numerical Stroop dual task. Those data (n=30) showed
+no effect of the dual task. Including this might add some
+empirical value.
+
+Thoughts?
 
@@
 defined by the across-experiment comparison in the current
-data. Adding a “stimulus reminder at feedback” condition to
-this new study might be helpful in clarifying the mechanisms
-producing the delay effect.
+data. Adding a ``stimulus reminder at feedback'' condition
+to this new study might be helpful in clarifying the
+mechanisms producing the delay effect.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for their careful reading of our
+manuscript and their thoughtful comments overall. Here, we
+agree that a replication / extension study would be valuable
+but feel the paper as is represents a suitable contribution
+to the field.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for their positive assessment of our
+work.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for this suggestion. We have added
+more detail about the experiments to the abstract, but have
+elected to retain the summary of the experiments in the
+introduction as we feel it helps orient the reader to the
+structure of the paper.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for raining this point and for the
+opportunity to clarify. In essence the reviewer is correct.
+Equation 1 indicates that the response is made immediately
+upon stimulus presentation. We therefore are not explicitly
+modelling the time-extended cognitive processes that govern
+evidence accumulation and the corresponding response time.
+We now say this explicitly in the revision.
 
@@
 What was the reason behind the choice to have equation 2
-only be sensitive to negative feedback? How is "optimal"
+only be sensitive to negative feedback? How is ``optimal''
 defined in the current case, given that response time does
@@
 \subsection{Author response}
-...
+We thank the reviewer for the opportunity to clarify this
+point. The rationale behind the choice to have equation 2
+only be sensitive to negative feedback is that if the
+response is correct, then the observer has effectively
+gained zero information about how their criterion should be
+modified. Also, in this case otimal refers to the true
+criterion value that the observer is trying to learn. We now
+say this explicitly in the revision.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for raising this important point. We
+have revised the manuscript accordingly.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for this suggestion. We have heavily
+revised the simulation results figures. We hope that the new
+figures are clearer and address the reviewer's concerns.
 
@@
 \subsection{Author response}
-...
+We agree that these are interesting questions. The models in
+their current form do not account for RT distributions or
+speed-accuracy tradeoff functions. However, we believe that
+this could be a fruitful avenue for future work.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for catching this confusion. No study
+was deleted from the manuscript. Rather, whether or not the
+computational modelling section should be considered an
+``experiment'' was a point of oscillation when writing. We
+hope our revision clarifies this point.
 
@@
 \subsection{Author response}
-...
+We thank the reviewer for this suggestion. We now estimate
+both a single-component truncated ExGaussian and a
+two-component truncated ExGaussian mixture via MLE with
+proper normalization on the tasks bounded support $[9,
+512]$.
 
@@
 \subsection{Author response}
-...
+Thank you for raising this but we respectfully disagree that
+our models do not provide a detailed understanding of
+mechanism. To the contrary, our models all offer detailed
+mechanistic accounts for how criterial learning may occur in
+cognition. They articulate that delay effects may arise from
+any of several distinct cognitive mechanisms including the
+decay of a memory representation over time, the discounting
+of learning magnitude by feedback delay of this
+representation, or the discounting of learning in a
+reinforcement-learning stimulus-response framework. In this
+way, the models contribute the most important piece: they
+pin down the process-level mechanism consistent with the
+experiments.
 

### 2025-09-11 write/main_revision.tex
new figures etc

--- 2025-06-11 write/main.tex
@@
     }
-\abstract{The notion of a response criterion is ubiquitous
-    in psychology, yet its cognitive and neural
-    underpinnings remain poorly understood. To address this
-    shortcoming, three computational models that capture
-    different hypotheses about criterial learning were
-    developed and tested. The time-dependent drift model
-    assumes the criterion is stored in working memory and
-    that its value drifts over time. The delay-sensitive
-    learning model assumes that the magnitude of criterial
-    learning is temporally discounted by feedback delay. The
+
+\abstract{
+    The notion of a response criterion is ubiquitous in
+    psychology, yet its cognitive and neural underpinnings
+    remain poorly understood. To address this shortcoming,
+    three computational models that capture different
+    hypotheses about criterial learning were developed and
+    tested. The time-dependent drift model assumes the
+    criterion is stored in working memory and that its value
+    drifts over time. The delay-sensitive learning model
+    assumes that the magnitude of criterial learning is
+    temporally discounted by feedback delay. The
     reinforcement-learning model assumes that criterial
@@
     reinforcement-learning models predicted impairments only
-    with feedback delays. Two behavioral experiments, which
-    tested these predictions, showed that human criterial
-    learning is impaired by delayed feedback but not by long
-    ITIs.  These results support the delay-sensitive and
-    reinforcement-learning models, and suggest that even in
-    tasks that appear to rely on explicit, rule-based
+    with feedback delays.  Two behavioral experiments tested
+    these predictions. In Experiment 1, participants were
+    explicitly instructed about the relevant dimension to
+    isolate criterial learning, and short feedback delays
+    impaired learning relative to immediate feedback. In
+    Experiment 2, using binary-feature stimuli without
+    instruction (isolating rule selection/switching),
+    neither longer ITIs nor delayed feedback affected
+    performance. These results support the delay-sensitive
+    and reinforcement-learning models, and suggest that even
+    in tasks that appear to rely on explicit, rule-based
     reasoning, criterial learning may have strong
@@
 
-\section{Experiment 1: The Models}
-We developed three computational models, each representing
-different architectures of criterial learning, and examined
-their sensitivity to the duration of ITIs and feedback
-delay. Two of the models assume that the criterion is stored
-in memory and that response decisions are made by comparing
-the current stimulus-driven percept to this stored referent.
-Both of these models further posit that the internal
-representation of the criterion is updated following errors
-using a simple gradient-descent rule. The third model
-assumes that no criterion is stored nor used to generate
-responses. Instead, reinforcement learning is used to form
-stimulus-response associations and these associations drive
-responding without any appeal to the notion of a criterion.
-
-For each of the three models, denote the time of stimulus
+\section{New Computational Models of Criterial Learning}
+We developed computational models of criterial learning
+within two broad architectures and examined their
+sensitivity to ITI duration and feedback delay. The
+architectures differ in whether they assume that the
+response criterion is explicitly represented in memory. In
+the first architecture, the criterion is stored in memory,
+and responses are generated by comparing the current percept
+with this stored referent. The criterion is updated after
+errors using a gradient-descent rule. Models in this class
+can exhibit sensitivity to ITI and feedback delay in three
+distinct ways: (1) perceptual drift, where stimulus
+representations change over time, (2) criterial drift, where
+the stored criterion shifts over time, and (3)
+delay-sensitive updating, where the magnitude of criterion
+updates decreases with longer feedback delays. In the second
+architecture, no explicit criterion is stored.  Instead,
+reinforcement learning forms direct stimulus--response
+associations that drive performance.  Sensitivity to ITI and
+feedback delay arises from either perceptual drift or
+delay-sensitive updating, but criterial drift is not
+possible since no criterion is represented.
+
+For each of these models, denote the time of stimulus
 presentation on the current trial by T$_\text{S}$, the
@@
 that begins the next trial is displayed by T$_{\text{S}^+}$.
-Then note that the feedback delay equals
-$\text{t}_\text{FD} = \text{T}_\text{F} - \text{T}_\text{R}$
-and the duration of the ITI equals $\text{t}_\text{ITI} =
-\text{T}_{\text{S}^+} - \text{T}_\text{F}$.
-
-\subsection{Time-Dependent Drift Model}
-The time-dependent drift model assumes that the observer
-constructs a criterion, holds this value in working memory,
-and then makes response decisions by comparing the percept
-to the criterion. The model also assumes that both the
-memory representation of the criterion and the perceived
-value of the stimulus gradually drift over time. The model
-further assumes that the extent of this drift increases with
-time, both within a trial and between consecutive trials.
+Then note that the feedback delay equals $\text{t}_\text{FD}
+= \text{T}_\text{F} - \text{T}_\text{R}$ and the duration of
+the ITI equals $\text{t}_\text{ITI} = \text{T}_{\text{S}^+}
+- \text{T}_\text{F}$.
+
+\subsection{Models that assume a stored criterion}
+These models assume that the observer constructs a response
+criterion, maintains it in working memory, and makes
+decisions by comparing the percept to this criterion. They
+may further posit that both the stored criterion and the
+perceived stimulus value drift over time, with the extent of
+drift increasing both within a trial and across consecutive
+trials. In addition, they may assume that the learning rate
+for updating the criterion decreases as the feedback delay
+increases.
 
@@
 \end{equation}
+Note that this equation indicates that the response is made
+immediately upon stimulus presentation. We therefore are not
+explicitly modelling the time-extended cognitive processes
+that govern evidence accumulation and the corresponding
+response time.
+
 If positive feedback is received, then the value of the
 criterion remains unchanged for the next trial (except for
-drift -- see Equation 3). If negative feedback is received,
-then the criterion is modified according to the standard
-model \parencite{SuttonBarto1998}:
+drift -- see Equation \ref{eq:criterion}).  The rationale
+here is that if the response is correct, then the observer
+has effectively gained zero information about how their
+criterion should be modified.  If negative feedback is
+received, then the criterion is modified according to the
+standard model
+\parencite{SuttonBarto1998}:
 \begin{equation}
-c_n(\text{T}_\text{F}+ \Delta_t) = c_n(\text{T}_\text{F}) + \alpha [x_n(\text{T}_\text{F}) - c_n(\text{T}_\text{F})],
+c_n(\text{T}_\text{F}+ \Delta_t) = c_n(\text{T}_\text{F}) + \frac{\alpha}{t_{FD}} [x_n(\text{T}_\text{F}) - c_n(\text{T}_\text{F})],
 \label{RL}
@@
 trials \parencite[e.g.,][]{Ashby2017}. This updating rule
-will gradually converge on the optimal criterion value.
-Since this model assumes that criterial learning relies on
-working memory -- and the available evidence suggests that
-logical reasoning and working memory are unaffected by
-feedback delays of several seconds \parencite[e.g., in
-one-dimensional rule-based category learning
-tasks;][]{ell2009critrial, MaddoxAshbyBohil2003,
-MaddoxIng2005} -- we assume that the learning process
-described in Equation \ref{RL} is likewise unaffected by
-feedback delay.
-
-Although the criterial-learning process described by
-Equation \ref{RL} is not affected by feedback delay, we
-assume that both the stimulus and the criterion
-representations drift randomly throughout the duration of
-time they are maintained in working memory. The
+will gradually converge on the optimal criterion value
+(i.e., the true criterion value the observer is trying to
+learn).
+
+The term $\frac{\alpha}{t_{FD}}$ captures the notion that
+the magnitude of the update decreases as the feedback delay
+increases.  As a result, when feedback is delayed, the
+system becomes less responsive to errors, leading to slower
+learning compared to immediate feedback conditions. However,
+since this model is consistent with the idea that criterial
+learning relies on working memory -- and the available
+evidence suggests that logical reasoning and working memory
+are unaffected by feedback delays of several seconds
+\parencite[e.g., in one-dimensional rule-based category
+learning tasks;][]{ell2009critrial, MaddoxAshbyBohil2003,
+MaddoxIng2005} -- the learning process described in Equation
+\ref{RL} likewise be unaffected by feedback delay. In this
+case, the term $\frac{\alpha}{t_{FD}}$ is replaced by
+$\alpha$, making the learning rate independent of feedback
+delay.
+
+This model further assumes that both the stimulus and the
+criterion representations may drift randomly throughout the
+duration of time they are maintained in working memory. The
 representation of the criterion must always be maintained in
@@
 affects performance up until the feedback is presented, but
-not afterwards.
-
-We modeled the drift in both the criterion and the percept
-by adding white noise to their initial values. Specifically,
-we assumed that for all $t>\text{T}_\text{S}$
+not afterwards. We modeled the drift in both the criterion
+and the percept by adding white noise to their initial
+values. Specifically, we assumed that for all
+$t>\text{T}_\text{S}$
 \begin{equation}
@@
 are parameters that determine the amount of drift over time.
+Note that these parameters are permitted to be equal to zero
+which is equivalent to assuming that the criterion and/or
+the percept do not drift over time at all.
+
 This model predicts that at the time of feedback, the
@@
 
-\subsection{Delay-Sensitive Learning Model}
-Like the time-dependent drift model, the delay-sensitive
-learning model assumes that decisions are based on comparing
-the current stimulus to a stored referent. However, unlike
-the time-dependent drift model, the criterion remains stable
-over time and does not drift. For this reason, the memory
-system used to store the criterion may be different from
-working memory. In other words, $c_n(t)=c_n$ for all
-$t>\text{T}_\text{S}$. This model also assumes that the
-magnitude of error-driven updates to the criterion decreases
-in proportion to the length of the feedback delay.  As a
-result, when feedback is delayed, the system becomes less
-responsive to errors, leading to slower learning compared to
-immediate feedback conditions.
-
-This model also assumes that the percept does not drift over
-time. Instead, it models perceptual noise as a
-time-invariant perturbation. Specifically, the
-delay-sensitive learning model assumes the observer uses the
-following decision rule:
-\begin{equation}
-  \text{Respond } R_n =
-  \begin{cases}
-    \text{A}, & \text{if ~} x_n(\text{T}_\text{S}) + \epsilon_p \leq c_n  \\
-    \text{B}, & \text{if ~} x_n(\text{T}_\text{S}) + \epsilon_p > c_n,
-  \end{cases}
-  \label{eq:DR}
-\end{equation}
-where the time-invariant perceptual noise term $\epsilon_p$
-is normally distributed with mean 0 and variances
-$\sigma_p^2$. This same value of perceptual noise corrupts
-all future values of the percept. Therefore, $x_n(t) =
-x_n(\text{T}_\text{S}) + \epsilon_p$ for all $t >
-\text{T}_\text{S}$.
-
-The delay-sensitive learning model assumes that longer
-feedback delays slow criteral learning. Specifically, the
-model assumes that if negative feedback is received, the
-criterion is updated as follows:
-\begin{equation}
-  c_{n+1} = c_n + \frac{\alpha}{t_\text{FD}} [x_n(\text{T}_\text{F}) - c_n].
-  \label{eq:DSL_learning}
-\end{equation}
-The scale factor $(1/t_\text{FD})$ captures the notion that
-the length of the feedback delay slows the \textit{rate} of
-procedural learning.
-
-\subsection{Reinforcement-Learning Model}
-The reinforcement-learning model gradually associates
-responses with stimuli, and therefore does not include an
-explicit representation of the response criterion. The model
-is based on an actor-critic architecture and uses
+\subsection{Reinforcement-learning models}
+The reinforcement-learning models gradually associate
+responses with stimuli, and therefore do not include an
+explicit representation of the response criterion. The
+models are based on an actor-critic architecture and use
 reinforcement learning to form stimulus-response
-associations \parencite{SuttonBarto1998}.
-
-This model assumes that the perceptual representation of
-each stimulus is defined by the pattern of activation across
-25 sensory units that are characterized by overlapping
-tuning curves. Each unit is maximally excited by one
-specific stimulus, which we call the unit's preferred
-stimulus. Specifically, the activation of the $i^\text{th}$
-sensory unit on trial $n$ is given by
+associations \parencite{SuttonBarto1998}. These models
+assumes that the perceptual representation of each stimulus
+is defined by the pattern of activation across 25 sensory
+units that are characterized by overlapping tuning curves.
+Each unit is maximally excited by one specific stimulus,
+which we call the unit's preferred stimulus. Specifically,
+the activation of the $i^\text{th}$ sensory unit on trial
+$n$ is given by
 \begin{equation}
   A_i(n) = \text{exp} \left( \frac{-\left[d_{i,S_n}\right]^2}{\sigma^2} \right)
+  \label{eq:DSL_activation}
 \end{equation}
@@
 $\delta(n-1)$ is the reward prediction error on trial $n-1$.
-Note that as in the delay-sensitive learning model, the
-learning rate is scaled by the inverse of the feedback
-delay. Much evidence suggests that this type of
-stimulus-response learning is mediated largely within the
-striatum, and is facilitated by a dopamine (DA) mediated
-reinforcement learning signal that is time dependent
+Note that as in Equation~\ref{RL}, the learning rate is
+scaled by the inverse of the feedback delay.
+
+Much evidence suggests that this type of stimulus-response
+learning is mediated largely within the striatum, and is
+facilitated by a dopamine (DA) mediated reinforcement
+learning signal that is time dependent
 \parencite[e.g.,][]{ValentinMaddoxAshby2014}. Specifically,
@@
 increasing feedback delays \parencite{YagishitaEtAl2014}.
-The scaling of $\alpha_{actor}$ by $1/t_\text{FD}$ models this
-phenomenon.
+The scaling of $\alpha_{actor}$ by ${1}{t_\text{FD}}$ is one
+way to model this phenomenon. However, it is possible that
+criterial learning via stimulus-response associations is
+mediated by other brain systems that are less sensitive to
+feedback delay. In this case, the learning rate would be
+independent of feedback delay, and the term
+$\frac{\alpha_{actor}}{t_\text{FD}}$ would be replaced by
+$\alpha_{actor}$.
 
@@
 \begin{equation}
-  P(n+1) = P(n) + \frac{\alpha_{critic}}{t_\text{FD}} \delta(n),
+  P(n) = P(n-1) + \frac{\alpha_{critic}}{t_\text{FD}} \delta(n-1),
 \end{equation}
 where $\alpha_{critic}$ is the learning rate of the critic.
-This learning rate is again scaled by the inverse of the
-feedback delay.
+This learning rate may again be scaled by the inverse of the
+feedback delay. If sensitivity to feedback delay is not
+assumed in Equation~\ref{RL}, then the term
+$\frac{\alpha_{critic}}{t_\text{FD}}$ is replaced by
+$\alpha_{critic}$.
+
+Because this class of models does not store an explicit
+response criterion, it cannot incorporate criterial drift.
+It can, however, incorporate perceptual drift. In this case,
+the value of the current stimulus used to compute sensory
+unit activation (Equation~\ref{eq:DSL_activation}) is
+assumed to drift over time according to
+\begin{equation}
+  S_n(t) = S_n(\text{T}_\text{S}) + \eta_S \epsilon(t),
+  \label{eq:stimulus_drift_RL}
+\end{equation}
+where $\eta_c$ scales the magnitude of the drift and
+$\epsilon(t)$ is a noise process.
 
@@
 We investigated how changing the duration of the feedback
-delay and ITI affect criterial learning for each of these
-three models. More specifically, we simulated performance of
+delay and ITI affect criterial learning for each class of
+model. More specifically, we simulated performance of
 each model in a categorization task that included two
@@
 observed performance metric measured in trials-to-criterion
-for each condition. All scores were then normalized by the
-largest observed value, and as a result, all axes in the
-figures that describe the simulation results range from zero
-to one.
+for each condition.
 
@@
 (PSP) to evaluate the performance of each model
-\parencite[PSP;][]{pitt2006global}. PSP calculates the
-proportion of the parameter space where a model makes
+\parencite[PSP;][]{pitt2006global}.  Classical PSP is
+defined for non-stochastic models and is often implemented
+by exploring the parameter space via MCMC.  In contrast, we
+used brute force search over a grid of plausible parameter
+values. Because our models are stochastic, at each sampled
+parameter setting we generated multiple simulations and
+classified the expected qualitative pattern.  PSP calculates
+the proportion of the parameter space where a model makes
 specific qualitative predictions. We focused on four such
@@
 
-\subsubsection{Time-dependent drift model}
-We investigated predictions of this model across a wide
+\subsubsection{Models that assume a stored criterion}
+We investigated predictions of these models across a wide
 range of values for the parameters $\eta_{x}$, $\eta_{c}$,
 and $\alpha$.  Specifically, in the case of $\alpha$,  we
-stepped through every value in the interval $[0, 1]$, with a
-step size of .01. In the case of both $\eta_{x}$, and
-$\eta_{c}$, we searched over the interval $[0.1, 5]$, with a
-step size of 0.5. The results are shown in Figure
-\ref{fig:TDD_results}.
-
-\begin{figure} \centering
-  \includegraphics[width=1\textwidth]{../figures/model_working_memory.pdf}
+stepped through every value in the interval $[0, 0.2]$, with
+a step size of .01.  In the case of both $\eta_{x}$, and
+$\eta_{c}$, we searched over the interval $[0, 1]$, with a
+step size of $0.1$.  We also explored the binary case of
+whether or not the update rate $\alpha$ was scaled by the
+feedback delay. The results are shown in Figure
+\ref{fig:model_class_II_results}.
+
+\begin{figure} 
+  \centering
+  \includegraphics[width=1\textwidth]{../figures/model_new_class_II.pdf}
   \caption{ 
-      \textbf{A:} The proportion of parameter space where
-      the time-dependent drift model predicted each of four
+      \textbf{A:} Assuming a delay-sensitive update, the
+      proportion of parameter space where the models that
+      assume a stored criterion predicted each of four
       qualitative patterns: (1) slower learning under
@@
       % 
-      \textbf{B:} Simulated trials-to-criterion for each
-      condition. All scores were normalized by the largest
-      trials-to-criterion value within a given parameter
-      set, so all axes range from zero to one. 
+      \textbf{B:} Assuming a delay-sensitive update, Boxplot
+      of the parameter ranges leading to each PSP pattern.
+      All parameter values were normalized by the largest
+      value in the search range, so the ordinate ranges from
+      zero to one for all parameters.
       % 
-      \textbf{C:} Boxplot of the parameter ranges leading to
-      each PSP pattern. All parameter values were normalized
-      by the largest value in the search range, so the
-      ordinate ranges from zero to one for all parameters.
-      % 
-      \textbf{D:} Scatter plot of the parameter ranges
-      associated with each PSP pattern. 
-      %
-      \textit{Note:} In all panels, color indicates the PSP
-      pattern.
-}
-  \label{fig:TDD_results}
-\end{figure}
-
-Figure \ref{fig:TDD_results}A shows that the model
-essentially always predicts that increasing the feedback
-delay or the ITI have similar detrimental effects on
-learning. The model makes this prediction because it assumes
-that the criterion drifts during both the feedback delay and
-during the ITI. So the critical variable for the model is
-the time between the response on trial $n$ and the
-presentation of the stimulus on trial $n+1$. The model
-assumes that the criterion drifts during this entire time,
-and therefore how this interval is divided between feedback
-delay and ITI is relatively unimportant to the model
-predictions.
-
-\subsubsection{Delay-sensitive learning model}
-We simulated the performance of the delay-sensitive learning
-across a wide range of parameter values for $\sigma_p$ and
-$\alpha$. Specifically, in the case of $\alpha$ we stepped
-through every value in the interval $[0, 1]$, with a step
-size of .1. In the case of $\sigma_p$, we searched over the
-interval $[0.1, 5]$, with a step size of 0.1.
-
-The results are shown in Figure \ref{fig:DSL_results}. Note
-that virtually all combinations of the parameter values
-predict that feedback delay impairs criterial learning more
-than increasing the ITI. This makes sense because Equation
-\ref{eq:DSL_learning} shows that the delay-sensitive
-learning model predicts that increasing the feedback delay
-(i.e., increasing $t_{FD}$) will impair learning -- for any
-value of $\alpha>0$. Large values of $\sigma_p$ will also
-impair performance because of distortion to the percept, but
-this interference will be the same in all conditions.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=1\textwidth]{../figures/model_procedural_supervised.pdf}
-  \caption{ 
-      \textbf{A:} The proportion of parameter space where
-      the delay-sensitive learning model predicted each of
-      four qualitative patterns: (1) slower learning under
+      \textbf{C:} Assuming no delay-sensitive update, the
+      proportion of parameter space where the models that
+      assume a stored criterion predicted each of four
+      qualitative patterns: (1) slower learning under
       feedback delay, (2) slower learning in the Long-ITI
@@
       % 
-      \textbf{B:} Simulated trials-to-criterion for each
-      condition. All scores were normalized by the largest
-      trials-to-criterion value within a given parameter
-      set, so all axes range from zero to one. 
-      % 
-      \textbf{C:} Boxplot of the parameter ranges leading to
-      each PSP pattern. All parameter values were normalized
-      by the largest value in the search range, so the
-      ordinate ranges from zero to one for all parameters.
-      % 
-      \textbf{D:} Scatter plot of the parameter ranges
-      associated with each PSP pattern. 
+      \textbf{D:} Assuming no delay-sensitive update, Boxplot
+      of the parameter ranges leading to each PSP pattern.
+      All parameter values were normalized by the largest
+      value in the search range, so the ordinate ranges from
+      zero to one for all parameters.
       %
@@
 }
-  \label{fig:DSL_results}
+  \label{fig:model_class_II_results}
 \end{figure}
 
-\subsubsection{Reinforcement-learning model}
+Figure~\ref{fig:model_class_II_results} shows that this
+class of models makes qualitatively different predictions
+depending on whether the update rule is sensitive to
+feedback delay. When delay sensitivity is assumed (panels A
+and B), the model consistently predicts that feedback delays
+impair criterion learning. When delay sensitivity is not
+assumed (panels C and D), the model instead predicts that
+long ITIs impair learning, except in cases with very small
+learning rates $\alpha$ (panel D).
+
+\subsubsection{Reinforcement-learning models}
 We investigated the effects on performance predicted by the
@@
 every value in the interval $[0, 0.2]$, with a step size of
-.01. We constrained our search over these parameters to this
+$.02$. We constrained our search over these parameters to this
 interval because reinforcement-learning models are prone to
@@
 within the allowable 200 trials. In the case of $\sigma$, we
-searched over the interval $[1, 10]$, with a step size of 1.
-As with the other models, all scores were normalized by the
+searched over the interval $[1, 10]$, with a step size of
+$3$.  We also explored the binary case of whether or not the
+update rate $\alpha$ was scaled by the feedback delay. As
+with the other models, all scores were normalized by the
 largest observed value.
 
-The results are described in Figure \ref{fig:RL_results}.
-Note this model predicts that delayed feedback impairs
-criterial learning more than a long ITI across a wide volume
-of parameters settings. As can be seen in Figure
-\ref{fig:RL_results}C, the only exceptions tend to occur
-when the value of any of the three parameters approaches the
-maximum end of the range explored. The logic here is similar
-to the delay-sensitive learning model -- that is, both
-models predict that increasing the feedback delay $t_{FD}$
-will impair learning -- for all values of the learning rate.
+The results are described in
+Figure~\ref{fig:model_class_I_results}.  Note this model
+predicts that delayed feedback impairs criterial learning
+more than a long ITI across the vast majority of parameters
+explored. As can be seen in
+Figure~\ref{fig:model_class_I_results}D, the few exceptions
+tend to occur when perceptual drift has relatively small
+values.
 
@@
   \centering
-  \includegraphics[width=1\textwidth]{../figures/model_procedural_reinforcement.pdf}
+  \includegraphics[width=1\textwidth]{../figures/model_new_class_I.pdf}
   \caption{ 
-      \textbf{A:} The proportion of parameter space where
-      the reinforcement-learning model predicted each of
-      four qualitative patterns: (1) slower learning under
+      \textbf{A:} Assuming a delay-sensitive update, the
+      proportion of parameter space where the models that
+      assume a stored criterion predicted each of four
+      qualitative patterns: (1) slower learning under
       feedback delay, (2) slower learning in the Long-ITI
@@
       % 
-      \textbf{B:} Simulated trials-to-criterion for each
-      condition. All scores were normalized by the largest
-      trials-to-criterion value within a given parameter
-      set, so all axes range from zero to one. 
+      \textbf{B:} Assuming a delay-sensitive update, Boxplot
+      of the parameter ranges leading to each PSP pattern.
+      All parameter values were normalized by the largest
+      value in the search range, so the ordinate ranges from
+      zero to one for all parameters.
       % 
-      \textbf{C:} Boxplot of the parameter ranges leading to
-      each PSP pattern. All parameter values were normalized
-      by the largest value in the search range, so the
-      ordinate ranges from zero to one for all parameters.
+      \textbf{C:} Assuming no delay-sensitive update, the
+      proportion of parameter space where the models that
+      assume a stored criterion predicted each of four
+      qualitative patterns: (1) slower learning under
+      feedback delay, (2) slower learning in the Long-ITI
+      condition, (3) slower learning in the control
+      condition, and (4) any other pattern. 
       % 
-      \textbf{D:} Scatter plot of the parameter ranges
-      associated with each PSP pattern. 
+      \textbf{D:} Assuming no delay-sensitive update, Boxplot
+      of the parameter ranges leading to each PSP pattern.
+      All parameter values were normalized by the largest
+      value in the search range, so the ordinate ranges from
+      zero to one for all parameters.
       %
@@
 }
-  \label{fig:RL_results}
+  \label{fig:model_class_I_results}
 \end{figure}
@@
 \subsubsection{Summary of modeling results}
-We investigated three criterial-learning models that make
-different predictions about how increases in the feedback
-delay and ITI affect learning. The time-dependent drift
-model predicts that the criterion always drifts, so the
-critical variable is the sum of the feedback delay and ITI.
-How this sum is divided into separate feedback delay and ITI
-time intervals has little effect on the model's predictions.
-In contrast, the delay-sensitive learning model predicts
-that feedback delays are necessarily more detrimental to
-performance than increases in the ITI.  Finally, the
-reinforcement-learning model also predicts that, in general,
-feedback delays should impair learning more than long ITIs,
-although this model is somewhat more flexible than the
-delay-sensitive learning model and can account for a small
-or null effect of feedback delay within a restricted region
-of its parameter space. The obvious next question is how
-human criterial learning is affected by these independent
-variables. Experiments 2 and 3 were designed to address this
-question, and therefore also to test the predictions of the
-three models.
-
-\section{Experiment 2}
-Experiment 2 investigated how feedback delay and the length
+We examined two classes of criterial-learning models that
+make different predictions about the effects of feedback
+delay and ITI. The first class assumes that the criterion is
+stored in memory and compared against incoming stimuli to
+generate responses. These models predict greater impairment
+from feedback delay only when the update rule is explicitly
+delay-sensitive. Otherwise, they are generally more impaired
+by long ITIs than by feedback delay. The second class
+assumes no stored criterion, with responses driven instead
+by reinforcement learning through gradual tuning of
+stimulus–response associations. These models almost
+uniformly predict greater impairment from feedback delay
+than from long ITIs.
+
+\section{Experiment 1}
+Experiment 1 investigated how feedback delay and the length
 of the ITI affect criterial learning in humans. As a result,
@@
 
-The stimuli in Experiment 2 were circular sine-wave gratings
+The stimuli in Experiment 1 were circular sine-wave gratings
 that varied across trials in bar width and bar orientation.
@@
 illustrated in Figure \ref{fig_design_exp_1_space}. Each
-criterial-learning problem was created by first randomly
-selecting a relevant dimension, and then randomly selecting
-one of the 7 bins defined on that dimension. Each bin was
-also associated with a corresponding unique value on the
-irrelevant dimension. We buffered the to-be-learned response
-criterion by 10\% of total bin width on either side with a
-no-stimulus region. Random uniform samples from the
-remaining eligible region of each bin were then selected and
-presented to the participant until 9 correct responses out
-of any 10 responses in a row advanced the participant to the
-next problem. Note that every category problem was a simple
-one-dimensional rule in which optimal accuracy was 100\%.
-Note also that the relative location of the optimal response
-criterion varied across problems.
+criterial-learning problem was created by first selecting a
+relevant dimension. This was spatial frequency for the
+problems 1 thorugh 7 and orientation for problems 8 through
+14. We then randomly selected one of the 7 bins defined on
+that dimension. Each bin was also associated with a
+corresponding unique value on the irrelevant dimension. We
+buffered the to-be-learned response criterion by 10\% of
+total bin width on either side with a no-stimulus region.
+Random uniform samples from the remaining eligible region of
+each bin were then selected and presented to the participant
+until 9 correct responses out of any 10 responses in a row
+advanced the participant to the next problem. Note that
+every category problem was a simple one-dimensional rule in
+which optimal accuracy was 100\%.  Note also that the
+relative location of the optimal response criterion varied
+across problems.
 
@@
 \subsubsection{Participants}
-Fifty-nine participants participated in Experiment 2. All
+Fifty-nine participants participated in Experiment 1. All
 were UCSB undergraduates and received course credit for
@@
 \subsection{Results}
-Figure \ref{fig_gmm_hist_1} shows a relative frequency
+Figure \ref{fig_mm_hist_1} shows a relative frequency
 histogram of the trials-to-criterion observed across all
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../figures/fig_exp_1_gmm.png}
+  \includegraphics[width=.8\textwidth]{../figures/fig_exp_1_mm_trunc_exgauss_compare.png}
   \caption{
@@
       trials-to-criterion observed across all three
-      conditions in Experiment 2. The black line are the
-      predictions of the best-fitting one-component Gaussian
-      Mixture Model to these data, whereas the red line
-      represents the best fit of the two-component Gaussian
-      Mixture Model. The two-component model provided a
-      significantly better fit according to AIC, BIC, and a
-      likelihood ratio test.
-}
-  \label{fig_gmm_hist_1}
+      conditions in Experiment 1 and the fit of a single
+      Exponential-Gaussian distribution (Left) and a
+      two-component Exponential-Gaussian mixture model
+      (Right). In the right panel, the blue dashed line is
+      the first higher performance component, the orange
+      dashed line is the second lower performance component.
+      The gray solid line is the sum of the two components.
+      Rug plots along the x-axes show the individual data
+      colored by their assignment to either the higher or
+      lower performance component.  The two-component model
+      provided a significantly better fit according to AIC,
+      BIC, and a likelihood ratio test.
+}
+  \label{fig_mm_hist_1}
 \end{figure}
 
-% AIC for 1 component: 735.7652217198266
-% AIC for 2 components: 662.8095009936025
-% 
-% BIC for 1 component: 739.8861077409194
-% BIC for 2 components: 673.1117160463345
-%
-% LRT statistic: 78.95572072622417
-% Degrees of freedom: 3
-% P-value: 5.140608761896481e-17
-
 We modeled the distribution shown in Figure
-\ref{fig_gmm_hist_1} using a Gaussian mixture model with two
-components. The first component captured participants with a
+\ref{fig_mm_hist_1} using a mixture of Exponential-Gaussian
+distributions with two components, each truncated to the
+range of possible trials-to-criterion values (i.e., $[9,
+512]$). The first component captured participants with a
 relatively low mean number of trials-to-criterion, while the
 second component identified outliers according to the above
-rationale. We compared the fit of this two-component model
-to a single-component model using the Akaike Information
-Criterion (AIC) and Bayesian Information Criterion (BIC),
-both of which indicated a better fit for the two-component
-model (AIC: 662.81 vs. 735.77; BIC: 673.11 vs. 739.89).
-Additionally, a likelihood ratio test confirmed that the
-two-component model provided a significantly better fit than
-the one-component model ($\chi^2(3) = 78.96$, $p < .001$).
-Outlier participants were then excluded from further
+rationale.  We estimated this model via MLE with proper
+normalization on the tasks bounded support and compared them
+via AIC, BIC, and a parametric bootstrap likelihood-ratio
+test (LRT).  We compared the MLE fit of this two-component
+model to a single-component Exponential-Gaussian model using
+the Akaike Information Criterion (AIC) and Bayesian
+Information Criterion (BIC), both of which indicated a
+better fit for the two-component model 
+(
+$\mathrm{AIC}_1=692.15$ vs. $\mathrm{AIC}_2=652.81$; 
+$\mathrm{BIC}_1=698.33$ vs. $\mathrm{BIC}_2=667.23$
+)
+Additionally, a bootstrap likelihood ratio test confirmed
+that the two-component model provided a significantly better
+fit than the one-component model ($D=42.84$,
+$p_{\text{boot}}\approx 0.0099$) Participants that were
+deemed to most likely to belong in the low-performance
+Gaussian by this analysis were excluded from further
 analysis. After making these exlcusions, we were left with
@@
 Figure \ref{fig_exp_1_t2c}A shows the mean
-trials-to-criterion in each condition of Experiment 2.  A
+trials-to-criterion in each condition of Experiment 1.  A
 one-way ANOVA revealed a significant effect of condition
@@
 
-%   Source  ddof1  ddof2         F     p-unc       np2
-% 0    cnd      2     39  6.171331  0.004693  0.240398
-%
-%           A          B         T        dof     p-unc
-% 0     Delay   Long ITI  2.938217  20.959026  0.007864
-% 1     Delay  Short ITI  2.701083  22.997274  0.012748
-% 2  Long ITI  Short ITI  0.219862  17.987044  0.828454
-
 Figure \ref{fig_exp_1_t2c}B shows the mean number of
-problems solved in each condition of Experiment 2. A one-way
+problems solved in each condition of Experiment 1. A one-way
 ANOVA revealed a significant effect of condition ($F(2, 39)
@@
 
-%   Source  ddof1  ddof2         F     p-unc       np2
-% 0    cnd      2     39  5.766358  0.006399  0.228223
-% 
-%           A          B         T        dof     p-unc
-% 0     Delay   Long ITI -2.143777  21.590392  0.043579
-% 1     Delay  Short ITI -3.277500  18.536603  0.004061
-% 2  Long ITI  Short ITI -1.501591  25.965920  0.145267
-
 \begin{figure}
@@
         \textbf{A}: Mean trials to criterion in each
-        condition of Experiment 2. Error bars are standard
+        condition of Experiment 1. Error bars are standard
         errors of the mean.
@@
         \textbf{B}: Mean number of problems solved in each
-        condition of Experiment 2. Error bars are standard
+        condition of Experiment 1. Error bars are standard
         errors of the mean.
@@
 
-\section{Experiment 3}
-Experiment 3 was designed to reinforce the findings of
-Experiment 2 by using the same design but with stimuli that
+\section{Experiment 2}
+Experiment 2 was designed to reinforce the findings of
+Experiment 1 by using the same design but with stimuli that
 have binary-valued dimensions, thereby eliminating criterial
-learning. Unlike Experiment 2, we did not explicitly
+learning. Unlike Experiment 1, we did not explicitly
 instruct participants on the relevant stimulus dimension in
-Experiment 3.  While Experiment 2 isolated criterial
+Experiment 2.  While Experiment 1 isolated criterial
 learning from rule selection and rule switching, this
 experiment isolates rule selection and switching from
-criterial learning. Therefore, Experiment 3 was designed to
+criterial learning. Therefore, Experiment 2 was designed to
 test whether the impaired learning that we observed in
-Experiment 2 when the feedback was delayed can be attributed
+Experiment 1 when the feedback was delayed can be attributed
 principally to criterial learning or whether it could be due
@@
 \subsubsection{Apparatus}
-The apparatus was the same as in Experiment 2.
+The apparatus was the same as in Experiment 1.
 
@@
   \caption{
-      \textbf{A:} An example trial from Experiment 3.
+      \textbf{A:} An example trial from Experiment 2.
       \textbf{B:} Eight example stimuli (out of 64 total
@@
 Participants received detailed instructions about the task.
-However, unlike in Experiment 2, they were not informed of
+However, unlike in Experiment 1, they were not informed of
 the relevant dimension or the correct rule. The trial timing
 for the Delayed-Feedback and Long-ITI conditions in
-Experiment 3 were the same as in Experiment 2. Experiment 3
+Experiment 2 were the same as in Experiment 1. Experiment 2
 did not include a condition that was analogous to the
-Control condition of Experiment 2. Participants practiced
+Control condition of Experiment 1. Participants practiced
 each problem until they responded correctly on 12
@@
 \subsubsection{Participants}
-Thirty-four participants participated in Experiment 3. All
+Thirty-four participants participated in Experiment 2. All
 were Macquarie University undergraduates and received course
@@
 \subsection{Results}
-Figure \ref{fig_gmm_hist_2} shows a relative frequency
+Figure \ref{fig_mm_hist_2} shows a relative frequency
 histogram of the trials-to-criterion observed in both
@@
 participants were able to learn each problem on average in
-less than 100 trials.  Unlike Experiment 2, there were no
+less than 100 trials.  Unlike Experiment 1, there were no
 highly suspicious outliers in this data set. Nevertheless,
-for symmetry with Experiment 2, we performed the same
+for symmetry with Experiment 1, we performed the same
 Gaussian mixture model analysis as used there. The
 two-component model provided a slightly smaller AIC value
-(317.40 vs. 317.78), but a slightly larger BIC value (325.03
-vs. 320.83). The likelihood-ratio test was not significant
-($\chi^2(3) = 6.38$, $p = .09$). We therefore did not
-exclude any participants from further analysis in Experiment
-2.
+(324.75 vs. 317.18), and a slightly larger BIC value (329.33
+vs. 327.87). The bootstrap likelihood-ratio test was not
+significant ($D = 11.41, p_{\text{boot}} \approx 0.2277$).
+We therefore did not exclude any participants from further
+analysis in Experiment 1.
 
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../figures/fig_exp_2_gmm.png}
+  \includegraphics[width=.8\textwidth]{../figures/fig_exp_2_mm_trunc_exgauss_compare.png}
   \caption{
@@
       trials-to-criterion observed across all three
-      conditions in Experiment 3. The black line represents
+      conditions in Experiment 2. The black line represents
       the best fit of the one-component Gaussian Mixture
@@
 }
-  \label{fig_gmm_hist_2}
+  \label{fig_mm_hist_2}
 \end{figure}
 
-% AIC for 1 component: 317.77535652645616
-% AIC for 2 components: 317.3995958509652
-% 
-% BIC for 1 component: 320.8280775756885
-% BIC for 2 components: 325.03139847404606
-% 
-% LRT statistic: 6.3757606754909375
-% Degrees of freedom: 3
-% P-value: 0.09469309595624516
-
 Figure \ref{fig_exp_2_t2c}A shows the mean
-trials-to-criterion in each condition of Experiment 3. An
+trials-to-criterion in each condition of Experiment 2. An
 independent-samples t-test revealed no significant effect of
@@
 
-%   Source  ddof1  ddof2        F     p-unc       np2
-% 0    cnd      1     32  0.47046  0.497715  0.014489
-% 
-%        A          B         T   dof     p-unc
-% 0  delay  immediate  0.685901  32.0  0.497715
-
-%   Source  ddof1  ddof2         F    p-unc       np2
-% 0    cnd      1     32  0.235294  0.63093  0.007299
-% 
-%        A          B         T   dof    p-unc
-% 0  delay  immediate -0.485071  32.0  0.63093
-
 \begin{figure}
@@
         \textbf{A}: Mean trials to criterion in each
-        condition of Experiment 3. Error bars are standard
+        condition of Experiment 2. Error bars are standard
         errors of the mean.
@@
         \textbf{B}: Mean number of problems solved in each
-        condition of Experiment 3. Error bars are standard
+        condition of Experiment 2. Error bars are standard
         errors of the mean.
@@
 
-\subsection{Discussion of Experiments 2 and 3}
-Experiment 2 clearly showed that a short feedback delay of
+\subsection{Discussion of Experiments 1 and 2}
+Experiment 1 clearly showed that a short feedback delay of
 3.5 s slowed criterial learning. In contrast, increasing the
@@
 delay had its interfering effects on criterial learning,
-Experiment 3 was designed to confirm this inference. The
+Experiment 2 was designed to confirm this inference. The
 goal here was to examine the effects of the same feedback
@@
 and switching). If the feedback delay effects observed in
-Experiment 2 were acting on some general category-learning
+Experiment 1 were acting on some general category-learning
 skill then we should have seen the same interfering effects
-of feedback delay in Experiment 3. However, if the
-feedback-delay interference of Experiment 2 was operating
+of feedback delay in Experiment 2. However, if the
+feedback-delay interference of Experiment 1 was operating
 selectively on criterial learning, then it should disappear
-in Experiment 3, since no criterial learning was required.
+in Experiment 2, since no criterial learning was required.
 Our results strongly supported this latter prediction.
@@
 \section{General Discussion}
-We developed three novel models of criterial learning that
-were designed to explore how feedback delay and ITI duration
-affect criterial learning. Two of these models assumed that
+We developed novel models of criterial learning that were
+designed to explore how feedback delay and ITI duration
+affect criterial learning.  One class of model assumed that
 decisions are made by comparing the current percept to a
@@
 of this criterion is updated following error feedback via a
-gradient-descent learning rule. The first of these models,
-the time-dependent drift model, posits that both the
-criterion and the percept drift over time, leading to
-similar impairments when feedback is delayed or the ITI is
-increased. The second model, the delay-sensitive learning
-model, assumes that although neither the percept nor the
-criterion drift over time, optimal criterial learning
-requires immediate feedback. As a result, this model
-predicts that increasing the feedback delay will slow the
-rate of criterial learning. The third model, the
-reinforcement-learning model, assumes that criterial
-learning arises through the gradual formation of
-stimulus-response associations, rather than via the
-construction of a response criterion. This model also
-assumes that the rate at which these associations are
-learned is slowed by feedback delays.
-
-Simulations of these models in a task that was structurally
-identical to Experiment 2 revealed that the time-dependent
-drift model typically predicts that the most important
-variable for criterial learning is the total amount of time
-between the response and the stimulus presentation that
-defines the next trial. Where the feedback is presented
-within this interval is relatively unimportant. In contrast,
-both the delay-sensitive learning model and the
-reinforcement-learning model consistently predict that
-feedback delays should impair performance more than a long
-ITI.
+gradient-descent learning rule.  This class of model
+generates sensitivity to feedback delay and ITI duration
+through (1) directly assuming that the update rule is
+inversely scaled by the feedback delay, (2) assuming that
+the representation of the percept drifts over time, or (3)
+assuming the representation of the criterion drifts over
+time.  Simulations of these models in a task that was
+structurally identical to Experiment 1 revealed that this
+class of model typically predicts that feedback delay in
+more impairing than a long ITI only when the update rule is
+explicitly delay-sensitive.  Otherwise, it typically
+predicts that long ITIs are more impairing than feedback
+delays.
+
+The second model class, reinforcement-learning models,
+assumes that criterial learning arises through the gradual
+formation of stimulus-response associations, rather than via
+the construction of a response criterion.  These models can
+generate sensitivity to feedback delay and ITI duration
+through the same two mechanisms as the first class of
+models. However, since these models do not explicitly
+represent a criterion, they do not assume that the
+representation of the criterion drifts over time.   This
+class of model  typically predicts that feedback delay is
+more impairing than a long ITI across wide ranges of
+parameter values.
 
@@
 sensitive to feedback delay but not to the ITI duration.
-This suggests that the delay-sensitive learning model and
-the reinforcement-learning model provide a better account of
-human behavior across wide ranges of their respective
-parameter spaces than the time-dependent drift model.
-
-The time-dependent drift model was based on the idea that
-criterial learning might be entirely supported by working
-memory. The assumption of drift in this model stems from the
-understanding that maintaining items in working memory is
-inherently challenging.  The longer an item is held, the
-more likely it is to deteriorate. Therefore, the model
-predicts that performance should deteriorate as the time
-between the response on trial $n$ and stimulus presentation
-on trial $n+1$ increases, regardless of how this interval is
-divided between feedback delay and ITI. The results of
-Experiments 1 and 2 strongly disconfirm this prediction.
-
-In contrast, the delay-sensitive learning model was inspired
-by the idea that a response criterion might be encoded in a
-more stable memory system. One appealing candidate for this
-function is the cerebellum, where synaptic plasticity has
-been shown to follow a gradient descent learning rule and
-also to be sensitive to feedback timing
-\parencite{brudner2016delayed, held_adaptation_1966,
+These results shed important constraints on the new models
+of criterial learning developed here. In particular, the
+reinforcement-learning models are consistent with the
+observed behavioural data across wide ranges of their
+parameter spaces, regardless of whether or not the update
+rule is delay-sensitive.  However, the models that assume a
+stored criterion are also consistent with the observed
+behavioral data, but only when the update rule is
+delay-sensitive.
+
+One natural neural underpinning for models that assume a
+stored criterion is that the criterion is maintained in
+working memory. However, since it is widely accepted that
+working memory is largely insensitive to both feedback delay
+and ITI duration, this interpretation is difficult to
+reconcile with our results suggesting a delay-sensitive
+update rule is required for this class of model. An
+alternative possibility for where the criterion might be
+encoded in a more stable memory system.  One appealing
+candidate for this function is the cerebellum, where
+synaptic plasticity has been shown to follow a gradient
+descent learning rule and also to be sensitive to feedback
+timing \parencite{brudner2016delayed, held_adaptation_1966,
 honda_adaptation_2012, kitazawa_effects_1995,
@@
 
-Our results do not strongly favor the delay-sensitive
-learning model over the reinforcement learning model or vice
-versa. The critical disagreement between these accounts is
-whether the criterion has any real psychological meaning. In
-the delay-sensitive learning model it does, and thus this
-model is psychologically similar to the time-dependent drift
-model -- the main disagreement being about whether the
-representation of the criterion is stable and if the
-updating of the criterion is sensitive to feedback delay.
-
-The reinforcement-learning model makes very different
-psychological assumptions. This model assumes that behaviors
-described as ``criterial learning'' are actually mediated by
-the learning of stimulus-response associations. According to
-this account, the criterion has no internal representation
-and therefore no psychological meaning. Testing between
-these two very different accounts of criterial learning
-should be a focus of future research.
+Our results do not strongly favor one model over another.
+The critical disagreement between these accounts is whether
+the criterion has any real psychological meaning. One class
+assumes the criterion is stored in memory and compared
+against incoming stimuli to generate responses.  The
+reinforcement-learning model, in contrast, makes very
+different psychological assumptions. This model assumes that
+behaviors described as ``criterial learning'' are actually
+mediated by the learning of stimulus-response associations.
+According to this account, the criterion has no internal
+representation and therefore no psychological meaning.
+Testing between these two very different accounts of
+criterial learning should be a focus of future research.
 
@@
 question of whether a response criterion is a fundamental
-psychological construct, as the delay-sensitive learning
-model suggests, or whether it is unnecessary, as assumed by
-the reinforcement learning model. Even so, there are some
-results in the category-learning literature that support the
-interpretation of the reinforcement-learning model. In
-information-integration (II) category-learning tasks, the
-optimal strategy is similarity-based, and difficult or
-impossible to describe verbally
-\parencite[e.g.,][]{AshbyValentin2018}. When the stimuli
-vary on two dimensions, the stimuli from contrasting
+psychological construct, or whether it is unnecessary. Even
+so, there are some results in the category-learning
+literature that support the interpretation of the
+reinforcement-learning model. In information-integration
+(II) category-learning tasks, the optimal strategy is
+similarity-based, and difficult or impossible to describe
+verbally \parencite[e.g.,][]{AshbyValentin2018}. When the
+stimuli vary on two dimensions, the stimuli from contrasting
 categories can be partitioned by a decision bound that is
@@
 contrasting response. Furthermore, in agreement with
-Experiment 2, II category learning is impaired by short
+Experiment 1, II category learning is impaired by short
 feedback delays \parencite{MaddoxAshbyBohil2003,

### 2025-09-11 code/imports.py

@@ -1,10 +1,21 @@
 import os
 import glob
+import string
+
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 import seaborn as sns
+
 import pingouin as pg
 from sklearn.mixture import GaussianMixture
-from scipy.stats import chi2
-import string
+
+from scipy.stats import chi2, exponnorm, norm, skew
+from scipy.optimize import minimize
+from scipy.special import erfcx, logsumexp, softmax, expit
+
+import pymc as pm
+import arviz as az
+import pytensor.tensor as pt
+
+

### 2025-09-11 code/inspect_model_new.py

@@ -3,214 +3,553 @@ import pandas as pd
 import matplotlib.pyplot as plt
 import seaborn as sns
 
-d1 = pd.read_csv('../output/param_record_class_I_multi.csv')
-d2 = pd.read_csv('../output/param_record_class_II_multi.csv')
+def inspect_class_I():
 
-d1 = d1.groupby([
-    'sigma_perceptual_noise', 'alpha_actor', 'alpha_critic',
-    'eta_perceptual_drift', 'delay_sensitive_update'
-])[['t2c_delay', 't2c_liti', 't2c_siti']].mean().reset_index()
+    print()
+    print("Inspecting Class I models...")
 
-d2 = d2.groupby([
-    'alpha', 'eta_perceptual_drift', 'eta_criterion_drift',
-    'delay_sensitive_update'
-])[['t2c_delay', 't2c_liti', 't2c_siti']].mean().reset_index()
+    # NOTE: Class I setup
+    d1 = pd.read_csv('../output/param_record_class_I_multi.csv')
 
-d1['psp_class'] = 'Other'
+    d1 = d1.groupby([
+        'sigma_perceptual_noise', 'alpha_actor', 'alpha_critic',
+        'eta_perceptual_drift', 'delay_sensitive_update'
+    ])[['t2c_delay', 't2c_liti', 't2c_siti']].mean().reset_index()
 
-# delay is greater than long iti and short iti
-d1.loc[(d1['t2c_delay'] > d1['t2c_liti']) &
-       (d1['t2c_delay'] > d1['t2c_siti']),
-       'psp_class'] = 'Delay Impaired'
 
-# long iti is greater than delay and short iti
-d1.loc[(d1['t2c_liti'] > d1['t2c_delay']) &
-       (d1['t2c_liti'] > d1['t2c_siti']),
-       'psp_class'] = 'Long ITI Impaired'
+    d1['psp_class'] = 'Other'
 
-# short iti is greater than delay and long iti
-d1.loc[(d1['t2c_siti'] > d1['t2c_delay']) &
-       (d1['t2c_siti'] > d1['t2c_liti']),
-       'psp_class'] = 'Short ITI Impaired'
+    # delay is greater than long iti and short iti
+    d1.loc[(d1['t2c_delay'] > d1['t2c_liti']) & (d1['t2c_delay'] > d1['t2c_siti']),
+           'psp_class'] = 'Delay Impaired'
 
-# make psp_class an ordered category
[…]

### 2025-09-11 code/inspect_results_behave.py

@@ -1,20 +1,6 @@
 from util_func import *
 from imports import *
 
-d = load_data()
-
-inspect_duration_vs_trials(d)
-
-dd = d.groupby(['exp', 'cnd', 'sub'])[['t2c', 'nps']].mean().reset_index()
-print(dd.groupby(['exp', 'cnd'])['sub'].nunique())
-
-dd1 = dd.loc[dd['exp'] == 'Exp 1'].copy()
-dd2 = dd.loc[dd['exp'] == 'Exp 2'].copy()
-
-report_exp_1(dd1)
-report_exp_2(dd2)
-
-d3 = load_data_exp_3()
-dd3 = d3.groupby(['exp', 'cnd', 'sub'])[['t2c', 'nps']].mean().reset_index()
-print(dd3.groupby(['exp', 'cnd'])['sub'].nunique())
-report_exp_3(dd3)
+# report_exp_1()
+report_exp_2()
+# report_exp_3()

### 2025-09-11 code/model_new.py

@@ -39,7 +39,7 @@ def simulate_class_I(params, args):
     t2c = 0
 
     # static params
-    dim = 25
+    dim = 100
     w_base = 0.5
     w_noise = 0.01
     vis_amp = 1.0
@@ -64,6 +64,8 @@ def simulate_class_I(params, args):
         else:
             cat = 2
 
+        vis_act = np.zeros(dim)
+
         # compute input activation via radial basis functions
         vis_dist_x = 0.0
         for i in range(0, dim):
@@ -84,12 +86,15 @@ def simulate_class_I(params, args):
         act_A = act_A if resp == 1.0 else 0.0
         act_B = act_B if resp == 2.0 else 0.0
 
+        vis_act_pre = vis_act.copy()
+        vis_act = np.zeros(dim)
+
         # perceptual drift may occur during the feedback delay
         # TODO: check with Greg for thoughts about this implementation
         if eta_perceptual_drift:
             vis_dist_x = 0.0
+            xx = x + eta_perceptual_drift * t_delay * np.random.normal(0, 1)
             for i in range(0, dim):
-                xx = eta_perceptual_drift * t_delay * np.random.normal(0, 1)
                 vis_dist_x = xx - i
                 vis_act[i] = vis_amp * math.exp(-(vis_dist_x**2) / sigma_perceptual_noise)
 

### 2025-09-11 code/util_exgauss.py

@@ -1,0 +1,478 @@
+from imports import *
+
+try:
+    from sklearn.base import BaseEstimator
+except Exception:
+
+    class BaseEstimator:  # lightweight fallback
+        pass
+
+
+class _ExG:
+
+    @staticmethod
+    def logpdf(x, mu, sigma, tau):
+        """Stable log-pdf of ExGaussian (untruncated)."""
+        x = np.asarray(x)
+        z = (mu - x) / (np.sqrt(2.0) * sigma) + (sigma / (np.sqrt(2.0) * tau))
+        return (-np.log(2.0 * tau) + (sigma**2) / (2.0 * tau**2) +
+                (mu - x) / tau - z**2 + np.log(erfcx(z)))
+
+    @staticmethod
+    def cdf(x, mu, sigma, tau):
+        K = tau / sigma
+        return exponnorm.cdf(x, K, loc=mu, scale=sigma)
+
+    @staticmethod
+    def ppf(u, mu, sigma, tau):
+        K = tau / sigma
+        return exponnorm.ppf(u, K, loc=mu, scale=sigma)
+
+
+class ExGaussianMixture(BaseEstimator):
+    """
+    Truncated ExGaussian mixture with a scikit-learn-like API.
+
+    Parameters
+    ----------
+    n_components : int
+        Number of mixture components (>=1).
+    a, b : float
+        Truncation bounds; model is normalized on [a, b].
+    n_init : int
+        Number of random initializations; best (highest log-lik) is kept.
+    random_state : int or np.random.Generator or None
+        RNG seed or Generator.
+    maxiter : int
+        Max optimizer iterations per start.
+    tol : float
+        Optimizer tolerance.
+    init_method : {'quantiles','random'}
+        How to seed component means.
+
+    Attributes (after fit)
+    ----------------------
+    weights_ : (K,) mixture weights
+    mu_      : (K,) component Normal means
+    sigma_   : (K,) component Normal std
+    tau_     : (K,) component Exp means
+    loglik_  : float, total log-likelihood on training X
+    n_params_ : int, number of free parameters (3K + (K-1))
+    converged_ : bool
+    """
+
+    def __init__(
+        self,
+        n_components=1,
+        a=12.0,
+        b=600.0,
+        n_init=10,
+        random_state=None,
+        maxiter=1000,
+        tol=1e-6,
+        init_method="quantiles",
+    ):
[…]

### 2025-09-11 code/util_func.py

@@ -1,7 +1,9 @@
 from imports import *
+from util_exgauss import *
 
 
 def load_data_exp_3():
+
     dir_delay = '../data/delay'
     dir_immed = '../data/immed'
     dir_maip = '../data/maip'
@@ -54,7 +56,9 @@ def load_data_exp_3():
     d_maip = pd.concat(d_maip, ignore_index=True)
     d_dual = pd.concat(d_dual, ignore_index=True)
 
-    d_dual = d_dual[['t', 't_prob', 'bnd', 'cat', 'x', 'y', 'rsp', 'rt', 'sub']]
+    d_dual = d_dual[[
+        't', 't_prob', 'bnd', 'cat', 'x', 'y', 'rsp', 'rt', 'sub'
+    ]]
 
     d_delay['cnd'] = 'Delay'
     d_immed['cnd'] = 'Long ITI'
@@ -71,14 +75,18 @@ def load_data_exp_3():
     d['experiment_duration'] = d['experiment_duration'] * 5 / 60
     d['trials_completed'] = d.groupby(['cnd', 'sub'])['t'].transform('max')
 
-    d['prob_num'] = d.groupby(['cnd', 'sub'])['t_prob'].transform(lambda x: (x < x.shift(1)).cumsum() + 1)
-    d['t2c'] = d.groupby(['cnd', 'sub', 'prob_num'])['t_prob'].transform('count')
+    d['prob_num'] = d.groupby(
+        ['cnd',
+         'sub'])['t_prob'].transform(lambda x: (x < x.shift(1)).cumsum() + 1)
+    d['t2c'] = d.groupby(['cnd', 'sub',
+                          'prob_num'])['t_prob'].transform('count')
     d['nps'] = d.groupby(['cnd', 'sub'])['prob_num'].transform('max')
 
     return d
 
 
 def load_data():
+
     d1 = load_data_exp_1()
     d2 = load_data_exp_2()
 
@@ -162,6 +170,16 @@ def load_data_exp_2():
 
     d = pd.concat([pd.read_csv(f) for f in f_names])
 
+    f_names = [
+        os.path.join('../data/bin_data_v1', f)
+        for f in os.listdir('../data/bin_data_v1') if f.endswith('.csv')
+    ]
+
+    d1 = pd.concat([pd.read_csv(f) for f in f_names])
+    d1['sub_num'] = d1['sub_num'] + 100
+
+    d = pd.concat([d, d1], ignore_index=True)
+
[…]

### 2025-10-17 write/response_to_reviews_GA.tex
updated analysis et al for revision

\documentclass[12pt]{article}
% Language and font
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
% Bibliography with APA style
\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
\DeclareLanguageMapping{american}{american-apa}
\addbibresource{sample.bib}
% Layout and formatting
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\setstretch{1.0}
\usepackage{parskip} % disables indentation and adds vertical space between paragraphs
\setlength{\parindent}{0pt}
% Math and symbols
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{gensymb}
% Graphics and floats
\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
% Optional: nicer font
\usepackage{lmodern}
% Turn off section numbers
\setcounter{secnumdepth}{0}

\title{\begin{large}\textbf{Response to Reviews} \\ \vspace{.1in}
Is there a Criterion in Criterial Learning? \\ Insights from Studying Feedback Delays \\ \vspace{.1in}
\textit{Journal of Experimental Psychology: Learning, Memory, and Cognition} \\ \vspace{-.1in}
Manuscript number: XLM-2024-0436 \end{large}}
\author{
    Matthew J. Crossley, 
    Benjamin O. Pelzer,
    F. Gregory Ashby
}
\date{}

\begin{document}

\maketitle

We would like to thank the editor and reviewers for their
excellent suggestions and thoughtful comments. We worked
very hard to respond to all of these. As a result, we are
confident that the revised manuscript is significantly
stronger than our original submission. Please note that we
changed the title of the manuscript to more directly convey
the main point of our paper.

Our specific response to each reviewer comment is as
follows.

\section{Reviewer 1}
We were pleased that Reviewer 1 ``really enjoyed reading the
manuscript'' and ``highly recommend(ed) this paper to be
published after addressing a number of mostly minor
concerns.''

\subsection{Reviewer comment 1}
1. Exploration of model space: When reading the description
of the models it was not entirely clear to me whether the
three presented models explore the space of possible models
to a sufficient degree or whether there might be some models
missing from consideration that could change the overall
message.

\subsubsection{Author response}
This is an excellent and important point. In response, we
greatly expanded the set of models we explored, and we also
now classify these models more systematically. The previous
version examined 3 models, whereas the revision examines 12.
These 12 models are partitioned into two classes: one that
assumes a criterion is learned and one that assumes SR
learning with no criterion. The criterial-learning class
includes 8 different models that were created by factorially
combining three different binary assumptions: perceptual
drift (yes or no), criterial drift (yes or no), and
delay-sensitive updating (yes or no). The SR learning class
includes four models created by factorially combining two
binary assumptions: perceptual drift (yes or no) and
delay-sensitive feedback (yes or no).

\subsection{Reviewer comment 2}
2. Simulation results figures: There were some aspects of
the results figures for the simulation that I found
difficult to understand:

(a) Even after reading the full paper and thinking about it
for a while, I do not understand what panels B are supposed
to show. Does this only show the subset of the parameter
space in which one of the two relevant patterns was
observed? If not, shouldn't there be data points everywhere?

(b) I do not understand why in Figure 1, for example, panel
C shows some cases in which the predicted qualitative
pattern is either ``control impaired'' or ``other'', but there
is no corresponding bar in panel A.

(c) The bars in Figure 2A are smaller than in Figures 1A and
3A. Why?

\subsubsection{Author response}
With the new approach to model exploration, we have heavily
revised the simulation results figures.  We hope that the
new figures are clearer and address the reviewer's concerns.

\subsection{Reviewer comment 3}
3. Calling the simulation Experiment 1: I found it extremely
confusing that the simulation using parameter-space
partitioning was called ``Experiment 1''. If I read
``Experiment'' I expect to see some empirical data obtained
from participants and not solely a simulation. I suggest
renaming this section to something clearer. For example
``Simulation Study'' or ``Parameter-Space Partitioning
Across
Models''.

\subsubsection{Author response}
After thinking about this a bit, we agree that referring to
the computational modeling analyses as Experiment 1 was a
mistake. First, we agree with the reviewer that this
nomenclature is confusing, but we also came to realize that
the equations in this section might dissuade empirically
oriented readers from reading about our empirical results.
As a result, the revised manuscript renames the section
describing the computational modeling ``New Computational
Models of Criterial Learning'' and this more technical
section has been moved to later in the manuscript -- after
the two empirical experiments are described.

\subsection{Reviewer comment 4}
4. Outliers in Exp. 2:

- While I think the argument for removing outliers in Exp. 2
is fine, the actual criterion defining an outlier is not
actually given (which feels a bit ironic in a paper about
criteria).

\subsubsection{Author response}
Good point. We now say ``for each participant, we used the
best-fitting two component model to compute the likelihood
that their performance was best described by each component.
Participants whose data were estimated to most likely belong
to the low-performance distribution were classified as
outliers and excluded from further analysis.''

\subsection{Reviewer comment 5}
4. Outliers in Exp. 2:

- How do the results look with these outliers included? Does
the difference between condition vanish? Adding a footnote
or so should be enough.

- The fact that some participants did not even learn a
single criterion feels very surprising to me. Is this common
in such experiments? Please add a sentence embedding this
finding within your general experience running such
experiments.

\subsubsection{Author response}
When the outliers are included in the analysis, the
difference between conditions does indeed vanish. We have
added the following footnote: ``Including the excluded
outliers in these analyses eliminates the difference between
conditions.''

In our experience, it is not unusual for a small subset of
participants to perform very poorly in experiments of this
kind. This typically reflects a lack of engagement or
misunderstanding of the task instructions rather than an
inability to perform the task itself. We now state exactly
that in the revision.

\subsection{Reviewer comment 6}
Minor points:

- Does Eq. 13 show results for trial n or n + 1? The text
and equation are inconsistent.

\subsubsection{Author response}
We corrected this typo.

\subsection{Reviewer comment 7}
- The description regarding the randomisation of Experiment
2 on page 12 (paragraph before ``Procedure'') seems to be
inconsistent with Figure 4. Is the dimension picked at
random for each trial or always the same for certain problem
numbers (as implied by Figure 4)?

\subsubsection{Author response}
We agree that the description of the design of this
experiment (now Experiment 1) needed improvement. We revised
the paragraph Reviewer 1 refers to in this comment as
follows.

``Each participant practiced each of 14 one-dimensional
category-learning tasks, or problems, until they responded
correctly on 9 of the 10 previous trials, at which point the
problem changed. The 14 different category structures are
described in Figure 1. The relevant dimension was bar
thickness in problems 1 – 7 and bar angle in problems 8 –
14. Each problem included stimuli in two distinct clusters
that varied over a restricted range of the relevant
dimension. Critically, the optimal criterion value varied
from problem-to-problem with respect to its position within
the stimulus range. For example, for some problems the
optimal criterion was below the midpoint of the range and
for other problems it was above the midpoint.''

\subsection{Reviewer comment 8}
- In figure 4, the top says Problems ``1 thru 13'', but only
shows 7 problems.

\subsubsection{Author response}
We corrected this typo.

\vspace{.2in}

\section{Reviewer 2}

We were pleased that Reviewer 2 believed our  ``manuscript
reports a sophisticated approach to an important problem''
and that the reviewer ``especially like(d) the use of
mathematical models.''

\subsection{Reviewer comment 1}
I was concerned that the models seem to have multiple
varying features. It might be more useful to isolate
specific processes while holding other model elements
constant.

\subsubsection{Author response}
This is the same as Reviewer 1's first comment. As described
there, we strongly agree with this concern, and as a result,
we greatly expanded the number of models we examined, and we
reclassified these according to the fundamental assumption
of whether or not they assumed a criterion value is
specifically learned and stored trial-by-trial in memory.

\subsection{Reviewer comment 2}
I think it makes more sense to have a prediction simulation
section and Experiments 1 and 2 rather than to consider the
simulation the first experiment.

\subsection{Author response}
This is the same as Reviewer 1's third comment. As noted
there, we agree and as a result, we now refer to the two
empirical experiments as Experiment 1 and 2, and the
computational modeling section has been renamed to ``New
Computational Models of Criterial Learning'' and moved to
after the two experiments.

\subsection{Reviewer comment 3}
The data in Figure 6 can likely be fit much more closely by
an ExGaussian distribution than either the 1 or 2 component
Gaussian mixtures that are shown.

\subsubsection{Author response}
We thank the reviewer for this suggestion. We now estimate
both a single-component truncated ExGaussian and a
two-component truncated ExGaussian mixture via MLE with
proper normalization on the tasks bounded support $[9,
512]$, and compared them via AIC, BIC, and a parametric
bootstrap likelihood-ratio test (LRT).  The results from
moving from the mixture of Gaussians used in our orginal
submission to the truncated ExGaussian mixture used in the
revision does not change the qualitative pattern of our
results in any way. However, we agree that the ExGaussian
and ExGaussian mixture better capture the skewed and bounded
distribution of our trials-to-criterion data so we have
adopted it in the revision.

\subsection{Reviewer comment 4}
Replication of the criterial learning results from the
currently-labeled Experiment 2 seems like a good idea given
the fairly low sample sizes per condition.

\subsubsection{Author response}
We agree with the reviewer that a replication would be
valuable, but we believe the data and the expanded modeling
analysis currently presented represent a significant
contribution to the field.

\subsection{Reviewer comment 5}
Did stimuli remain on the screen across the feedback delay?
I don't think so. If not, what happens if the stimulus
reappears at the time of feedback? I could imagine this
would eliminate or dramatically reduce the negative effect
of feedback delay. If so, is that consistent with the
conclusions? That is, would the models that predict a
deleterious effect of feedback delay also predict that it
should be eliminated with ``reminder'' stimuli at the time
of feedback?

\subsubsection{Author response}
As shown in Figures 2 and 5, the stimulus was presented up
until a response was made. At this time the stimulus was
removed and replaced by a noise mask until feedback was
delivered. The question of how our results might have
changed if the stimulus were to reappear at the time of
feedback is both interesting and insightful. In the very
first study that investigated the effects of feedback delays
on category learning (Maddox, Ashby, \& Bohil, 2003,
JEP:LM$\&$C) we found that a short 2.5 s feedback delay
impaired information-integration category learning, but not
rule-based learning -- but only if a noise mask was
presented during the time between response and feedback. The
published 2003 article doesn't explore this issue, but we
hypothesized at the time that in the absence of the mask,
participants might be able to use eidetic imagery to keep
the relevant synapses active during the delay period. As a
result, we agree with the reviewer's prediction that
re-presenting the stimulus at the time of feedback would
eliminate the impaired learning that we observed. This
interesting point seems tangential to our main point,
however, and as a result, we did not address this point in
the revision.

\subsection{Reviewer comment 6}
For the currently-labeled Experiment 3, reporting
fail-to-reject results for t-tests is not a valid method for
establishing that feedback delay has no effect. This is
especially a concern given the low participant count per
condition and the fact that empirical effects go in the
``right'' direction (lower performance for delayed
feedback).

\subsubsection{Author response}
This is a completely valid point. As in almost all
psychological experiments, our Experiment 2 lacks the
statistical power to make strong inferences about the null
results we obtained there. We now explicitly acknowledge
this point in paragraph 2 of the section entitled
``Discussion of Experiments 1 and 2.'' Also, rather than try
to make a statistical argument in support of our null
results, we now use a converging operations approach by
appealing to the literature. The new relevant paragraph is
as follows.

``Whereas the results of Experiment 1 seem clear, the
Experiment 2 results are weaker since they suggest a null
result. Although insufficient power makes it impossible to
rule out the possibility that Experiment 2 missed some small
effect of feedback delay, the null results found there are
consistent with a wide variety of other evidence. In
particular, a variety of other studies have reported that
feedback delays of up to 10 s do not significantly slow
one-dimensional rule-based learning (Dunn et al., 2012; Ell
et al., 2009; Maddox et al., 2003; Maddox \& Ing, 2005;
Worthy et al., 2013). In addition, many studies have
reported evidence that one-dimensional rule-based learning
recruits working memory and executive attention, but not
procedural learning (for reviews, see, e.g., Ashby \&
Valentin, 2017; Ashby et al., 2020). So the results of
Experiment 2 are consistent with all of these studies. On
the other hand, several of these feedback-delay studies did
not use binary-valued stimulus dimensions, so some criterial
learning was required. If so, then why did they all report
no effect of feedback delay? It is important to note that
criterial learning was not the focus of any of these
studies, and as a result, in all of these previous studies
the optimal criterion was always set exactly midway between
the category prototypes, which makes criterial learning
trivial. For example, under these conditions, criterial
learning might not even require feedback. The unsupervised
category-learning experiments reported by Ashby et al.
(1999) provide strong support for this because all of their
rule-based participants learned the correct criterion (which
was midway between the category means), even though the task
was completely unsupervised.''

\subsection{Reviewer comment 7}
I recommend running a bigger replication study that combines
both tasks to demonstrate the claimed interaction (feedback
delay impairs performance only for the criterion learning
task). The replication study should be preregistered with a
sample size justification based on the interaction effect
defined by the across-experiment comparison in the current
data. Adding a ``stimulus reminder at feedback'' condition
to this new study might be helpful in clarifying the
mechanisms producing the delay effect.

\subsubsection{Author response}
We agree that a replication / extension study would
strengthen our manuscript -- as it would for almost all
empirical papers. We chose not to pursue this suggestion,
however.  Even so, we believe we did significantly
strengthen our manuscript by greatly expanding the class of
computational models that we investigated, and by better
situating our results within the existing literature.

\section*{Action Editor}

\subsection{AE comment 1}
The final paragraph of the introduction summarizes the
experiments. This seems more appropriate in the abstract,
given that no information about the experiments are
presented in the abstract.

\subsubsection{Author response}
Thanks for this suggestion. The revised abstract now has
more detail about the experiments.

\subsection{AE comment 2}
Equation 1 needs some explanation, as this implies that the
decision Rn is made immediately on stimulus presentation. In
other words, Rn is not an outcome of some time-extended
cognitive process. Am I misunderstanding equation 1?

\subsubsection{Author response}
This is a valid and insightful point. Equation 1 does assume
that the response is made immediately upon stimulus
presentation. Because we are not trying to account for
response time, we decided to not explicitly model the
time-extended cognitive processes that govern evidence
accumulation. We now state this explicitly in the revision.
Specifically, immediately after equation 1, we added the
following text:

``Note that this equation indicates that the response is
made immediately upon stimulus presentation. We therefore
are not explicitly modeling the time-extended cognitive
processes that govern evidence accumulation and the
corresponding response time.''

\subsection{AE comment 3}
What was the reason behind the choice to have equation 2
only be sensitive to negative feedback?

\subsubsection{Author response}
The rationale behind the choice to have equation 2 only be
sensitive to negative feedback is now stated explicitly in
the first two sentences of the paragraph that includes
equation 2. Those sentences read as follows.

``If positive feedback is received, then the value of the
criterion remains unchanged for the next trial (except for
drift – see equation 3). The rationale here is that if the
response is correct, then the observer has effectively
gained zero information about how their criterion should be
modified.''

\subsection{AE comment 4}
How is ``optimal'' defined in the current case, given that
response time does not factor in the update rule?

\subsubsection{Author response}
By ``optimal'', we mean the value of the criterion that
maximizes accuracy. We now say this explicitly in the
revision. This is now explained in the last sentence of the
paragraph that includes equation 2, which reads as follows.

``This updating rule will gradually converge on the optimal
criterion value (i.e., the value of the criterion that
maximizes accuracy).''

\subsection{AE comment 5}
The use of parameter space partitioning is sensible given
the many differences among the three models. However, PSP
operates on non-stochastic models. In addition, the way you
have approached PSP is not how the original PSP is
implemented (i.e., using MCMC). This should be mentioned
explicitly to avoid confusion.

\subsubsection{Author response}
Thanks for raising this important point. We have revised the
manuscript to make this point clear by adding the following
text  to the beginning of the third paragraph of the section
entitled ``Simulation Results.''

``Classical PSP is defined for non-stochastic models and is
often implemented by exploring the parameter space via a
Markov chain Monte Carlo algorithm. In contrast, we used
brute force search over a grid of plausible parameter
values. Because our models are stochastic, at each sampled
set of parameter values we generated multiple simulations
and classified the results into one of a small set of
qualitative outcome patterns.''

\subsection{AE comment 6}
The three-dimensional plots in figures 1B, 2B, and 3B do not
come out right on 2D. You could improve these by having 2D
projections on the planes. Better still, three boxplots
would be more appropriate as there is no additional
information by using 3D plots.

\subsubsection{Author response}
This was a good suggestion. We revised the two figures (8
and 9) that describe the simulation results accordingly --
i.e., both figures are 2D and include boxplots.

\subsection{AE comment 7}
Although not a focus in this manuscript, I wondered whether
the models are able to account for finer-grained data
patterns such as complete RT distributions and
speed-accuracy tradeoff functions.

\subsubsection{Author response}
As described in our response to AE comment 2, the models
make no attempt to account for response time (RT). The
models could all be generalized to make RT predictions, but
any such generalization would require adding extra
assumptions about the time course of the decision-making
process. The focus of our research was on \textit{what} is
updated trial-by-trial (i.e., a criterion vs SR
associations), rather than on the dynamics of the updating
process. For this reason, we chose a theoretically minimal
landscape. Specifically, we were concerned that adding any
extra assumptions that were not directly related to our
primary goal would weaken the inferences possible from our
model simulations. For example, it would be difficult to
rule out the possibility that any qualitative result we
established was a consequence of one of these added
assumptions, rather than because of a difference in what was
being updated on each trial. Even so, we agree that
extending the models proposed here to make RT predictions
could be a fruitful avenue for future work.

\subsection{AE comment 8}
The experiments are numbered 2 and 3, instead of 1 and 2.
The models were following the methods of experiment 1, which
is missing. Experiment 1 is also mentioned in the discussion
separately with experiment 2. It looks like experiment 1 has
been deleted from this manuscript, but the model simulations
and discussion still refer to it.

\subsubsection{Author response}
We thank the reviewer for catching this confusion. No study
was deleted from the manuscript. Rather, whether or not the
computational modeling section should be considered an
``experiment'' was a point of oscillation when writing. This
comment, along with comment 3 of Reviewer 1 and comment 2 of
Reviewer 2 helped us see that referring to the modeling
section as Experiment 1 was a mistake. As explained in our
responses to both of those comments, the revision refers to
the two empirical experiments as Experiments 1 and 2. These
are now presented first, and then following by a modeling
section entitled ``New Computational Models of Criterial
Learning.''

\subsection{AE comment 9}
In figure 6, it is clear that the participants form two
subcohorts. However, the use of a Gaussian seems odd given
that there can not be any negative values for
trials-to-criterion. Given that the authors model this
frequency histogram to exclude participants, more
appropriate distributions, such as Poisson should be used.

\subsubsection{Author response}
This is a good point that is similar to comment 3 of
Reviewer 2. Rather than use a mixture model based on the
Poisson distribution, we took the advice of Reviewer 2 and
used the ExGaussian as our base distribution. Specifically,
as described in our response to comment 3 of Reviewer 2, we
now fit the trials-to-criterion data collapsed across
conditions and participants with both a single-component
truncated ExGaussian and a two-component truncated
ExGaussian mixture via MLE with proper normalization on the
tasks bounded support $[9,512]$.

\subsection{AE comment 10}
Given the experimental results, what is the added benefit of
the three models? Experiment 2 is the critical experiment
showing that 3.5-0.5 leads to more trials-to-criterion than
0.5-3.5 or 0.5-0.5. The models do not provide any detailed
understanding of mechanisms and they are not fitted to
actual data. The general message from the experiments is
that any updating rule should be sensitive to feedback
delay. The models do not go beyond this.

\subsubsection{Author response}
The experiments show that feedback delay impairs criterial
learning, but they do not speak to the question of
underlying mechanism. The modeling is critical to this
question. Specifically, the computational analysis attempts
to establish constraints on underlying mechanisms that are
compatible with the empirical results. We believe that PSP
is a better approach, given this goal, than the more
traditional approach of fitting the models to the observed
data. This is now explained in the following new paragraph
(paragraph 2 of the ``Simulation Results'' section):

``Our approach was to generate predictions for each of the
three experimental conditions across a wide range of
parameter values. We then used parameter-space partitioning
(PSP) to evaluate the performance of each model (Pitt et
al., 2006). Traditional computational modeling identifies
parameter values that allow the model to provide the best
possible fit to the observed data. This traditional approach
therefore treats the single observed data set being fit as a
gold standard. Our goal instead was to investigate the range
of possible predictions of each model. We were interested in
questions such as: can a model predict any possible outcome
or is it constrained to always predict, for example, that
feedback delays impair learning? PSP was designed to answer
such questions.''

Furthermore, we believe this PSP analysis produced some
surprising and important results. First, it showed that
SR-learning accounts are compatible with our results -- even
if they assume that the amount of synaptic updating that
occurs trial-by-trial is unaffected by feedback delay.
Second, the PSP analysis showed that, in contrast,
stored-criterion accounts of Experiments 1 and 2 must assume
that feedback delays impair the criterion-updating process
-- a result that greatly constrains the type of memory that
could be used to store and update the criterion. For
example, it rules out working memory because many previous
results show that working memory is immune to short 3 s
delays. We believe these inferences would be impossible
without our computational modeling approach. In this way,
the models contribute the most important piece: they pin
down the process-level mechanisms consistent with the
experiments.

\end{document}

### 2025-10-17 write/main_revision.tex
updated analysis et al for revision

@@
 
-\title{Criterial Learning and Feedback Delay: Insights from
-Computational Models and Behavioral Experiments}
-\shorttitle{Computational Criterion Learning}
+\title{Is there a Criterion in Criterial Learning? \\ \vspace{.08in}
+Insights from Studying Feedback Delays
+}
+\shorttitle{Criterial Learning}
 
@@
      psychology, yet its cognitive and neural underpinnings
-    remain poorly understood. To address this shortcoming,
-    three computational models that capture different
-    hypotheses about criterial learning were developed and
-    tested. The time-dependent drift model assumes the
-    criterion is stored in working memory and that its value
-    drifts over time. The delay-sensitive learning model
-    assumes that the magnitude of criterial learning is
-    temporally discounted by feedback delay. The
-    reinforcement-learning model assumes that criterial
-    learning emerges from stimulus-response association
-    learning without an explicit representation of the
-    criterion, with learning rate also temporally discounted
-    by feedback delay. The performance of these models was
-    investigated under varying feedback delay and intertrial
-    interval (ITI) durations. The time-dependent drift model
-    predicted that long ITIs and feedback delays both impair
-    criterial learning. In contrast, the delay-sensitive and
-    reinforcement-learning models predicted impairments only
-    with feedback delays.  Two behavioral experiments tested
-    these predictions. In Experiment 1, participants were
-    explicitly instructed about the relevant dimension to
-    isolate criterial learning, and short feedback delays
-    impaired learning relative to immediate feedback. In
-    Experiment 2, using binary-feature stimuli without
-    instruction (isolating rule selection/switching),
-    neither longer ITIs nor delayed feedback affected
-    performance. These results support the delay-sensitive
-    and reinforcement-learning models, and suggest that even
-    in tasks that appear to rely on explicit, rule-based
-    reasoning, criterial learning may have strong
-    associative underpinnings.
+     remain poorly understood. Two experiments and extensive
+     computational modeling were used to test between two
+     strikingly different interpretations of the criterion.
+     The traditional account is that decisions are made by
+     comparing the stimulus value to a stored value of the
+     criterion. A conceptually different interpretation is
+     that learning instead is a process of associating
+     responses with stimuli, and that the criterion is
+     simply the hypothetical value that separates stimuli
+     associated with contrasting responses. The experiments
+     and modeling tested between these two interpretations
+     by contrasting the effects on criterial learning of
+     feedback delays versus increases in the duration of the
+     intertrial interval in a one-dimensional
+     category-learning task. The empirical results strongly
+     suggested that human criterial learning is sensitive to
+     feedback delay but not to the duration of the
+     intertrial interval. The computational modeling showed
+     that these results are compatible with a
+     stimulus-response learning account, and incompatible
+     with all versions of the stored-criterion account,
+     except for the subset of these models that explicitly
+     assume the criterial updating process is sensitive to
+     feedback delay.
 }
@@
 This article addresses this shortcoming through a
-combination of computational modeling and empirical data
-collection. Specifically, we develop and test three
-different computational models that make qualitatively
-different assumptions about how the criterion is learned.
-The models differ in the role they assign to working memory
-and in whether they treat the response criterion as a
-fundamental psychological construct, or instead assume that
-behavior is driven purely by stimulus-response associations
-without any criterion guiding responses. These models are
-then tested in two behavioral experiments. The modeling and
-empirical focus are on how feedback delays and the length of
-the intertrial interval (ITI) affect criterial learning. All
-criterial-learning models assume that updating (i.e.,
-learning) of the criterion occurs during the time interval
-between feedback presentation and the stimulus presentation
-that defines the onset of the next trial. So feedback delay
-and the length of the ITI are the independent variables that
-most clearly differentiate the conflicting predictions of
-criterial-learning models. Furthermore, feedback delays are
-known to impair some forms of learning (i.e., procedural)
-much more than others (i.e., learning that relies on
-declarative memory), so feedback-delay manipulations offer a
-powerful method of disambiguating the nature of criterial
-learning \parencite{ell2009critrial, MaddoxAshbyBohil2003,
-MaddoxIng2005}.
-
-Although the models could be tested in any task that depends
-on criterial learning, the two experiments we describe used
-a one-dimensional category-learning task. In such tasks,
-stimuli vary across trials on two or more dimensions -- one
-relevant to categorization and one or more that are
-irrelevant. The observer's goal is to identify the relevant
-dimension and learn the response criterion that maximizes
-accuracy. This task has been used in hundreds of studies,
-and all current models of performance in this task emphasize
-the role of criterial learning.
-
-In the first behavioral experiment, participants were
-explicitly instructed about the relevant dimension, thereby
-isolating criterial learning from rule selection and
-switching processes. The results showed that short feedback
-delays impaired learning compared to immediate feedback.
-The second behavioural experiment used stimuli composed of
-binary features -- where no criterial learning is needed --
-and did not instruct participants about the relevant
-dimension. Thus, this experiment isolated rule selection and
-switching from criterial learning. The results found that
-increasing feedback delay or ITI did not affect performance.
-These findings suggest that feedback delays impact criterial
-learning but not the discovery of the relevant stimulus
-dimension.  More broadly, they indicate that criterial
-learning may recruit procedural learning, even in tasks that
-seem to rely on explicit, rule-based processes.
-
-\section{New Computational Models of Criterial Learning}
-We developed computational models of criterial learning
-within two broad architectures and examined their
-sensitivity to ITI duration and feedback delay. The
-architectures differ in whether they assume that the
-response criterion is explicitly represented in memory. In
-the first architecture, the criterion is stored in memory,
-and responses are generated by comparing the current percept
-with this stored referent. The criterion is updated after
-errors using a gradient-descent rule. Models in this class
-can exhibit sensitivity to ITI and feedback delay in three
-distinct ways: (1) perceptual drift, where stimulus
-representations change over time, (2) criterial drift, where
-the stored criterion shifts over time, and (3)
-delay-sensitive updating, where the magnitude of criterion
-updates decreases with longer feedback delays. In the second
-architecture, no explicit criterion is stored.  Instead,
-reinforcement learning forms direct stimulus--response
-associations that drive performance.  Sensitivity to ITI and
-feedback delay arises from either perceptual drift or
-delay-sensitive updating, but criterial drift is not
-possible since no criterion is represented.
-
-For each of these models, denote the time of stimulus
-presentation on the current trial by T$_\text{S}$, the
-response time by T$_\text{R}$, the time when feedback is
-displayed by T$_\text{F}$, and the time when the stimulus
-that begins the next trial is displayed by T$_{\text{S}^+}$.
-Then note that the feedback delay equals $\text{t}_\text{FD}
-= \text{T}_\text{F} - \text{T}_\text{R}$ and the duration of
-the ITI equals $\text{t}_\text{ITI} = \text{T}_{\text{S}^+}
-- \text{T}_\text{F}$.
-
-\subsection{Models that assume a stored criterion}
-These models assume that the observer constructs a response
-criterion, maintains it in working memory, and makes
-decisions by comparing the percept to this criterion. They
-may further posit that both the stored criterion and the
-perceived stimulus value drift over time, with the extent of
-drift increasing both within a trial and across consecutive
-trials. In addition, they may assume that the learning rate
-for updating the criterion decreases as the feedback delay
-increases.
-
-Let $x_n(t)$ and $c_n(t)$ denote the values of the percept
-and the criterion, respectively, on trial $n$ at time $t$.
-Then the decision rule on trial $n$ is:
-\begin{equation}
-  \text{Respond } R_n =
-  \begin{cases}
-    \text{A}, & \text{if ~} x_n(\text{T}_\text{S}) \leq c_n(\text{T}_\text{S})  \\
-    \text{B}, & \text{if ~} x_n(\text{T}_\text{S}) > c_n(\text{T}_\text{S}).
-  \end{cases}
-  \label{eq:DR}
-\end{equation}
-Note that this equation indicates that the response is made
-immediately upon stimulus presentation. We therefore are not
-explicitly modelling the time-extended cognitive processes
-that govern evidence accumulation and the corresponding
-response time.
-
-If positive feedback is received, then the value of the
-criterion remains unchanged for the next trial (except for
-drift -- see Equation \ref{eq:criterion}).  The rationale
-here is that if the response is correct, then the observer
-has effectively gained zero information about how their
-criterion should be modified.  If negative feedback is
-received, then the criterion is modified according to the
-standard model
-\parencite{SuttonBarto1998}:
-\begin{equation}
-c_n(\text{T}_\text{F}+ \Delta_t) = c_n(\text{T}_\text{F}) + \frac{\alpha}{t_{FD}} [x_n(\text{T}_\text{F}) - c_n(\text{T}_\text{F})],
-\label{RL}
-\end{equation}
-where $\alpha$ is a learning-rate parameter, and $\Delta_t$
-is the time it takes to complete the updating.  It is
-straightforward to show that the iterative Equation \ref{RL}
-is equivalent to computing a weighted mean (weighted by
-recency) of the values of all percepts that occur on error
-trials \parencite[e.g.,][]{Ashby2017}. This updating rule
-will gradually converge on the optimal criterion value
-(i.e., the true criterion value the observer is trying to
-learn).
-
-The term $\frac{\alpha}{t_{FD}}$ captures the notion that
-the magnitude of the update decreases as the feedback delay
-increases.  As a result, when feedback is delayed, the
-system becomes less responsive to errors, leading to slower
-learning compared to immediate feedback conditions. However,
-since this model is consistent with the idea that criterial
-learning relies on working memory -- and the available
-evidence suggests that logical reasoning and working memory
-are unaffected by feedback delays of several seconds
-\parencite[e.g., in one-dimensional rule-based category
-learning tasks;][]{ell2009critrial, MaddoxAshbyBohil2003,
-MaddoxIng2005} -- the learning process described in Equation
-\ref{RL} likewise be unaffected by feedback delay. In this
-case, the term $\frac{\alpha}{t_{FD}}$ is replaced by
-$\alpha$, making the learning rate independent of feedback
-delay.
-
-This model further assumes that both the stimulus and the
-criterion representations may drift randomly throughout the
-duration of time they are maintained in working memory. The
-representation of the criterion must always be maintained in
-working memory, whereas drift in the stimulus representation
-affects performance up until the feedback is presented, but
-not afterwards. We modeled the drift in both the criterion
-and the percept by adding white noise to their initial
-values. Specifically, we assumed that for all
-$t>\text{T}_\text{S}$
-\begin{equation}
-  c_n(t) = c_n(\text{T}_\text{S}) + \eta_c \epsilon(t),
-  \label{eq:criterion}
-\end{equation}
-and
-\begin{equation}
-  x_n(t) = x_n(\text{T}_\text{S}) + \eta_x \epsilon(t),
-  \label{eq:percept}
-\end{equation}
-where $\epsilon(t)$ is white noise and $\eta_c$ and $\eta_x$
-are parameters that determine the amount of drift over time.
-Note that these parameters are permitted to be equal to zero
-which is equivalent to assuming that the criterion and/or
-the percept do not drift over time at all.
-
-This model predicts that at the time of feedback, the
-response criterion $c_n(\text{T}_\text{F})$ will be normally
-distributed with mean $c_n(\text{T}_\text{S})$ and variance
-$\text{t}_\text{FD} \eta_c^2$. Similarly, at the time when
-the stimulus that defines the next trial is presented, the
-criterion $c_n(\text{T}_{\text{S}^+}) =
-c_{n+1}(\text{T}_\text{S})$ is normally distributed with
-mean $c_n(\text{T}_\text{S})$ and variance
-$(\text{t}_\text{FD}+\text{t}_\text{ITI}) \eta_c^2$. The
-predictions for the percept are similar. Specifically, at
-the time when feedback is presented, the percept
-$x_n(\text{T}_\text{F})$ is normally distributed with mean
-$x_n(\text{T}_\text{S})$ and variance
-$(\text{T}_\text{F}-\text{T}_\text{S}) \eta_x^2$.
-
-\subsection{Reinforcement-learning models}
-The reinforcement-learning models gradually associate
-responses with stimuli, and therefore do not include an
-explicit representation of the response criterion. The
-models are based on an actor-critic architecture and use
-reinforcement learning to form stimulus-response
-associations \parencite{SuttonBarto1998}. These models
-assumes that the perceptual representation of each stimulus
-is defined by the pattern of activation across 25 sensory
-units that are characterized by overlapping tuning curves.
-Each unit is maximally excited by one specific stimulus,
-which we call the unit's preferred stimulus. Specifically,
-the activation of the $i^\text{th}$ sensory unit on trial
-$n$ is given by
-\begin{equation}
-  A_i(n) = \text{exp} \left( \frac{-\left[d_{i,S_n}\right]^2}{\sigma^2} \right)
-  \label{eq:DSL_activation}
-\end{equation}
-where $d_{i,S_n}$ is the (Euclidean) distance between the preferred
-stimulus value of the $i^\text{th}$ sensory unit and the
-value of the stimulus that was presented on trial $n$, and
-$\sigma$ is a constant that increases with perceptual noise.
-
-The model includes two decision or actor units -- one
-associated with each of the two possible responses.
-Initially, each sensory unit is connected to both actor
-units with some random connection strength. Let
-$\omega_{iJ}(n)$ denote the strength of the connection
-between sensory unit $i$ and actor unit $J$ (for $J$ = A or
-B) on trial $n$. Then the activation in actor unit $J$ on
-trial $n$, denoted by $V_J(n)$, equals
-\begin{equation}
-  V_J(n) = \sum_{i} \omega_{iJ}(n) A_i(n)
-\end{equation}
-Responses are generated by the decision rule:
-\begin{equation}
- \text{Respond } R_n =
-  \begin{cases}
-    A, & \text{if $V_A(n) > V_B(n)$}\\
-    B, & \text{if $V_A(n) \leq V_B(n).$}
-  \end{cases}
-\end{equation}
-The connection strengths $\omega_{iA}(n)$ and
-$\omega_{iB}(n)$ are updated after feedback is received on
-each trial according to standard reinforcement learning
-rules \parencite{SuttonBarto1998}:
-\begin{equation}
-  \omega_{iA}(n) = \omega_{iA}(n-1) + \frac{\alpha_{actor}}{t_\text{FD}} \delta(n-1)
-\end{equation}
-\begin{equation}
-  \omega_{iB}(n) = \omega_{iB}(n-1) + \frac{\alpha_{actor}}{t_\text{FD}} \delta(n-1),
-\end{equation}
-where $\alpha_{actor}$ is the learning-rate of the actor, and
-$\delta(n-1)$ is the reward prediction error on trial $n-1$.
-Note that as in Equation~\ref{RL}, the learning rate is
-scaled by the inverse of the feedback delay.
-
-Much evidence suggests that this type of stimulus-response
-learning is mediated largely within the striatum, and is
-facilitated by a dopamine (DA) mediated reinforcement
-learning signal that is time dependent
-\parencite[e.g.,][]{ValentinMaddoxAshby2014}. Specifically,
-the dopamine signal generated by positive feedback appears
-to peak at around 500 ms after feedback and then decay back
-to baseline levels within 2 or 3 s
-\parencite{YagishitaEtAl2014}. As a result, synaptic
-plasticity at cortical-striatal synapses is attenuated with
-increasing feedback delays \parencite{YagishitaEtAl2014}.
-The scaling of $\alpha_{actor}$ by ${1}{t_\text{FD}}$ is one
-way to model this phenomenon. However, it is possible that
-criterial learning via stimulus-response associations is
-mediated by other brain systems that are less sensitive to
-feedback delay. In this case, the learning rate would be
-independent of feedback delay, and the term
-$\frac{\alpha_{actor}}{t_\text{FD}}$ would be replaced by
-$\alpha_{actor}$.
-
-The reward prediction error on trial $n-1$ is defined as the
-value of the obtained reward, denoted by $R(n-1)$ minus the
-value of the predicted reward, denoted by $P(n-1)$:
-\begin{equation}
-  \delta(n-1) = R(n-1) - P(n-1).
-\end{equation}
-The predicted reward on trial $n$ is determined by the
-critic via
-\begin{equation}
-  P(n) = P(n-1) + \frac{\alpha_{critic}}{t_\text{FD}} \delta(n-1),
-\end{equation}
-where $\alpha_{critic}$ is the learning rate of the critic.
-This learning rate may again be scaled by the inverse of the
-feedback delay. If sensitivity to feedback delay is not
-assumed in Equation~\ref{RL}, then the term
-$\frac{\alpha_{critic}}{t_\text{FD}}$ is replaced by
-$\alpha_{critic}$.
-
-Because this class of models does not store an explicit
-response criterion, it cannot incorporate criterial drift.
-It can, however, incorporate perceptual drift. In this case,
-the value of the current stimulus used to compute sensory
-unit activation (Equation~\ref{eq:DSL_activation}) is
-assumed to drift over time according to
-\begin{equation}
-  S_n(t) = S_n(\text{T}_\text{S}) + \eta_S \epsilon(t),
-  \label{eq:stimulus_drift_RL}
-\end{equation}
-where $\eta_c$ scales the magnitude of the drift and
-$\epsilon(t)$ is a noise process.
-
-\subsection{Simulation Results}
-We investigated how changing the duration of the feedback
-delay and ITI affect criterial learning for each class of
-model. More specifically, we simulated performance of
-each model in a categorization task that included two
-categories of stimuli that varied on a single stimulus
-dimension and that could be categorized perfectly by
-comparing each stimulus to an appropriate value of the
-response criterion (We used the same experimental design as
-in Experiment 1, so see those methods for more details). Our
-simulations included three experimental conditions: Delayed
-Feedback, Long ITI, and Control. The Delayed-Feedback
-condition included a long feedback delay (3.5 s) and a short
-ITI (0.5 s). The Long-ITI condition matched the total trial
-duration of the Delayed-Feedback condition but with a short
-feedback delay (0.5 s) and a long ITI (3.5 s). The Control
-condition included both a short feedback delay (0.5 s) and a
-short ITI (0.5 s). Each simulation continued for 200 trials
-or until the model responded correctly for 12 trials in a
-row. For each set of parameter values, we simulated the
-model 100 times and averaged the results, yielding one
-observed performance metric measured in trials-to-criterion
-for each condition.
-
-Our approach was to generate predictions for each of the
-three experimental conditions across a wide range of
-parameter values. We then used parameter-space partitioning
-(PSP) to evaluate the performance of each model
-\parencite[PSP;][]{pitt2006global}.  Classical PSP is
-defined for non-stochastic models and is often implemented
-by exploring the parameter space via MCMC.  In contrast, we
-used brute force search over a grid of plausible parameter
-values. Because our models are stochastic, at each sampled
-parameter setting we generated multiple simulations and
-classified the expected qualitative pattern.  PSP calculates
-the proportion of the parameter space where a model makes
-specific qualitative predictions. We focused on four such
-predictions. The first was that learning under feedback
-delay would be slower than in the other two conditions. The
-second was that learning in the Long ITI condition would be
-slower than in the other two. The third was that learning in
-the control condition would be slower than in either of the
-other two conditions. The fourth was a catch-all category
-encompassing any other possible patterns of results. For
-each model, the PSP analysis quantified the proportions of
-the explored parameter space where the model made these
-qualitative predictions.
-
-\subsubsection{Models that assume a stored criterion}
-We investigated predictions of these models across a wide
-range of values for the parameters $\eta_{x}$, $\eta_{c}$,
-and $\alpha$.  Specifically, in the case of $\alpha$,  we
-stepped through every value in the interval $[0, 0.2]$, with
-a step size of .01.  In the case of both $\eta_{x}$, and
-$\eta_{c}$, we searched over the interval $[0, 1]$, with a
-step size of $0.1$.  We also explored the binary case of
-whether or not the update rate $\alpha$ was scaled by the
-feedback delay. The results are shown in Figure
-\ref{fig:model_class_II_results}.
+combination of empirical data collection and computational
+modeling. Specifically, our goal is to test between two
+general possibilities. One is that the criterion is learned
+in the sense that it is updated trial-by-trial and that its
+current value is stored in memory. This is the hypothesis
+implicitly assumed by most decision-making models (e.g.,
+signal-detection theory). The second possibility is that the
+criterion has no psychological meaning and learning instead
+is of stimulus-response (SR) associations. According to this
+account, the criterion is simply the sensory (or cognitive)
+value that separates percepts associated with the
+contrasting responses. For example, there is evidence that
+procedural learning works in this way
+\parencite{AshbyWaldron1999}.
+
+Two independent variables that seem especially likely to
+discriminate between these two alternatives are the duration
+of the intertrial interval (ITI) and the delay between the
+response and the feedback. If the value of a criterion is
+updated following feedback and then held in memory until
+needed again, then the longer it must be held in memory, the
+more time there is for its value to drift. In other words,
+there should be more criterial drift with long ITIs than
+with short ITIs, and as a result, increasing the ITI should
+impair performance. In contrast, if the criterion has no
+psychological meaning and instead learning is of SR
+associations, then learning should be impaired by increases
+in feedback delay. The idea is that positive feedback
+strengthens and negative feedback weakens recently active
+synapses that associate a response with the presented
+stimulus. As a result, increasing the feedback delay will
+weaken the trace of recently active synapses, and therefore
+reduce this type of SR learning. This model is supported by
+many previous reports that feedback delays as short as 3 s
+impair procedural learning much more than  declarative
+learning that depends on executive attention and working
+memory  \parencite{DunnEtAl2012, ell2009critrial,
+MaddoxAshbyBohil2003, MaddoxIng2005,  Worthyetal2013}.
+
+Our general approach is as follows. First, we establish an
+empirical database by describing the results of two
+experiments that report the results of manipulating ITI and
+feedback delay on criterial learning. Although a variety of
+different tasks could be used to study criterial learning,
+the two experiments we describe used a one-dimensional
+category-learning task. In such tasks, stimuli vary across
+trials on two or more dimensions -- one relevant to
+categorization and one or more that are irrelevant. The
+observer's goal is to identify the relevant dimension and
+learn the response criterion that maximizes accuracy. This
+task has been used in hundreds of studies, and all current
+models of performance in this task emphasize the role of
+criterial learning.
+
+In the first experiment, participants were explicitly
+instructed about the relevant dimension, thereby isolating
+criterial learning from rule selection and switching
+processes. The results showed that increasing the feedback
+delay impaired learning but increasing the ITI did not.
+Experiment 2 used stimuli that varied on six binary-valued
+dimensions and participants were not told which of these
+dimensions was relevant to the categorization decision. The
+observer's task was to discover the single relevant
+dimension, but after this dimension was identified, no
+criterial learning was required since there were two
+categories and the stimuli had exactly two values on the
+relevant dimension. Thus, Experiment 2 isolated rule
+selection and switching from criterial learning. The results
+showed that performance was unaffected either by increases
+in feedback delay or ITI. These findings therefore suggest
+that feedback delays impact criterial learning but not the
+discovery of the relevant stimulus dimension.
+
+The results of Experiments 1 and 2 seem to suggest that
+criterial learning is a form of procedural learning, even in
+tasks that seem to rely on explicit, rule-based processes,
+and therefore that the criterion is an emergent construct
+with no psychological meaning. Before accepting this
+conclusion however, it is important to ask whether some type
+of criterion-learning model can account for these results.
+To investigate this question, we developed a variety of
+computational models of criterial learning of two different
+types. Eight different models assumed the criterion is
+learned and updated trial-by-trial. These were created by
+factorially combining three different binary assumptions:
+the percept drifts randomly over time (yes or no), the
+criterion drifts randomly over time (yes or no), and the
+updating of the criterion is sensitive to the feedback delay
+(yes or no). Four different models assumed that learning is
+of SR associations and that the criterion therefore has no
+psychological meaning. These were created by factorially
+combining two binary assumptions: the percept drifts
+randomly over time (yes or no) and SR strengthening is
+sensitive to feedback delay (yes or no). Next, using a
+technique called parameter-space partitioning
+\parencite{pitt2006global}, we investigated whether each
+model could account for the qualitative results of
+Experiments 1 and 2. The results of these analyses
+reinforced the conclusion that criterial learning may have
+strong associative underpinnings, even in tasks that appear
+to rely on explicit, rule-based reasoning,
+
+\section{Experiment 1}
+Experiment 1 used a one-dimensional category-learning task
+to investigate how feedback delay and ITI affect criterial
+learning. The stimuli in this experiment were circular
+sine-wave gratings that varied across trials in bar width
+and bar orientation. These stimuli were divided into two
+categories according to their value on one of the two
+dimensions. In other words, the optimal strategy was to set
+a response criterion on the single relevant dimension, and
+then choose a categorization response based on whether the
+value of the presented stimulus on this dimension was larger
+or smaller than the criterion value.  The experiment
+isolated criterial learning by (1) explicitly instructing
+participants about the relevant stimulus dimension and rule
+structure (e.g., thick bars are ``A'', thin bars are ``B''),
+and (2) eliminating all variability along the irrelevant
+stimulus dimension. In other words, the instructions
+identified the relevant stimulus dimension, and as a result,
+the only learning required was criterial learning.
+
+Each participant practiced each of 14 one-dimensional
+category-learning tasks, or problems, until they responded
+correctly on 9 of the 10 previous trials, at which point the
+problem changed. The 14 different category structures are
+described in Figure \ref{fig_design_exp_1_space}. The
+relevant dimension was bar thickness in problems 1 -- 7 and
+bar angle in problems 8 -- 14. Each problem included stimuli
+in two distinct clusters that varied over a restricted range
+of the relevant dimension. Critically, the optimal criterion
+value varied from problem-to-problem with respect to its
+position within the stimulus range. For example, for some
+problems the optimal criterion was below the midpoint of the
+range and for other problems it was above the midpoint.
+Feedback delay and ITI were varied across three conditions.
+In the Delayed-Feedback condition, feedback was delayed
+after the observer responded (by 3.5 s) and the ITI was
+short (500 ms). In the Long-ITI condition, the feedback
+delay was short (500 ms) and the ITI was long (3.5 s).
+Finally, in the Control condition, feedback delay and the
+ITI were both 500 ms.
 
@@
   \centering
-  \includegraphics[width=1\textwidth]{../figures/model_new_class_II.pdf}
+  \includegraphics[width=.5\textwidth]{../figures/fig_design_exp_1_space.png}
   \caption{
-      \textbf{A:} Assuming a delay-sensitive update, the
-      proportion of parameter space where the models that
-      assume a stored criterion predicted each of four
-      qualitative patterns: (1) slower learning under
-      feedback delay, (2) slower learning in the Long-ITI
-      condition, (3) slower learning in the control
-      condition, and (4) any other pattern. 
-      % 
-      \textbf{B:} Assuming a delay-sensitive update, Boxplot
-      of the parameter ranges leading to each PSP pattern.
-      All parameter values were normalized by the largest
-      value in the search range, so the ordinate ranges from
-      zero to one for all parameters.
-      % 
-      \textbf{C:} Assuming no delay-sensitive update, the
-      proportion of parameter space where the models that
-      assume a stored criterion predicted each of four
-      qualitative patterns: (1) slower learning under
-      feedback delay, (2) slower learning in the Long-ITI
-      condition, (3) slower learning in the control
-      condition, and (4) any other pattern. 
-      % 
-      \textbf{D:} Assuming no delay-sensitive update, Boxplot
-      of the parameter ranges leading to each PSP pattern.
-      All parameter values were normalized by the largest
-      value in the search range, so the ordinate ranges from
-      zero to one for all parameters.
-      %
-      \textit{Note:} In all panels, color indicates the PSP
-      pattern.
+      Category sample space. Different colors represent
+      different category problems. Dashed lines are category
+      boundaries (criterion) and the surrounding solid lines
+      mark the no-stimulus region in which no stimuli were
+      sampled.
 }
-  \label{fig:model_class_II_results}
+  \label{fig_design_exp_1_space}
 \end{figure}
-
-Figure~\ref{fig:model_class_II_results} shows that this
-class of models makes qualitatively different predictions
-depending on whether the update rule is sensitive to
-feedback delay. When delay sensitivity is assumed (panels A
-and B), the model consistently predicts that feedback delays
-impair criterion learning. When delay sensitivity is not
-assumed (panels C and D), the model instead predicts that
-long ITIs impair learning, except in cases with very small
-learning rates $\alpha$ (panel D).
-
-\subsubsection{Reinforcement-learning models}
-We investigated the effects on performance predicted by the
-reinforcement-learning model of three parameters -- the
-perceptual-noise variance $\sigma^2$ and the actor and
-critic learning rates (i.e., $\alpha_{actor}$ and
-$\alpha_{critic}$, respectively). In the case of both
-$\alpha_{actor}$ and $\alpha_{critic}$, we stepped through
-every value in the interval $[0, 0.2]$, with a step size of
-$.02$. We constrained our search over these parameters to this
-interval because reinforcement-learning models are prone to
-instability at very high learning rates
-\parencite{SuttonBarto1998}. As evidence of this, at higher
-learning rates, the model failed to learn with any
-consistency -- that is, in most cases, it failed to reach
-the learning criterion (12 correct responses in a row)
-within the allowable 200 trials. In the case of $\sigma$, we
-searched over the interval $[1, 10]$, with a step size of
-$3$.  We also explored the binary case of whether or not the
-update rate $\alpha$ was scaled by the feedback delay. As
-with the other models, all scores were normalized by the
-largest observed value.
-
-The results are described in
-Figure~\ref{fig:model_class_I_results}.  Note this model
-predicts that delayed feedback impairs criterial learning
-more than a long ITI across the vast majority of parameters
-explored. As can be seen in
-Figure~\ref{fig:model_class_I_results}D, the few exceptions
-tend to occur when perceptual drift has relatively small
-values.
-
-\begin{figure} 
-  \centering
-  \includegraphics[width=1\textwidth]{../figures/model_new_class_I.pdf}
-  \caption{ 
-      \textbf{A:} Assuming a delay-sensitive update, the
-      proportion of parameter space where the models that
-      assume a stored criterion predicted each of four
-      qualitative patterns: (1) slower learning under
-      feedback delay, (2) slower learning in the Long-ITI
-      condition, (3) slower learning in the control
-      condition, and (4) any other pattern. 
-      % 
-      \textbf{B:} Assuming a delay-sensitive update, Boxplot
-      of the parameter ranges leading to each PSP pattern.
-      All parameter values were normalized by the largest
-      value in the search range, so the ordinate ranges from
-      zero to one for all parameters.
-      % 
-      \textbf{C:} Assuming no delay-sensitive update, the
-      proportion of parameter space where the models that
-      assume a stored criterion predicted each of four
-      qualitative patterns: (1) slower learning under
-      feedback delay, (2) slower learning in the Long-ITI
-      condition, (3) slower learning in the control
-      condition, and (4) any other pattern. 
-      % 
-      \textbf{D:} Assuming no delay-sensitive update, Boxplot
-      of the parameter ranges leading to each PSP pattern.
-      All parameter values were normalized by the largest
-      value in the search range, so the ordinate ranges from
-      zero to one for all parameters.
-      %
-      \textit{Note:} In all panels, color indicates the PSP
-      pattern.
-}
-  \label{fig:model_class_I_results}
-\end{figure}
-
-\subsubsection{Summary of modeling results}
-We examined two classes of criterial-learning models that
-make different predictions about the effects of feedback
-delay and ITI. The first class assumes that the criterion is
-stored in memory and compared against incoming stimuli to
-generate responses. These models predict greater impairment
-from feedback delay only when the update rule is explicitly
-delay-sensitive. Otherwise, they are generally more impaired
-by long ITIs than by feedback delay. The second class
-assumes no stored criterion, with responses driven instead
-by reinforcement learning through gradual tuning of
-stimulus–response associations. These models almost
-uniformly predict greater impairment from feedback delay
-than from long ITIs.
-
-\section{Experiment 1}
-Experiment 1 investigated how feedback delay and the length
-of the ITI affect criterial learning in humans. As a result,
-it also provides rigorous tests of the predictions of the
-time-dependent drift model, the delay-sensitive learning
-model, and the reinforcement-learning model.
-
-The stimuli in Experiment 1 were circular sine-wave gratings
-that varied across trials in bar width and bar orientation.
-These stimuli were divided into two categories according to
-their value on one of the two dimensions. In other words,
-the optimal strategy was to set a response criterion on the
-single relevant dimension, and then choose a categorization
-response based on whether the value of the presented stimuli
-on this dimension was larger or smaller than the criterion
-value.
-
-The experiment isolated criterial learning by (1) explicitly
-instructing participants on the relevant stimulus dimension
-and rule structure (e.g., thick bars are ``A'', thin bars
-are ``B''), and (2) eliminating variability along the
-irrelevant stimulus dimension. In other words, the
-instructions identified the relevant stimulus dimension, and
-as a result, the only learning required was criterial
-learning.
 
@@
 \textcite{treutwein1989perceptual}, which roughly equate the
-salience of each dimension (for details, see also
-\textcite{CrossleyAshby2015}).
+salience of each dimension. For details, see also
+\textcite{CrossleyAshby2015}.
 
@@
 relevant dimension. This was spatial frequency for the
-problems 1 thorugh 7 and orientation for problems 8 through
+problems 1 through 7 and orientation for problems 8 through
 14. We then randomly selected one of the 7 bins defined on
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.5\textwidth]{../figures/fig_design_exp_1_space.png}
-  \caption{
-      Category sample space. Different colors represent
-      different category problems. Dashed lines are category
-      boundaries (criterion) and the surrounding solid lines
-      mark the no-stimulus region in which no stimuli were
-      sampled.
-}
-  \label{fig_design_exp_1_space}
-\end{figure}
-
 \subsubsection{Procedure}
@@
 \subsection{Results}
-Figure \ref{fig_mm_hist_1} shows a relative frequency
-histogram of the trials-to-criterion observed across all
-three conditions. The histogram shows that the majority of
-participants were able to learn each problem on average in
-less than 100 trials. However, this histogram also shows
-that a subset of participants required many more trials to
-learn each problem. Given that each problem is very simple
-and that the relevant dimension of each problem is
-explicitly instructed to participants, it is likely that the
-participants in the tails of this distribution did not pay
-attention to the instructions, were not motivated to learn
-the task, were distracted by other factors, or should be
-considered outliers for some other reason.
+Both panels of Figure \ref{fig_mm_hist_1} show the same
+relative frequency histogram of the trials-to-criterion
+collapsed across all participants and conditions. The
+histogram shows that the majority of participants were able
+to learn each problem on average in less than 100 trials.
+However, this histogram also shows that a subset of
+participants required many more trials to learn each
+problem. Given that participants were explicitly instructed
+about the relevant dimension of each problem, the task was
+relatively simple and as a result, it seems likely that the
+participants represented by the tails of this distribution
+did not pay attention to the instructions, were not
+motivated to learn the task, were distracted by other
+factors, or should be considered outliers for some other
+reason. In our experience, it is not unusual for a small
+subset of participants to perform very poorly in experiments
+of this kind. This typically reflects a lack of engagement
+or misunderstanding of the task instructions rather than an
+inability to perform the task itself.
 
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../figures/fig_exp_1_mm_trunc_exgauss_compare.png}
+  \includegraphics[width=1\textwidth]{../figures/fig_exp_1_mm_trunc_exgauss_compare.png}
   \caption{
       Relative frequency histogram of the
-      trials-to-criterion observed across all three
-      conditions in Experiment 1 and the fit of a single
-      Exponential-Gaussian distribution (Left) and a
-      two-component Exponential-Gaussian mixture model
-      (Right). In the right panel, the blue dashed line is
-      the first higher performance component, the orange
-      dashed line is the second lower performance component.
-      The gray solid line is the sum of the two components.
-      Rug plots along the x-axes show the individual data
-      colored by their assignment to either the higher or
-      lower performance component.  The two-component model
-      provided a significantly better fit according to AIC,
-      BIC, and a likelihood ratio test.
+      trials-to-criterion collapsed across all three
+      conditions and participants in Experiment 1. The left
+      panel shows the best-fitting single ExGaussian
+      distribution and the right panel shows the
+      best-fitting two-component ExGaussian mixture model.
+      In the right panel, the blue dashed line is the first
+      higher performance component, the orange dashed line
+      is the second lower performance component.       The
+      gray solid line is the mixture distribution that
+      combines these two components. Rug plots along the
+      x-axes show the individual data      colored by their
+      assignment to either the higher or lower performance
+      component. The two-component model provided a
+      significantly better fit according to AIC, BIC, and a
+      likelihood ratio test.
 }
@@
 
-We modeled the distribution shown in Figure
-\ref{fig_mm_hist_1} using a mixture of Exponential-Gaussian
-distributions with two components, each truncated to the
-range of possible trials-to-criterion values (i.e., $[9,
-512]$). The first component captured participants with a
-relatively low mean number of trials-to-criterion, while the
-second component identified outliers according to the above
-rationale.  We estimated this model via MLE with proper
-normalization on the tasks bounded support and compared them
-via AIC, BIC, and a parametric bootstrap likelihood-ratio
-test (LRT).  We compared the MLE fit of this two-component
-model to a single-component Exponential-Gaussian model using
-the Akaike Information Criterion (AIC) and Bayesian
-Information Criterion (BIC), both of which indicated a
-better fit for the two-component model 
-(
-$\mathrm{AIC}_1=692.15$ vs. $\mathrm{AIC}_2=652.81$; 
-$\mathrm{BIC}_1=698.33$ vs. $\mathrm{BIC}_2=667.23$
-)
-Additionally, a bootstrap likelihood ratio test confirmed
-that the two-component model provided a significantly better
-fit than the one-component model ($D=42.84$,
-$p_{\text{boot}}\approx 0.0099$) Participants that were
-deemed to most likely to belong in the low-performance
-Gaussian by this analysis were excluded from further
-analysis. After making these exlcusions, we were left with
-the following sample sizes:  Control condition ($ N = 11$);
-Delayed-Feedback condition ($N = 14$); Long-ITI condition
-($N = 17$).
+We identified and removed from further analyses these
+outlier participants using the following algorithm. First,
+we modeled the distribution shown in Figure
+\ref{fig_mm_hist_1} using a mixture of two ExGaussian
+distributions,\footnote{The ExGaussian is the distribution
+of the sum of two independent random variables -- one with
+an exponential distribution and one with a normal or
+Gaussian distribution.} each truncated to the range of
+possible trials-to-criterion values (i.e., $[9,512]$). The
+first distribution or component was assumed to model
+performance of participants with a relatively low mean
+number of trials-to-criterion, whereas the second component
+was assumed to identify outlier participants. We estimated
+parameters of this model using maximum likelihood estimation
+with proper normalization on the tasks bounded support and
+compared them via the AIC and BIC goodness-of-fit
+statistics, and a parametric bootstrap likelihood-ratio
+test. We then compared the fit of this two-component model
+to a single-component ExGaussian model using AIC and BIC,
+both of which indicated a better fit for the two-component
+model ($\mathrm{AIC}_1=692.15$ vs. $\mathrm{AIC}_2=652.81$;
+$\mathrm{BIC}_1=698.33$ vs. $\mathrm{BIC}_2=667.23$). The
+bootstrap likelihood-ratio test also confirmed that the
+two-component model provided a significantly better fit than
+the one-component model ($D=42.84$, $p_{\text{boot}}\approx
+.01$). Next, for each participant, we used the
+best-fitting two component model to compute the likelihood
+that their performance was best described by each component.
+Participants whose data were estimated to most likely belong
+to the low-performance distribution were classified as
+outliers and excluded from further analysis. After making
+these exclusions, we were left with the following sample
+sizes:  Control condition ($ N = 11$); Delayed-Feedback
+condition ($N = 14$); Long-ITI condition ($N = 17$).
 
 Figure \ref{fig_exp_1_t2c}A shows the mean
-trials-to-criterion in each condition of Experiment 1.  A
-one-way ANOVA revealed a significant effect of condition
-($F(2, 39) = 6.17$, $p < .01$, $\eta^2 = .24$) and planned
-comparisons revealed that performance in the
-Delayed-Feedback condition was significantly worse than in
-either the Control condition ($t(23.00) = 2.70$, $p < .05$)
-or the Long-ITI condition ($t(20.96) = 2.94$, $p < .01$).
+trials-to-criterion for all non-excluded participants in
+each condition of Experiment 1.  A one-way ANOVA revealed a
+significant effect of condition ($F(2, 39) = 6.17$, $p <
+.01$, $\eta^2 = .24$) and planned comparisons revealed that
+performance in the Delayed-Feedback condition was
+significantly worse than in either the Control condition
+($t(23.00) = 2.70$, $p < .05$, $d=1.06$) or the Long-ITI
+condition ($t(20.96) = 2.94$, $p < .01$, $d=1.10$).
 Performance in the Control and Long-ITI conditions did not
-differ significantly from each other ($t(17.99) = 0.22$, $p
-= .83$).
+differ significantly from each other ($t(17.99) = .22$, $p
+= .83$, $d=.09$). \footnote{Including the excluded outliers
+in these analyses eliminates the difference between
+conditions.}
 
@@
 was significantly worse than in the Control condition
-($t(18.54) = -3.28$, $p < .01$) and worse than in the Long
-ITI condition ($t(21.60) = -2.14$, $p < .05$).  Performance
-in the Control and Long-ITI conditions did not differ
-significantly from each other ($t(25.97) = -1.50$, $p=.15$).
+($t(18.54) = -3.28$, $p < .01$, $d=-1.21$) and worse than in
+the Long ITI condition ($t(21.60) = -2.14$, $p < .05$,
+$-.80$).  Performance in the Control and Long-ITI conditions
+did not differ significantly from each other ($t(25.97) =
+-1.50$, $p=.15$, $d=-.53$).
 
@@
     \caption{
-        \textbf{A}: Mean trials to criterion in each
+        \textbf{A}: Mean trials-to-criterion in each
         condition of Experiment 1. Error bars are standard
@@
 \section{Experiment 2}
-Experiment 2 was designed to reinforce the findings of
-Experiment 1 by using the same design but with stimuli that
-have binary-valued dimensions, thereby eliminating criterial
-learning. Unlike Experiment 1, we did not explicitly
-instruct participants on the relevant stimulus dimension in
-Experiment 2.  While Experiment 1 isolated criterial
-learning from rule selection and rule switching, this
-experiment isolates rule selection and switching from
-criterial learning. Therefore, Experiment 2 was designed to
-test whether the impaired learning that we observed in
-Experiment 1 when the feedback was delayed can be attributed
-principally to criterial learning or whether it could be due
-to some more general category-learning process.
+According to a popular model of one-dimensional rule-based
+category learning \parencite[i.e.,
+COVIS;][]{AshbyCOVIS1998}, successful learning requires at
+least three separate cognitive processes: selecting a rule
+for testing, switching from one candidate rule to another
+(following negative feedback), and criterial learning.
+Experiment 1 isolated criterial learning from this set.
+Experiment 2 complements Experiment 1 by using a task that
+required rule selection and rule switching, but no criterial
+learning.
+
+Experiment 2 used colored geometric figures as stimuli,
+which varied across trials on six binary dimensions (i.e.,
+see Figure \ref{fig_design_exp_2}B). As in Experiment 1,
+only one dimension was relevant. Unlike Experiment 1,
+however, participants were given no instruction as to which
+dimension was relevant. Therefore, this task required
+participants to discover the relevant dimension. However,
+after this dimension was identified, no criterial learning
+was required because each stimulus always had one of two
+possible values on the relevant dimension and participants
+were instructed that there were two categories. Experiment 2
+was therefore designed to test whether the impaired learning
+that we observed in Experiment 1 when the feedback was
+delayed can be attributed principally to criterial learning
+or whether it could be due to some more general
+category-learning process.
 
@@
 \subsection{Results}
-Figure \ref{fig_mm_hist_2} shows a relative frequency
-histogram of the trials-to-criterion observed in both
-conditions. The histogram shows that the majority of
-participants were able to learn each problem on average in
-less than 100 trials.  Unlike Experiment 1, there were no
-highly suspicious outliers in this data set. Nevertheless,
-for symmetry with Experiment 1, we performed the same
-Gaussian mixture model analysis as used there. The
-two-component model provided a slightly smaller AIC value
-(324.75 vs. 317.18), and a slightly larger BIC value (329.33
-vs. 327.87). The bootstrap likelihood-ratio test was not
-significant ($D = 11.41, p_{\text{boot}} \approx 0.2277$).
-We therefore did not exclude any participants from further
-analysis in Experiment 1.
+Figure \ref{fig_mm_hist_2} shows the trials-to-criterion
+relative frequency histogram collapsed across participants
+and conditions. As in Experiment 1, the histogram shows that
+the majority of participants were able to learn each problem
+in well under 100 trials. Unlike Experiment 1, however,
+there were no highly suspicious outliers in this data set.
+Nevertheless, for symmetry with Experiment 1, we performed
+the same ExGaussian mixture model analysis as used there.
+The two-component model provided a slightly smaller AIC
+value (317.07 vs 325.38), and a slightly smaller BIC value
+(327.75 vs 329.96). The bootstrap likelihood-ratio test was
+not significant ($D = 13.05, p_{\text{boot}} \approx .08$).
+Given the weaker evidence for multiple components in this
+data set compared to Experiment 1, and the fact that there
+were no obvious outliers in this data set, we concluded that
+there was insufficient evidence to justify excluding any
+participants from further analysis.
 
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../figures/fig_exp_2_mm_trunc_exgauss_compare.png}
+  \includegraphics[width=1\textwidth]{../figures/fig_exp_2_mm_trunc_exgauss_compare.png}
   \caption{
       Relative frequency histogram of the
-      trials-to-criterion observed across all three
-      conditions in Experiment 2. The black line represents
-      the best fit of the one-component Gaussian Mixture
-      Model, and the red line represents the best fit of the
-      two-component Gaussian Mixture Model.
+      trials-to-criterion collapsed across all participants
+      and conditions in Experiment 2. The left panel shows
+      the best fitting one-component ExGaussian
+      distribution, and the right panel shows the two
+      components of the best-fitting two component
+      ExGaussian mixture model.
 }
@@
 independent-samples t-test revealed no significant effect of
-condition ($t(32.0)=0.69$, $p = .50$, $\eta^2 = .01$).  The
-mean number of problems solved by each participant is shown
-in Figure \ref{fig_exp_2_t2c}B. An independent-samples
-t-test revealed no significant effect of condition ($t(32) =
--0.49$, $p = .63$, $\eta^2 = .01$).
+condition ($t(32.0)=1.37$, $p = .18$, $d = .47$).  The mean
+number of problems solved by each participant is shown in
+Figure \ref{fig_exp_2_t2c}B. An independent-samples t-test
+revealed no significant effect of condition ($t(32) =
+-1.99$, $p = .06$, $d = -.68$).
 
@@
 \subsection{Discussion of Experiments 1 and 2}
-Experiment 1 clearly showed that a short feedback delay of
-3.5 s slowed criterial learning. In contrast, increasing the
-ITI to this same value had no effect on learning. The task
-we used was one-dimensional category learning, but we
-isolated criterial learning by instructing participants
-about the optimal strategy. Specifically, we told them which
-stimulus dimension was relevant and that there was no
-trial-by-trial variability on the irrelevant dimension.
-Although these results strongly suggest that the feedback
-delay had its interfering effects on criterial learning,
-Experiment 2 was designed to confirm this inference. The
-goal here was to examine the effects of the same feedback
-delays and ITIs on performance in a one-dimensional
-category-learning task that did not require any criterial
-learning, but did require category-learning processes that
-are thought to mediate rule discovery (e.g., rule selection
-and switching). If the feedback delay effects observed in
-Experiment 1 were acting on some general category-learning
-skill then we should have seen the same interfering effects
-of feedback delay in Experiment 2. However, if the
-feedback-delay interference of Experiment 1 was operating
-selectively on criterial learning, then it should disappear
-in Experiment 2, since no criterial learning was required.
-Our results strongly supported this latter prediction.
-Therefore, Experiments 2 and 3 together strongly suggest
-that criterial learning is impaired by feedback delays and
-is relatively unaffected by the length of the ITI.
-
-Two prior studies investigated similar issues. First, Ell
-and colleagues reported that feedback delay and also a
-concurrent memory-scanning task each impaired rule-based
-category learning \parencite{ell2009critrial}. Based on
-these results, they hypothesized that criterial learning may
-be tied to working-memory capacity and therefore to explicit
-cognitive mechanisms. However, criterial learning was
-confounded with rule selection and task difficulty in their
-design.  Furthermore, \textcite{ell2009critrial} found that
-feedback delay only impaired learning when working memory
-demand was high -- that is, when participants had to learn
-more than one response criterion for optimal performance. In
-contrast, we found that feedback delay impairs learning even
-when working memory demands are trivial (participants never
-have to keep in mind more than one criterion).
-
-Second, \textcite{bohil2014implicit} reported that the
-effects of unequal base rates on criterion placement in
-rule-based category learning were diminished under delayed
-feedback and also under an observational training protocol
-that is also thought to selectively impair basal
-ganglia-dependent associative-learning mechanisms. These
-results are consistent with our findings. However, because
-they were obtained by manipulating base rates, it is
-difficult to conclude that the delay affected criterial
-learning, \emph{per se}. For example, consider a simple
-two-stage model in which the first stage learns the base
-rates and the second stage uses what the first stage learned
-to adjust the response criterion appropriately. The
+Experiment 1 showed that a short feedback delay of only 3.5
+s slowed criterial learning. In contrast, increasing the ITI
+from 500 ms to 3.5 s had no effect on learning. The task we
+used was one-dimensional category learning, but we isolated
+criterial learning by instructing participants about the
+optimal strategy. Specifically, we told them which stimulus
+dimension was relevant and that there was no trial-by-trial
+variability on the irrelevant dimension. Although these
+results strongly suggest that the feedback delay had its
+interfering effects on criterial learning, Experiment 2 was
+designed to complement this inference. The goal here was to
+examine the effects of the same feedback delays and ITIs on
+performance in a one-dimensional category-learning task that
+did not require any criterial learning, but did require
+category-learning processes that are thought to mediate rule
+discovery (e.g., rule selection and switching). If the
+feedback delay effects observed in Experiment 1 were acting
+on some general category-learning skill then we should have
+seen the same interfering effects of feedback delay in
+Experiment 2. However, if the feedback-delay interference of
+Experiment 1 was operating selectively on criterial
+learning, then it should disappear in Experiment 2, since no
+criterial learning was required.  Our results strongly
+supported this latter prediction.  Therefore, Experiments 1
+and 2 together strongly suggest that criterial learning is
+impaired by feedback delays and is relatively unaffected by
+the length of the ITI.
+
+Whereas the results of Experiment 1 seem clear, the
+Experiment 2 results are weaker since they suggest a null
+result. Although it impossible to rule out the possibility
+that Experiment 2 missed some small effect of feedback delay
+due to insufficient power, the null results found there are
+consistent with a wide variety of other evidence. In
+particular, a variety of other studies have reported that
+feedback delays of up to 10 s do not significantly slow
+one-dimensional rule-based learning \parencite{DunnEtAl2012,
+    ell2009critrial, MaddoxAshbyBohil2003, MaddoxIng2005,
+Worthyetal2013}. In addition, many studies have reported
+evidence that one-dimensional rule-based learning recruits
+working memory and executive attention, but not procedural
+learning \parencite[for reviews, see,
+e.g.,][]{AshbySmithRosedahl2020, AshbyValentin2017}. So the
+results of Experiment 2 are consistent with all of these
+studies. On the other hand, several of these feedback-delay
+studies did not use binary-valued stimulus dimensions, so
+some criterial learning was required. If so, then why did
+they all report no effect of feedback delay? It is important
+to note that criterial learning was not the focus of any of
+these studies, and as a result, in all of these previous
+studies the optimal criterion was always set exactly midway
+between the category prototypes, which makes criterial
+learning trivial. For example, under these conditions,
+criterial learning might not even require feedback. The
+unsupervised category-learning experiments reported by
+\textcite{ashby1999dominance} provide strong support for
+this because all of their rule-based participants learned
+the correct criterion (which was midway between the category
+means), even though the task was completely unsupervised.
+
+We know of only one previous study that reported impaired
+one-dimensional rule-based category learning when feedback
+was delayed \parencite{ell2009critrial}. However, in this
+study, a feedback delay only impaired performance in a
+condition with four contrasting categories that required
+learning three different response criteria. So this task was
+considerably more complex than any of the other
+feedback-delay studies, which all used two contrasting
+categories that required the learning of only one response
+criterion.
+
+One other study with similar goals to ours studied how
+manipulations of category base rates affect criterial
+learning in one-dimensional rule-based categorization.
+Specifically, \textcite{bohil2014implicit} reported that
+changes in criterion placement in response to changes in
+base rate were diminished when feedback was delayed (and
+also by observational training). These results are
+consistent with our findings. However, because they were
+obtained by manipulating base rates, it is difficult to
+conclude that the delay affected criterial learning,
+\emph{per se}. For example, consider a simple two-stage
+model in which the first stage learns the base rates and the
+second stage uses what the first stage learned to adjust the
+response criterion appropriately. The
 \textcite{bohil2014implicit} results are also consistent
@@
 
+\section{New Computational Models of Criterial Learning}
+The results of Experiments 1 and 2 seem to suggest that
+criterial learning is a form of procedural learning, which
+previous research suggests is of SR associations
+\parencite{AshbyWaldron1999, CasaleEtAl2012}. If so, then
+the criterion is simply the sensory, perceptual, or
+cognitive value that separates representations associated
+with contrasting responses. However, before accepting this
+conclusion, it is important to ask whether our results might
+also be compatible with an account in which the criterion is
+held is some form of working or short-term memory and
+updated trial by trial.
+
+To address this question, we developed computational models
+of criterial learning within two broad architectures and
+examined their sensitivity to ITI duration and feedback
+delay. The architectures differed in whether they assume
+that the response criterion is explicitly represented in
+memory. In the first architecture, the criterion is stored
+in memory, and responses are generated by comparing the
+current percept with this stored referent. The criterion is
+updated after error feedback using a gradient-descent rule.
+Models in this class can exhibit sensitivity to ITI and
+feedback delay in three distinct ways: (1) perceptual drift,
+where stimulus representations change over time, (2)
+criterial drift, where the stored criterion shifts over
+time, and (3) updating that is sensitive to feedback delays,
+where the magnitude of criterion updates decreases with
+longer feedback delays. In the second architecture, no
+explicit criterion is stored. Instead, reinforcement
+learning forms direct stimulus--response associations that
+drive performance.  Sensitivity to ITI and feedback delay
+arise from perceptual drift and/or delay-sensitive updating,
+but criterial drift is not possible since no criterion is
+represented.
+
+For each of these models, denote the time of stimulus
+presentation on the current trial by T$_\text{S}$, the
+response time by T$_\text{R}$, the time when feedback is
+displayed by T$_\text{F}$, and the time when the stimulus
+that begins the next trial is displayed by T$_{\text{S}^+}$.
+Then note that the feedback delay equals $\text{t}_\text{FD}
+= \text{T}_\text{F} - \text{T}_\text{R}$ and the duration of
+the ITI equals $\text{t}_\text{ITI} = \text{T}_{\text{S}^+}
+- \text{T}_\text{F}$.
+
+\subsection{Models that assume a stored criterion}
+These models assume that the observer constructs a response
+criterion, maintains it in memory, and makes
+decisions by comparing the percept to this criterion. They
+may further posit that both the stored criterion and the
+perceived stimulus value drift over time, with the extent of
+drift increasing both within a trial and across consecutive
+trials. In addition, they may assume that the learning rate
+for updating the criterion decreases as the feedback delay
+increases.
+
+Let $x_n(t)$ and $c_n(t)$ denote the values of the percept
+and the criterion, respectively, on trial $n$ at time $t$.
+Then the decision rule on trial $n$ is:
+\begin{equation}
+  \text{Respond } R_n =
+  \begin{cases}
+    \text{A}, & \text{if ~} x_n(\text{T}_\text{S}) \leq c_n(\text{T}_\text{S})  \\
+    \text{B}, & \text{if ~} x_n(\text{T}_\text{S}) > c_n(\text{T}_\text{S}).
+  \end{cases}
+  \label{eq:DR}
+\end{equation}
+Note that this equation indicates that the response is made
+immediately upon stimulus presentation. We therefore are not
+explicitly modeling the time-extended cognitive processes
+that govern evidence accumulation and the corresponding
+response time.
+
+If positive feedback is received, then the value of the
+criterion remains unchanged for the next trial (except for
+drift -- see equation \ref{eq:criterion}).  The rationale
+here is that if the response is correct, then the observer
+has effectively gained zero information about how their
+criterion should be modified.  If negative feedback is
+received, then the criterion is modified according to the
+standard model
+\parencite{SuttonBarto1998}:
+\begin{equation}
+c_n(\text{T}_\text{F}+ \Delta_t) = c_n(\text{T}_\text{F}) + \frac{\alpha}{t_{FD}} [x_n(\text{T}_\text{F}) - c_n(\text{T}_\text{F})],
+\label{RL}
+\end{equation}
+where $\alpha$ is a learning-rate parameter, and $\Delta_t$
+is the time it takes to complete the updating.  It is
+straightforward to show that the iterative equation \ref{RL}
+is equivalent to computing a weighted mean (weighted by
+recency) of the values of all percepts that occur on error
+trials \parencite[e.g.,][]{Ashby2017}. This updating rule
+will gradually converge on the optimal criterion value
+(i.e., the value of the criterion that maximizes accuracy).
+
+The term $\alpha/t_{FD}$ captures the notion that the
+magnitude of the update decreases as the feedback delay
+increases.  As a result, when feedback is delayed, the
+system becomes less responsive to errors, leading to slower
+learning compared to immediate feedback conditions. However,
+since this model is consistent with the idea that criterial
+learning relies on working memory -- and the available
+evidence suggests that logical reasoning and working memory
+are unaffected by feedback delays of several seconds
+\parencite[e.g., in one-dimensional rule-based category
+learning tasks;][]{ell2009critrial, MaddoxAshbyBohil2003,
+MaddoxIng2005} -- the learning process described in equation
+\ref{RL} likewise may be unaffected by feedback delay. We
+therefore also investigated a version of the model in which
+the term $\alpha/t_{FD}$ is replaced by $\alpha$, making the
+learning rate independent of feedback delay.
+
+This model further assumes that both the stimulus and the
+criterion representations may drift randomly throughout the
+duration of time they are maintained in working memory. The
+representation of the criterion must always be maintained in
+working memory, whereas drift in the stimulus representation
+affects performance up until the feedback is presented, but
+not afterwards (i.e., because the equation \ref{RL} updating
+rule depends on the value of the current stimulus
+representation). We modeled the drift in both the criterion
+and the percept by adding white noise to their initial
+values. Specifically, we assumed that for all
+$t>\text{T}_\text{S}$
+\begin{equation}
+  c_n(t) = c_n(\text{T}_\text{S}) + \eta_c \epsilon(t),
+  \label{eq:criterion}
+\end{equation}
+and
+\begin{equation}
+  x_n(t) = x_n(\text{T}_\text{S}) + \eta_x \epsilon(t),
+  \label{eq:percept}
+\end{equation}
+where $\epsilon(t)$ is white noise and $\eta_c$ and $\eta_x$
+are parameters that determine the amount of drift over time.
+Note that these parameters are permitted to be equal to
+zero, which is equivalent to assuming that the criterion
+and/or the percept do not drift over time at all.
+
+This model predicts that between the time of stimulus
+presentation and feedback, the response criterion will drift
+randomly. The amount of drift during this time will be
+normally distributed with mean zero and variance
+$\text{t}_\text{FD} \eta_c^2$. Similarly, the amount of
+criterial drift after the updating that follows feedback and
+the presentation of the stimulus on the next trial will also
+be normally distributed with mean zero, but now with
+variance $(\text{t}_\text{ITI}) \eta_c^2$. Predictions for
+the percept are similar. Specifically, the amount of drift
+between stimulus presentation and feedback will be normally
+distributed with mean zero and variance $\text{t}_\text{FD}
+\eta_x^2$.
+
+\subsection{Models that assume no criterion}
+A qualitatively different class of models, based on an
+actor-critic architecture \parencite{SuttonBarto1998},
+assume that learning is a process of associating each
+stimulus with a response via reinforcement learning. These
+models do not include an explicit representation of the
+response criterion. Rather the criterion is simply the
+mental representation that separates percepts associated
+with contrasting responses.
+
+In this class of models, the perceptual representation of
+each stimulus is modeled as a pattern of activation across
+25 sensory units that are each characterized by an
+overlapping tuning curve. Each unit is maximally excited by
+one specific stimulus, which we call the unit's preferred
+stimulus. Specifically, the activation of the $i^\text{th}$
+sensory unit on trial $n$ at time $t$ is given by
+\begin{equation}
+  A_i(n,t) = \text{exp} \left( \frac{-\left[d_{i,S_n(t)}\right]^2}{\sigma^2} \right)
+  \label{eq:DSL_activation}
+\end{equation}
+where $d_{i,S_n(t)}$ is the (Euclidean) distance between the
+preferred stimulus value of the $i^\text{th}$ sensory unit
+and the value of the stimulus that was presented on trial
+$n$ at time $t$, and $\sigma$ is a constant that increases
+with perceptual noise.
+
+Because this class of models does not store an explicit
+response criterion, it cannot incorporate criterial drift.
+It can, however, incorporate perceptual drift. In this case,
+the value of the current stimulus used to compute sensory
+unit activation (Equation~\ref{eq:DSL_activation}) is
+assumed to drift over time according to
+\begin{equation}
+  S_n(t) = S_n(\text{T}_\text{S}) + \eta_S \epsilon(t),
+  \label{eq:stimulus_drift_RL}
+\end{equation}
+where $\eta_S$ scales the magnitude of the drift and
+$\epsilon(t)$ is a noise process.
+
+The model includes two decision or actor units -- one
+associated with each of the two possible responses.
+Initially, each sensory unit is connected to both actor
+units with some random connection strength. Let
+$\omega_{iJ}(n)$ denote the strength of the connection
+between sensory unit $i$ and actor unit $J$ (for $J$ = A or
+B) on trial $n$. Then the activation in actor unit $J$ on
+trial $n$ at time $t$, denoted by $V_J(n,t)$, equals
+\begin{equation}
+  V_J(n,t) = \sum_{i} \omega_{iJ}(n) A_i(n,t).
+\end{equation}
+A response is selected by comparing the activations in
+competing actor units immediately after stimulus
+presentation, that is, at time $t=\text{T}_\text{S}$,  using
+the following the decision rule:
+\begin{equation}
+ \text{Respond } R_n =
+  \begin{cases}
+    A, & \text{if $V_A(n,\text{T}_\text{S}) > V_B(n,\text{T}_\text{S})$}\\
+    B, & \text{if $V_A(n,\text{T}_\text{S}) \leq V_B(n,\text{T}_\text{S}).$}
+  \end{cases}
+\end{equation}
+
+The connection strengths $\omega_{iA}(n)$ and
+$\omega_{iB}(n)$ are updated after feedback is received on
+each trial according to standard reinforcement learning
+rules \parencite{SuttonBarto1998}. If the feedback is
+positive, then for J = A or B:
+\begin{equation}
+  \omega_{i\text{J}}(n) = \omega_{i\text{J}}(n-1) + \frac{\alpha_{actor}}{t_\text{FD}} A_i(n,\text{T}_\text{F}) V_\text{J}(n,\text{T}_\text{F}) \delta(n-1) [1-\omega_{i\text{J}}(n-1)],
+  \label{eq:W+}
+\end{equation}
+where $\alpha_{actor}$ is the learning-rate of the actor,
+and $\delta(n-1)$ is the reward prediction error on trial
+$n-1$. The last term prevents the weight from exceeding 1.
+Note that the increase in synaptic strength is scaled by the
+product of the pre- and postsynaptic activation values. Also
+note that, as in equation~\ref{RL}, the learning rate is
+scaled by the inverse of the feedback delay. If the feedback
+is negative, then 
+\begin{equation}
+  \omega_{i\text{J}}(n) = \omega_{i\text{J}}(n-1) + \frac{\alpha_{actor}}{t_\text{FD}} A_i(n,\text{T}_\text{F}) V_\text{J}(n,\text{T}_\text{F}) \delta(n-1) \omega_{i\text{J}}(n-1).
+  \label{eq:W-}
+\end{equation}
+Now the last term prevents the weight from falling below 0.
+Note that on error trials, $\delta(n-1)$ will be negative,
+so equation \ref{eq:W-} decreases the synaptic weight.
+
+The reward prediction error on trial $n-1$ is defined as the
+value of the obtained reward, denoted by $R(n-1)$ minus the
+value of the predicted reward, denoted by $P(n-1)$:
+\begin{equation}
+  \delta(n-1) = R(n-1) - P(n-1).
+\end{equation}
+The predicted reward on trial $n$ is determined by the
+critic via
+\begin{equation}
+  P(n) = P(n-1) + \frac{\alpha_{critic}}{t_\text{FD}} \delta(n-1),
+  \label{eq:Pcritic}
+\end{equation}
+where $\alpha_{critic}$ is the learning rate of the critic.
+Note that this learning rate is again scaled by the inverse
+of the feedback delay.
+
+Much evidence suggests that this type of stimulus-response
+learning is mediated largely within the striatum, and is
+facilitated by a dopamine-mediated reinforcement learning
+signal that is time dependent
+\parencite[e.g.,][]{ValentinMaddoxAshby2014}. Specifically,
+the dopamine signal generated by positive feedback appears
+to peak at around 500 ms after feedback and then decay back
+to baseline levels within 2 or 3 s
+\parencite{YagishitaEtAl2014}. As a result, synaptic
+plasticity at cortical-striatal synapses is attenuated with
+increasing feedback delays \parencite{YagishitaEtAl2014}.
+The scaling of $\alpha_{actor}$ and $\alpha_{critic}$ by
+$1/{t_\text{FD}}$ is one way to model this phenomenon.
+However, it is possible that stimulus-response association
+learning is mediated by other brain systems that are less
+sensitive to feedback delay. For this reason, we also
+investigated a version of the model in which the updating
+equations are independent of feedback delay. In this
+version, the term $\alpha_{actor}/t_\text{FD}$ is replaced
+by $\alpha_{actor}$ in equations \ref{eq:W+} and \ref{eq:W-}
+and the term $\alpha_{critic}/t_\text{FD}$ is replaced by
+$\alpha_{critic}$ in equation \ref{eq:Pcritic}.
+
+\subsection{Simulation Results}
+We investigated how changing the duration of the feedback
+delay and the ITI affect criterial learning for each class
+of model. More specifically, we simulated performance of
+each model in each of the three conditions of the Experiment
+1 categorization task. Each simulation continued for 200
+trials or until the model responded correctly for 12 trials
+in a row. For each set of parameter values, we simulated the
+model 100 times and averaged the results, yielding one
+observed performance metric measured in mean
+trials-to-criterion for each condition.
+
+Our approach was to generate predictions for each of the
+three experimental conditions across a wide range of
+parameter values. We then used parameter-space partitioning
+(PSP) to evaluate the performance of each model
+\parencite{pitt2006global}. Traditional computational
+modeling identifies parameter values that allow the model to
+provide the best possible fit to the observed data. This
+traditional approach therefore treats the single observed
+data set being fit as a gold standard. Our goal instead was
+to investigate the range of possible predictions of each
+model. We were interested in questions such as: can a model
+predict any possible outcome or is it constrained to always
+predict, for example, that feedback delays impair learning?
+PSP was designed to answer such questions.
+
+Classical PSP is defined for non-stochastic models and is
+often implemented by exploring the parameter space via a
+Markov chain Monte Carlo algorithm. In contrast, we used
+brute force search over a grid of plausible parameter
+values. Because our models are stochastic, at each sampled
+set of parameter values we generated multiple simulations
+and classified the results into one of a small set of
+qualitative outcome patterns. PSP calculates the proportion
+of the parameter space where the model makes specific
+predetermined qualitative predictions. We focused on the
+following four such predictions. 1) Delay impaired:
+Performance is worst in the Feedback-Delay condition. 2)
+Long ITI impaired: Performance is worst in the Long-ITI
+condition. 3) Short ITI impaired: Performance is worst in
+the Control condition. 4) Other: Two or more conditions are
+tied for worst performance. For each model, the PSP analysis
+quantified the proportions of the explored parameter space
+where the model made each of these qualitative predictions.
+
+\subsubsection{Models that assume a stored criterion}
+We investigated predictions of these models across a wide
+range of values for the parameters $\eta_{x}$, $\eta_{c}$,
+and $\alpha$.  Specifically, in the case of $\alpha$,  we
+stepped through every value in the interval $[0, .2]$, with
+a step size of .01.  In the case of both $\eta_{x}$, and
+$\eta_{c}$, we searched over the interval $[0, 5]$, with a
+step size of $.1$.  We also explored the binary case of
+whether or not the update rate $\alpha$ was scaled by the
+feedback delay.
+
+Figure~\ref{fig:criterion_results} shows the results. Panel
+A lists the proportions of parameter space over which the
+version of the model that is sensitive to feedback delay
+predicts each of the four qualitatively different outcomes.
+Note that this model predicts that the poorest performance
+will be in the Feedback Delay condition for all values of
+its parameters. At first glance, this result seems
+surprising because it means that there are no parameter
+values that make the model more sensitive to ITI than to
+feedback delay. This makes sense though because increasing
+the ITI has no effect on the mean criterion value -- the
+only effect is to increase the criterion value's variance.
+In contrast, increasing the feedback delay affects the mean
+criterion value because it slows the criterion updating
+process from converging on the optimal value.
+
+\begin{figure} 
+  \centering
+  \includegraphics[width=1\textwidth]{../figures/model_new_class_II_yes_stored_criterion.pdf}
+  \caption{Simulation results for models that assume a stored criterion.  
+      \textbf{A:} The proportions of parameter space where
+      each of the four qualitative outcomes is predicted by
+      the versions of the model that assume criterion
+      updating is sensitive to feedback delay. 
+      \textbf{B:} Boxplot showing the parameter ranges that
+      produced the panel A results. All parameter values
+      were normalized by the largest value in the search
+      range, so the ordinate ranges from zero to one for all
+      parameters.
+      % 
+      \textbf{C:} The proportions of parameter space where
+      the models that assume criterion updating is
+      insensitive to feedback delay predict each of the four
+      qualitative outcomes. 
+      \textbf{D:} Boxplot showing the parameter ranges that
+      produced the panel C results.     
+      %
+      \textit{Note:} In all panels, color indicates the PSP
+      pattern.
+}
+    \label{fig:criterion_results}
+\end{figure}
+
+Panel C of Figure \ref{fig:criterion_results} shows the same
+results as in panel A, except for the version of the model
+in which the criterion updating process is insensitive to
+feedback delay. Note that this version of the model predicts
+that performance must be worst in the Long-ITI condition --
+again for all values of its parameters. This result also
+shows that no amount of perceptual or criterial drift allows
+this version of the model to account for the Experiment 1
+results.
+
+\subsubsection{Models that assume no criterion}
+We investigated predictions of the SR-learning model by
+varying four parameters; namely, the perceptual-noise
+variance $\sigma^2$, the stimulus drift parameter $\eta_S$,
+and the actor and critic learning rates (i.e.,
+$\alpha_{actor}$ and $\alpha_{critic}$, respectively). In
+the case of $\sigma$, we searched over the interval $[1,
+10]$, with a step size of $3$. For $\eta_S$, we searched
+over the interval $[0, 1]$, with a step size of $.1$. For
+both $\alpha_{actor}$ and $\alpha_{critic}$, we stepped
+through every value in the interval $[0, .2]$, with a step
+size of $.02$. We constrained our search over these
+parameters to this interval because reinforcement-learning
+models are prone to instability at very high learning rates
+\parencite{SuttonBarto1998}. As evidence of this, at higher
+learning rates, the model failed to learn with any
+consistency -- that is, in most cases, it failed to reach
+the learning criterion (12 correct responses in a row)
+within the allowable 200 trials.   We also explored the
+binary case of whether or not the update rates
+$\alpha_{actor}$ and $\alpha_{critic}$ were scaled by the
+feedback delay. As with the other models, all scores were
+normalized by the largest observed value.
+
+The results are described in Figure \ref{fig:RL_results}.
+Panel A lists the proportions of parameter space over which
+the feedback-delay sensitive version of the model predicts
+each qualitatively different outcome. Note that across the
+vast majority of its parameter space, this model predicts
+that an increase in feedback delay will impair performance
+more than an increase in ITI.
+
+\begin{figure} 
+  \centering
+  \includegraphics[width=1\textwidth]{../figures/model_new_class_I_no_stored_criterion.pdf}
+  \caption{ Simulation results for models that assumed
+      stimulus-response learning and therefore no stored
+      criterion.  
+      \textbf{A:} The proportions of parameter space where
+      each of the four qualitative outcomes is predicted by
+      the versions of the model that assume synaptic
+      updating is sensitive to feedback delay.
+      \textbf{B:} Boxplot showing the parameter ranges that
+      produced the panel A results.     All parameter values
+      were normalized by the largest value in the search
+      range, so the ordinate ranges from zero to one for all
+      parameters.
+      % 
+      \textbf{C:} The proportions of parameter space where
+      the models that assume synaptic updating is
+      insensitive to feedback delay predict each of the four
+      qualitative outcomes. 
+      \textbf{D:} Boxplot showing the parameter ranges that
+      produced the panel C results.     
+      %
+      \textit{Note:} In all panels, color indicates the PSP
+      pattern.
+}
+  \label{fig:RL_results}
+\end{figure}
+
+Figure \ref{fig:RL_results}C shows the proportions of
+parameter space over which the models that assume synaptic
+updating is insensitive to feedback delay predict each of
+the four qualitative outcomes. Surprisingly, this version of
+the model also almost always predicts the poorest
+performance in the Feedback-Delay condition. What is going
+on here? First, note that even if
+$\alpha_{actor}/t_\text{FD}$ and
+$\alpha_{critic}/t_\text{FD}$ in equations \ref{eq:W+},
+\ref{eq:W-}, and \ref{eq:Pcritic} are replaced by
+$\alpha_{actor}$ and $\alpha_{critic}$, respectively, then
+equations \ref{eq:W+} and \ref{eq:W-} are still sensitive to
+feedback delay because the pre- and postsynaptic activations
+in equations \ref{eq:W+} and \ref{eq:W-} are computed at the
+time of feedback -- that is, at time T$_\text{F}$. So the
+longer that the feedback is delayed, the more time there is
+for the stimulus representation to drift from its true
+value. As a result, this model predicts that delayed
+feedback will impair learning, even though the amount of
+updating on each trial does not depend on feedback delay.
+Second, note that all versions of this model are generally
+insensitive to the length of the ITI. The stored-criterion
+models are adversely affected by a long ITI because the
+criterion drifts during this time and large drift means that
+on the next trial a decision will be made with a suboptimal
+value of the criterion. But there is no criterion to drift
+in the SR-learning models. The stimulus will continue to
+drift during the ITI in these models, but with no
+consequence, because after the synaptic updating that occurs
+with presentation of the feedback, the stimulus
+representation on the current trial is no longer needed. So
+all versions of the SR-learning model are more sensitive to
+feedback delay than to the duration of the ITI. In support
+of this interpretation, note from panel D of Figure
+\ref{fig:RL_results} that the few versions of the model in
+which synaptic updating is insensitive to feedback delay
+that predict poorest performance in the Long ITI or Control
+conditions all include almost no stimulus drift (i.e.,
+$\eta_\text{S}$ is small), which eliminates the one
+remaining component of the model that makes it sensitive to
+feedback delay.
+
+\subsubsection{Summary of modeling results}
+We examined two classes of criterial-learning models that
+make different predictions about the effects of feedback
+delay and ITI. The first class assumes that the criterion is
+stored in memory and compared against incoming stimuli to
+generate responses. These models predict poorest performance
+in the Feedback-Delay condition only when the update rule is
+explicitly delay-sensitive. Otherwise, they make the strong
+prediction that performance must be worst in the Long-ITI
+condition. The second class of models assumes no stored
+criterion, with responding driven instead by reinforcement
+learning of SR associations. Almost all versions of this
+general model make the strong a priori prediction that
+performance must be worst in the Feedback-Delay condition --
+even those versions that predict no effect of feedback delay
+on the strength of synaptic updating.
+
 \section{General Discussion}
-We developed novel models of criterial learning that were
-designed to explore how feedback delay and ITI duration
-affect criterial learning.  One class of model assumed that
-decisions are made by comparing the current percept to a
-stored referent, or criterion, and that the remembered value
-of this criterion is updated following error feedback via a
-gradient-descent learning rule.  This class of model
-generates sensitivity to feedback delay and ITI duration
-through (1) directly assuming that the update rule is
-inversely scaled by the feedback delay, (2) assuming that
-the representation of the percept drifts over time, or (3)
-assuming the representation of the criterion drifts over
-time.  Simulations of these models in a task that was
-structurally identical to Experiment 1 revealed that this
-class of model typically predicts that feedback delay in
-more impairing than a long ITI only when the update rule is
-explicitly delay-sensitive.  Otherwise, it typically
-predicts that long ITIs are more impairing than feedback
-delays.
-
-The second model class, reinforcement-learning models,
-assumes that criterial learning arises through the gradual
-formation of stimulus-response associations, rather than via
-the construction of a response criterion.  These models can
-generate sensitivity to feedback delay and ITI duration
-through the same two mechanisms as the first class of
-models. However, since these models do not explicitly
-represent a criterion, they do not assume that the
-representation of the criterion drifts over time.   This
-class of model  typically predicts that feedback delay is
-more impairing than a long ITI across wide ranges of
-parameter values.
-
-The experiments were designed to test these predictions. Our
-results strongly suggested that human criterial learning is
-sensitive to feedback delay but not to the ITI duration.
-These results shed important constraints on the new models
-of criterial learning developed here. In particular, the
-reinforcement-learning models are consistent with the
-observed behavioural data across wide ranges of their
-parameter spaces, regardless of whether or not the update
-rule is delay-sensitive.  However, the models that assume a
-stored criterion are also consistent with the observed
-behavioral data, but only when the update rule is
-delay-sensitive.
-
-One natural neural underpinning for models that assume a
-stored criterion is that the criterion is maintained in
-working memory. However, since it is widely accepted that
-working memory is largely insensitive to both feedback delay
-and ITI duration, this interpretation is difficult to
-reconcile with our results suggesting a delay-sensitive
-update rule is required for this class of model. An
-alternative possibility for where the criterion might be
-encoded in a more stable memory system.  One appealing
-candidate for this function is the cerebellum, where
-synaptic plasticity has been shown to follow a gradient
-descent learning rule and also to be sensitive to feedback
-timing \parencite{brudner2016delayed, held_adaptation_1966,
-honda_adaptation_2012, kitazawa_effects_1995,
-kitazawa_prism_2002}.   Finally, the reinforcement-learning
-model draws on the idea that criterial learning could emerge
-through stimulus-response association learning driven by
-dopamine-dependent synaptic plasticity in the striatum. This
-process aligns with established theories of category
-learning that posit a key role to this learning process in
-the basal ganglia \cite{AshbyCOVIS1998}.
-
-Our results do not strongly favor one model over another.
-The critical disagreement between these accounts is whether
-the criterion has any real psychological meaning. One class
-assumes the criterion is stored in memory and compared
-against incoming stimuli to generate responses.  The
-reinforcement-learning model, in contrast, makes very
-different psychological assumptions. This model assumes that
-behaviors described as ``criterial learning'' are actually
-mediated by the learning of stimulus-response associations.
-According to this account, the criterion has no internal
-representation and therefore no psychological meaning.
-Testing between these two very different accounts of
-criterial learning should be a focus of future research.
-
-We are not aware of any data that directly addresses the
-question of whether a response criterion is a fundamental
-psychological construct, or whether it is unnecessary. Even
-so, there are some results in the category-learning
-literature that support the interpretation of the
-reinforcement-learning model. In information-integration
-(II) category-learning tasks, the optimal strategy is
-similarity-based, and difficult or impossible to describe
-verbally \parencite[e.g.,][]{AshbyValentin2018}. When the
-stimuli vary on two dimensions, the stimuli from contrasting
-categories can be partitioned by a decision bound that is
-conceptually similar to a response criterion. In both cases,
-all stimuli on one side are associated with one response,
-and all stimuli on the other side are associated with the
-contrasting response. Furthermore, in agreement with
-Experiment 1, II category learning is impaired by short
-feedback delays \parencite{MaddoxAshbyBohil2003,
-MaddoxIng2005}. The analogous question in II category
-learning is whether the decision bound is learned directly
-or whether it is simply the set of points that divide the
-perceptual space into contrasting response regions.  In
-fact, the evidence strongly supports this latter
-interpretation \parencite{AshbyWaldron1999, CasaleEtAl2012}.
-For example, if the decision bound is learned, then it
-should be possible to apply this bound to novel stimuli.
-With rule-based categories, this is easy for participants,
-but with II categories there is no evidence that the
-response strategy that participants learn can be generalized
-to novel stimuli \parencite{CasaleEtAl2012} -- a result that
-strongly supports the hypothesis that the decision bound has
-no psychological meaning.
-
-Our results clearly demonstrate that criterial learning is
-impaired by delayed feedback, and not by extending the
-intertrial interval. These results are consistent with the
-hypothesis that criterial learning is a form of
-basal-ganglia mediated associative learning, and are
-inconsistent with hypotheses that criterial learning is a
-working-memory-based process. Thus, our results provide a
-critical constraint on future models of rule-based
-classification and decision making, and possibly also on
-more general accounts of criterion setting, such as in
-signal detection theory.
-
-Previous studies have failed to find that feedback delays
-impair rule-based category learning, and on the face of it,
-this seems to contradict our finding that feedback delays
-impair criterial learning. However, all earlier RB studies
-that looked for effects of a feedback delay, either used
-binary-valued stimulus dimension and so no criterial
-learning was needed, or else set the response criterion
-exactly midway between the category prototypes, which makes
-criterial learning trivial. For example, under these
-conditions, criterial learning might not even require
-feedback. The unsupervised category-learning experiments
-reported by \textcite{ashby1999dominance} provide strong
-support for this because all of their rule-based
-participants learned the correct criterion (which was midway
-between the category means), even though the task was
-completely unsupervised. Moreover, all previous studies
-failed to isolate criterial learning, so even if there was
-an effect of feedback delay on criterial learning, it could
-have been masked by larger effects caused by other
-rule-learning processes.
+The notion of a response criterion is ubiquitous in
+psychological models of decision making. Despite its
+theoretical importance, relatively little is known about its
+perceptual and cognitive basis. This article focused on two
+strikingly different interpretations of the criterion. The
+first, which seems to be assumed implicitly by most
+decision-making models -- including for example, signal
+detection theory -- is that decisions are made by comparing
+the stimulus value to a stored value of the criterion, and
+that this stored value is updated trial-by-trial after the
+feedback is presented. A second interpretation, which is
+motivated by results from the procedural-learning
+literature, is that learning instead is a process of
+associating responses with stimuli. According to this
+account, there is no response criterion -- at least not one
+with any mental representation. The value that the first
+class of models would refer to as the response criterion is
+simply the hypothetical value that separates stimuli
+associated with contrasting responses.
+
+As a first step in testing between these two very different
+interpretations, we explored, both empirically and
+theoretically, how feedback delays and increases in the ITI
+affect criterial learning in a one-dimensional
+category-learning task. Our empirical results strongly
+suggested that human criterial learning is sensitive to
+feedback delay but not to the ITI duration. To investigate
+the theoretical implications of this result, we examined
+predictions of two qualitatively different types of
+computational model. This analysis showed that SR-learning
+models almost always predict that feedback delays should
+impair learning more than long ITIs, even versions of the
+model in which the amount of synaptic updating is
+insensitive to feedback delay. In contrast, a broad class of
+models that assumed the criterion is learned and updated
+trial-by-trial predict that increasing the ITI should have
+greater effect than delaying feedback, except for the subset
+of these models that explicitly assume the criterial
+updating process is sensitive to feedback delay.
+
+In summary, our results suggest that if a criterion is
+learned and stored in memory, then its trial-by-trial
+updating must be sensitive to feedback delay. How likely is
+this model? If a criterion is learned and stored in memory,
+then one obvious possibility is that the storage is in
+working memory. After all, in the present experiments at
+least, the current value of the criterion would need to be
+accessed and updated every few seconds -- properties that
+are commonly attributed to working memory
+\parencite[e.g.,][]{Baddeley2010, Oberauer2002}. The problem
+with this account is that it is widely accepted that working
+memory is largely insensitive to feedback delay. For
+example, rule-based category-learning tasks that are known
+to recruit working memory and executive attention are
+generally unaffected by feedback delays of up to 10 s
+\parencite{DunnEtAl2012, ell2009critrial,
+MaddoxAshbyBohil2003, MaddoxIng2005,  Worthyetal2013}.
+Whereas working memory is generally attributed to circuits
+centered in prefrontal cortex
+\parencite[e.g.,][]{AshbyEtAl2005, Funahashi2017}, an
+alternative possibility is that the criterion is stored in
+the cerebellum. For example, the cerebellum has been
+associated with working-memory tasks
+\parencite[e.g.,][]{AshidaEtAl2019}, and synaptic plasticity
+there follows a gradient descent learning rule that is
+sensitive to feedback timing \parencite{brudner2016delayed,
+held_adaptation_1966, honda_adaptation_2012,
+kitazawa_effects_1995, kitazawa_prism_2002}. Although our
+results can not rule out this possibility, there is
+virtually no previous literature linking criterial learning
+to the cerebellum, so much more empirical and theoretical
+work would be needed to establish this as a viable model.
+
+Whereas the stored-criterion model struggles to  account for
+our results, the SR-learning account  is simple and
+straightforward. In particular, almost all versions of this
+model predict our results -- even those versions that assume
+the strength of synaptic updating is unaffected by feedback
+delays. Furthermore, there is a well-accepted neural account
+of this model. First, there is abundant evidence that the
+learning of arbitrary SR associations is mediated primarily
+within the striatum \parencite[e.g.,][]{Horvitz2009,
+PackardMcGaugh1992}. Second, there is also abundant evidence
+that striatal-mediated learning is impaired with feedback
+delays as short as 2.5 or 3 s
+\parencite{MaddoxAshbyBohil2003, MaddoxIng2005,
+YagishitaEtAl2014}.
+
+We are not aware of any other attempts to address directly
+the question of whether or not a response criterion is a
+fundamental psychological construct. Even so, there are some
+results in the category-learning literature that support a
+mixed view. The multivariate generalization of a response
+criterion is a decision bound that separates
+multidimensional stimuli belonging to contrasting categories
+\parencite[e.g.,][]{AshbyGott1988}. The multivariate
+analogue of the stored-criterion hypothesis is that the
+decision bound is learned and stored trial-by-trial in some
+form of memory and that categorization decisions are made by
+noting which side of the decision bound the current stimulus
+is on. A second possibility though is that category-learning
+in these tasks is of SR associations, in which case the
+decision bound has no psychological meaning and instead is
+just the set of percepts that divide the perceptual space
+into regions associated with contrasting responses. Although
+more research is needed on this question, the current
+evidence strongly suggests that both hypotheses are correct,
+but that they apply to different types of categorization
+tasks.
+
+With rule-based tasks in which the optimal strategy is a
+one-dimensional rule (like the tasks used here), the
+evidence strongly favors the decision-bound-is-learned
+hypothesis, whereas with information-integration tasks in
+which the optimal strategy is impossible to describe
+verbally and therefore requires some sort of
+similarity-based response strategy, the evidence strongly
+favors the SR-learning hypothesis
+\parencite{AshbyWaldron1999, CasaleEtAl2012}. For example,
+if a decision bound is learned, it should be possible to use
+this bound to respond to novel stimuli that fall in some
+nearby, but untrained region of perceptual space. In
+contrast, SR learning builds associations between specific
+stimuli and specific responses and would not generalize to
+novel stimuli -- at least, not if the novel stimuli are
+perceptually distinct. With rule-based categories,
+generalization to novel stimuli is nearly perfect -- that
+is, responses to novel stimuli are almost perfectly
+predicted by the one-dimensional decision bound that best
+describes the participant's training performance
+\parencite{CasaleEtAl2012}. However, with
+information-integration categories, there is no evidence of
+any generalization. Instead, performance on novel stimuli is
+at chance \parencite{CasaleEtAl2012}.
+
+Rule-based category learning depends on working memory and
+executive attention and is relatively unaffected by feedback
+delays as long as 10 s, whereas information-integration
+category learning recruits procedural learning and is
+impaired by feedback delays as short as 2.5 s \parencite[for
+a review of the scores of studies supporting this
+hypothesis, see, e.g.,][]{Ashby2025}. Our results suggest
+that criterial learning shares the same sensitivity to
+feedback delay as information-integration category learning.
+Therefore, our results support the hypothesis that criterial
+learning, like information-integration category learning, is
+a form of procedural learning -- that is, criterial learning
+is a form of SR associative learning that is mediated
+primarily in the basal ganglia. Similarly, our results are
+inconsistent with the hypothesis that criterial learning
+depends in any significant way on working memory and/or
+executive attention.
 

### 2025-10-17 code/inspect_model_new.py

@@ -43,8 +43,8 @@ def inspect_class_I():
     print(d1['psp_class'].value_counts())
 
     # Remove 'Other' class for plotting
-    d1 = d1[d1['psp_class'] != 'Other']
-    d1['psp_class'] = d1['psp_class'].cat.remove_unused_categories()
+    # d1 = d1[d1['psp_class'] != 'Other']
+    # d1['psp_class'] = d1['psp_class'].cat.remove_unused_categories()
 
     # report the number of models in each class
     print(d1['psp_class'].value_counts())
@@ -55,27 +55,27 @@ def inspect_class_I():
     print("Unique values of alpha_critic:", d1['alpha_critic'].unique())
     print("Unique values of eta_perceptual_drift:", d1['eta_perceptual_drift'].unique())
 
-    # NOTE: Inspect class I
-    fig, ax = plt.subplots(figsize=(5, 4))
-    sns.histplot(data=d1,
-                 x='t2c_delay',
-                 stat='density',
-                 common_norm=False,
-                 bins=30,
-                 ax=ax)
-    sns.histplot(data=d1,
-                 x='t2c_liti',
-                 stat='density',
-                 common_norm=False,
-                 bins=30,
-                 ax=ax)
-    sns.histplot(data=d1,
-                 x='t2c_siti',
-                 stat='density',
-                 common_norm=False,
-                 bins=30,
-                 ax=ax)
-    plt.show()
+#    # NOTE: Inspect class I
+#    fig, ax = plt.subplots(figsize=(5, 4))
+#    sns.histplot(data=d1,
+#                 x='t2c_delay',
+#                 stat='density',
+#                 common_norm=False,
+#                 bins=30,
+#                 ax=ax)
+#    sns.histplot(data=d1,
+#                 x='t2c_liti',
+#                 stat='density',
+#                 common_norm=False,
+#                 bins=30,
+#                 ax=ax)
+#    sns.histplot(data=d1,
+#                 x='t2c_siti',
+#                 stat='density',
+#                 common_norm=False,
+#                 bins=30,
+#                 ax=ax)
+#    plt.show()
 
     # NOTE: Figure style from first submission
     sns.set_palette("rocket", 4)
@@ -131,7 +131,7 @@ def inspect_class_I():
     ax.set_xticks(np.arange(0, 4, 1))
[…]

### 2025-10-17 code/inspect_results_behave.py

@@ -1,6 +1,6 @@
 from util_func import *
 from imports import *
 
-# report_exp_1()
-report_exp_2()
+report_exp_1()
+# report_exp_2()
 # report_exp_3()

### 2025-10-17 code/util_exgauss.py

@@ -422,18 +422,25 @@ def plot_exg_mixture(model, X, bins=30, title=None, ax=None):
             density=True,
             alpha=0.25,
             edgecolor="none",
-            label="Data (density)")
+            color="gray",
+            label="Human data")
 
     # lines
-    colors = ["tab:blue", "tab:orange", "tab:green", "tab:red"]
+
+    # use rocket colors 
+    from matplotlib.cm import get_cmap
+    cmap = get_cmap("rocket")
+    n_colors = 6
+    colors = [cmap(0.2), cmap(0.5)]
+
     for k, ck in enumerate(comps):
         ax.plot(xgrid,
                 ck,
                 ls="--",
                 lw=2,
                 color=colors[k % len(colors)],
-                label=f"Component {k+1} (weighted)")
-    ax.plot(xgrid, pdf_total, lw=2, label="Mixture total")
+                label=f"Component {k+1}")
+    # ax.plot(xgrid, pdf_total, lw=2, label="Mixture total")
 
     # rug colored by hard labels (ordered by mu for consistency)
     order = np.argsort(mu)
@@ -443,15 +450,14 @@ def plot_exg_mixture(model, X, bins=30, title=None, ax=None):
         xk = X[labels == k]
         if xk.size:
             ax.scatter(xk,
-                       np.full_like(xk, -0.002 - 0.001 * k),
+                       np.full_like(xk, -0.002 - 0.001 * 1),
                        s=10,
                        alpha=0.9,
                        color=colors[k % len(colors)],
-                       label=(f"Rug (comp {k+1})"
-                              if k == 0 else None))  # avoid dup legend
+                       label="")
 
-    ax.set_xlabel("x")
-    ax.set_ylabel("Density")
+    ax.set_xlabel("Trials to criterion", fontsize=12)
+    ax.set_ylabel("Density", fontsize=12)
     if title:
         ax.set_title(title)
     yl = ax.get_ylim()
@@ -471,7 +477,7 @@ def plot_exg_compare(exg_single, exg_mix, X, bins=30):
     plot_exg_mixture(exg_mix,
                      X,
                      bins=bins,
-                     title=f"Mixture ({exg_mix.n_components} comps)",
+                     title="Mixture Model",
                      ax=axes[1])
     axes[0].set_ylabel("Density")
     plt.tight_layout()

### 2025-10-17 code/util_func.py

@@ -170,16 +170,6 @@ def load_data_exp_2():
 
     d = pd.concat([pd.read_csv(f) for f in f_names])
 
-    f_names = [
-        os.path.join('../data/bin_data_v1', f)
-        for f in os.listdir('../data/bin_data_v1') if f.endswith('.csv')
-    ]
-
-    d1 = pd.concat([pd.read_csv(f) for f in f_names])
-    d1['sub_num'] = d1['sub_num'] + 100
-
-    d = pd.concat([d, d1], ignore_index=True)
-
     d.rename(
         {
             'sub_num': 'sub',
@@ -831,20 +821,20 @@ def report_exp_1():
     print("ANOVA on trials to criterion:")
     print(res)
 
-    res = pg.pairwise_tests(data=ddd, dv='t2c', between='cnd', parametric=True)
+    res = pg.pairwise_tests(data=ddd, dv='t2c', between='cnd', parametric=True, effsize='cohen')
     print()
     print("Pairwise comparisons on trials to criterion:")
-    print(res[['A', 'B', 'T', 'dof', 'p-unc']])
+    print(res[['A', 'B', 'T', 'dof', 'p-unc', 'cohen']])
 
     res = pg.anova(data=ddd, dv='nps', between='cnd', ss_type=3, effsize='np2')
     print()
     print("ANOVA on number of problems solved:")
     print(res)
 
-    res = pg.pairwise_tests(data=ddd, dv='nps', between='cnd', parametric=True)
+    res = pg.pairwise_tests(data=ddd, dv='nps', between='cnd', parametric=True, effsize='cohen')
     print()
     print("Pairwise comparisons on number of problems solved:")
-    print(res[['A', 'B', 'T', 'dof', 'p-unc']])
+    print(res[['A', 'B', 'T', 'dof', 'p-unc', 'cohen']])
 
 
 def report_exp_2():
@@ -875,22 +865,22 @@ def report_exp_2():
                              n_init=16).fit(X)
 
     # Bootstrap LRT p-value (use a small n_boot while iterating; increase at the end)
-    # D_obs, p_boot, D_boot, exg2_refit = bootstrap_lrt_pvalue(
-    #     X,
-    #     null_model=exg1,
-    #     alt_components=2,
-    #     n_boot=100,
-    #     n_init_null_boot=2,
-    #     n_init_alt_boot=4,
[…]

### 2025-10-17 code/inspect_model_new.py
basic cleanup

@@ -1,577 +1,0 @@
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-import seaborn as sns
-
-def inspect_class_I():
-
-    print()
-    print("Inspecting Class I models...")
-
-    # NOTE: Class I setup
-    d1 = pd.read_csv('../output/param_record_class_I_multi.csv')
-
-    d1 = d1.groupby([
-        'sigma_perceptual_noise', 'alpha_actor', 'alpha_critic',
-        'eta_perceptual_drift', 'delay_sensitive_update'
-    ])[['t2c_delay', 't2c_liti', 't2c_siti']].mean().reset_index()
-
-
-    d1['psp_class'] = 'Other'
-
-    # delay is greater than long iti and short iti
-    d1.loc[(d1['t2c_delay'] > d1['t2c_liti']) & (d1['t2c_delay'] > d1['t2c_siti']),
-           'psp_class'] = 'Delay Impaired'
-
-    # long iti is greater than delay and short iti
-    d1.loc[(d1['t2c_liti'] > d1['t2c_delay']) & (d1['t2c_liti'] > d1['t2c_siti']),
-           'psp_class'] = 'Long ITI Impaired'
-
-    # short iti is greater than delay and long iti
-    d1.loc[(d1['t2c_siti'] > d1['t2c_delay']) & (d1['t2c_siti'] > d1['t2c_liti']),
-           'psp_class'] = 'Short ITI Impaired'
-
-    # make psp_class an ordered category
-    d1['psp_class'] = pd.Categorical(d1['psp_class'],
-                                     ordered=True,
-                                     categories=[
-                                         'Delay Impaired', 'Long ITI Impaired',
-                                         'Short ITI Impaired', 'Other'
-                                     ])
-
-    # report the number of models in each class
-    print(d1['psp_class'].value_counts())
-
-    # Remove 'Other' class for plotting
-    # d1 = d1[d1['psp_class'] != 'Other']
-    # d1['psp_class'] = d1['psp_class'].cat.remove_unused_categories()
-
-    # report the number of models in each class
-    print(d1['psp_class'].value_counts())
-
-    # Report the unique values of each parameter
-    print("Unique values of sigma_perceptual_noise:", d1['sigma_perceptual_noise'].unique())
-    print("Unique values of alpha_actor:", d1['alpha_actor'].unique())
-    print("Unique values of alpha_critic:", d1['alpha_critic'].unique())
[…]

### 2025-10-17 code/inspect_results_model.py

@@ -1,0 +1,577 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+def inspect_class_I():
+
+    print()
+    print("Inspecting Class I models...")
+
+    # NOTE: Class I setup
+    d1 = pd.read_csv('../output/param_record_class_I_multi.csv')
+
+    d1 = d1.groupby([
+        'sigma_perceptual_noise', 'alpha_actor', 'alpha_critic',
+        'eta_perceptual_drift', 'delay_sensitive_update'
+    ])[['t2c_delay', 't2c_liti', 't2c_siti']].mean().reset_index()
+
+
+    d1['psp_class'] = 'Other'
+
+    # delay is greater than long iti and short iti
+    d1.loc[(d1['t2c_delay'] > d1['t2c_liti']) & (d1['t2c_delay'] > d1['t2c_siti']),
+           'psp_class'] = 'Delay Impaired'
+
+    # long iti is greater than delay and short iti
+    d1.loc[(d1['t2c_liti'] > d1['t2c_delay']) & (d1['t2c_liti'] > d1['t2c_siti']),
+           'psp_class'] = 'Long ITI Impaired'
+
+    # short iti is greater than delay and long iti
+    d1.loc[(d1['t2c_siti'] > d1['t2c_delay']) & (d1['t2c_siti'] > d1['t2c_liti']),
+           'psp_class'] = 'Short ITI Impaired'
+
+    # make psp_class an ordered category
+    d1['psp_class'] = pd.Categorical(d1['psp_class'],
+                                     ordered=True,
+                                     categories=[
+                                         'Delay Impaired', 'Long ITI Impaired',
+                                         'Short ITI Impaired', 'Other'
+                                     ])
+
+    # report the number of models in each class
+    print(d1['psp_class'].value_counts())
+
+    # Remove 'Other' class for plotting
+    # d1 = d1[d1['psp_class'] != 'Other']
+    # d1['psp_class'] = d1['psp_class'].cat.remove_unused_categories()
+
+    # report the number of models in each class
+    print(d1['psp_class'].value_counts())
+
+    # Report the unique values of each parameter
+    print("Unique values of sigma_perceptual_noise:", d1['sigma_perceptual_noise'].unique())
+    print("Unique values of alpha_actor:", d1['alpha_actor'].unique())
+    print("Unique values of alpha_critic:", d1['alpha_critic'].unique())
[…]

### 2025-10-17 code/inspect_results_models.py

@@ -1,23 +1,0 @@
-from experiment_imports import *
-import model_working_memory
-import model_procedural_supervised
-import model_procedural_reinforcement
-
-# model_working_memory.fit('../data/delay/', *(3.0, 0.5))
-# model_working_memory.fit_validate_ppt()
-
-# model_working_memory.fit()
-# model_procedural_supervised.fit()
-# model_procedural_reinforcement.fit()
-
-# model_working_memory.psp()
-# model_procedural_supervised.psp()
-# model_procedural_reinforcement.psp()
-
-model_working_memory.plot_psp()
-model_procedural_supervised.plot_psp()
-model_procedural_reinforcement.plot_psp()
-
-# model_working_memory.fit_validate()
-# model_procedural_supervised.fit_validate()
-# model_procedural_reinforcement.fit_validate()

### 2025-10-17 code/inspect_validate.py

@@ -1,9 +1,0 @@
-import numpy as np
-import scipy as sp
-import pandas as pd
-from ggplot import *
-
-d = pd.read_csv('fit_validate.csv', header=0)
-
-ggplot(d, aes(x=problem, y=alpha_fit)) +\
-    geom_point()

### 2025-10-17 code/model_new.py

@@ -1,513 +1,0 @@
-from experiment_imports import *
-
-"""
-Model class I: No criterion is learned
-- perceptual drift (y/n)
-- criterial drift (NA)
-- delay-sensitive updating (y/n)
-
-Model class II: Criterion is learned
-- perceptual drift (y/n)
-- criterial drift (y/n)
-- delay-sensitive updating (y/n)
-"""
-
-
-def simulate_class_I(params, args):
-
-    # Unpack params into sigma_perceptual_noise, alpha_actor, and alpha_critic
-    sigma_perceptual_noise = params[0]
-    alpha_actor = params[1]
-    alpha_critic = params[2]
-    eta_perceptual_drift = params[3]
-    delay_sensitive_update = params[4]
-
-    # Unpack args into xc_true, t_delay, t_iti, and stim
-    xc_true = args[0]
-    t_delay = args[1]
-    t_iti = args[2]
-    stim = args[3]
-
-    if delay_sensitive_update:
-        t_delay_actor = t_delay
-        t_delay_critic = t_delay
-    else:
-        t_delay_actor = 1
-        t_delay_critic = 1
-
-    num_consecutive_correct = 0
-    t2c = 0
-
-    # static params
-    dim = 100
-    w_base = 0.5
-    w_noise = 0.01
-    vis_amp = 1.0
-
-    # buffers
-    vis_act = np.zeros(dim)
-    w_A = np.zeros(dim)
-    w_B = np.zeros(dim)
-
-    # init
-    v = 0.5
-
-    # init weights: Note that we assume fresh weights every problem
-    w_A = np.random.normal(w_base, w_noise, dim)
-    w_B = np.random.normal(w_base, w_noise, dim)
-
-    for x in stim:
-
-        # Determine true category label
-        if x < xc_true:
-            cat = 1
-        else:
-            cat = 2
-
-        vis_act = np.zeros(dim)
-
-        # compute input activation via radial basis functions
-        vis_dist_x = 0.0
-        for i in range(0, dim):
-            vis_dist_x = x - i
-            vis_act[i] = vis_amp * math.exp(-(vis_dist_x**2) / sigma_perceptual_noise)
-
-        # Compute A and B unit activations via dot product
-        act_A = 0.0
-        act_B = 0.0
-        act_A = np.inner(vis_act, w_A)
-        act_B = np.inner(vis_act, w_B)
-
-        # Compute resp via max
-        discrim = act_A - act_B
[…]

### 2025-10-17 code/model_procedural_reinforcement.py

@@ -1,533 +1,0 @@
-from experiment_imports import *
-
-
-def simulate(xc_true, vis_width, alpha_critic, alpha_actor, delay, iti, stim):
-
-    data_record = {'cat': [], 'x': [], 'y': [], 'resp': []}
-
-    num_consecutive_correct = 0
-    t2c = 0
-
-    # static params:
-    dim = 25
-    w_base = 0.5
-    w_noise = 0.01
-    vis_amp = 1.0
-
-    # buffers
-    vis_act = np.zeros(dim)
-    w_A = np.zeros(dim)
-    w_B = np.zeros(dim)
-
-    # init
-    v = 0.5
-
-    # init weights: Note that we assume fresh weights every problem
-    w_A = np.random.normal(w_base, w_noise, dim)
-    w_B = np.random.normal(w_base, w_noise, dim)
-
-    for x in stim:
-
-        # Determine true category label
-        if x < xc_true:
-            cat = 1
-        else:
-            cat = 2
-
-        # compute input activation via radial basis functions
-        vis_dist_x = 0.0
-        for i in range(0, dim):
-            vis_dist_x = x - i
-            vis_act[i] = vis_amp * math.exp(-(vis_dist_x**2) / vis_width)
-
-        # Compute A and B unit activations via dot product
-        act_A = 0.0
-        act_B = 0.0
-        act_A = np.inner(vis_act, w_A)
-        act_B = np.inner(vis_act, w_B)
-
-        # Compute resp via max
-        discrim = act_A - act_B
-        resp = 1 if discrim > 0 else 2
-
-        # Implement strong lateral inhibition
-        act_A = act_A if resp == 1.0 else 0.0
-        act_B = act_B if resp == 2.0 else 0.0
-
-        # compute outcome
-        r = 1.0 if cat == resp else 0.0
-
-        # compute prediction error
-        delta = (r - v)
-
-        # update critic
-        v += (1 / delay) * alpha_critic * delta
-
-        # update actor
-        for i in range(0, dim):
-
-            if delta < 0:
-                w_A[i] += (
-                    1 /
-                    delay) * alpha_actor * vis_act[i] * act_A * delta * w_A[i]
-                w_B[i] += (
-                    1 /
-                    delay) * alpha_actor * vis_act[i] * act_B * delta * w_B[i]
-            else:
-                w_A[i] += (
-                    1 / delay) * alpha_actor * vis_act[i] * act_A * delta * (
-                        1 - w_A[i])
[…]

### 2025-10-17 code/model_procedural_supervised.py

@@ -1,297 +1,0 @@
-from experiment_imports import *
-
-
-def simulate(xc_init, xc_true, alpha, noise, delay, iti, stim):
-
-    xc = xc_init
-    num_consecutive_correct = 0
-    t2c = 0
-
-    for x in stim:
-
-        # percept is corrupted by perceptual noise
-        noise_err = np.random.normal(0, noise, 1)
-        x += noise_err
-
-        if x < xc_true:
-            cat = 1
-        else:
-            cat = 2
-
-        if x - xc < 0:
-            resp = 1
-        else:
-            resp = 2
-
-        if cat != resp:
-            xc += (1 / delay) * (alpha) * (x - xc)
-            num_consecutive_correct = 0
-        else:
-            num_consecutive_correct += 1
-
-        t2c += 1
-
-        if num_consecutive_correct >= 12:
-            break
-
-    return (t2c)
-
-
-def fit(fit_params, delay, iti, human_data):
-
-    xc_init = fit_params[0]
-    alpha = fit_params[1]
-    noise = fit_params[2]
-
-    # human_data must be a dict with keys: 'cat', 'x', 'y', 'resp'
-    num_trials = len(human_data['x'])
-    stim_range_x = max(human_data['x']) - min(human_data['x'])
-    stim_range_y = max(human_data['y']) - min(human_data['y'])
-    stim_range = max([stim_range_x, stim_range_y])
-
-    xc = xc_init
-    L = []
-
-    for trial in range(num_trials):
-
-        # get current stimulus
-        if stim_range_x > stim_range_y:
-            x = human_data['x'][trial]
-        else:
-            x = human_data['y'][trial]
-
-        # percept is corrupted by perceptual noise
-        noise_err = np.random.normal(0, noise, 1)
-        x += noise_err
-
-        # compute the liklihood of observing the human response given the current model
-        L.extend(abs(x - xc) / float(stim_range))
-
-        if human_data['cat'][trial] != human_data['resp'][trial]:
-            xc += (1 / delay) * (alpha) * (x - xc)
-
-    return (np.prod(L))
-
-
-def fit_validate():
-    # generate fake data from the model using simulate()
-    # find whether or not fit() recovers parameters
-    # find constraints on when fit() performs well and not
-    pass
-
-
-def fit_humans():
-    # load human data
-    # iterate through per human per problem
[…]

### 2025-10-17 code/model_working_memory.py

@@ -1,621 +1,0 @@
-from experiment_imports import *
-
-
-def simulate(x, *args):
-    alpha = x[0]
-    eta_stim = x[1]
-    eta_crit = x[2]
-
-    delay = args[0]
-    iti = args[1]
-
-    num_consecutive_correct = 0
-    t2c = 0
-    max_trials = 200
-
-    simulated_data = {
-        'cat': [],
-        'x': [],
-        'y': [],
-        'rsp': [],
-        'pe': [],
-        'prob': []
-    }
-
-    num_bins = 7
-    bin_width = 14
-    bin_bounds = np.arange(0, 100, bin_width)
-    bins = np.zeros((num_bins, 2))
-    for i in range(bin_bounds.shape[0] - 1):
-        bins[i, 0] = bin_bounds[i]
-        bins[i, 1] = bin_bounds[i + 1]
-
-    bin_dim = np.arange(0, num_bins, 1)
-    np.random.shuffle(bin_dim)
-
-    for prob in range(num_bins):
-        x_range = bins[bin_dim[prob]]
-        xc_true = np.random.uniform(x_range[0], x_range[1])
-        lb1 = x_range[0]
-        ub1 = xc_true - 0.1 * bin_width
-        lb2 = xc_true + 0.1 * bin_width
-        ub2 = x_range[1]
-        xc = np.random.uniform(x_range[0], x_range[1])
-
-        for i in range(max_trials):
-
-            # sample stimulus for current trial
-            x1 = np.random.uniform(lb1, ub1)
-            x2 = np.random.uniform(lb2, ub2)
-            x = np.random.choice([x1, x2])
-
-            # memory for the criterion drifts during the ITI
-            crit_noise = np.random.normal(0, 1)
-            xc += eta_crit * iti * crit_noise
-
-            if x < xc_true:
-                cat = 1
-            else:
-                cat = 2
-
-            if x - xc < 0:
-                rsp = 1
-            else:
-                rsp = 2
-
-            # the criterion and the stimulus drift over the feedback delay
-            stim_drift = np.random.normal(0, 1)
-            crit_drift = np.random.normal(0, 1)
-            x += eta_stim * delay * stim_drift
-            xc += eta_crit * delay * crit_drift
-
-            if cat != rsp:
-                xc += alpha * (x - xc)
-                num_consecutive_correct = 0
-            else:
-                num_consecutive_correct += 1
-
-            t2c += 1
-
-            simulated_data['cat'].append(cat)
-            simulated_data['x'].append(x)
[…]

### 2025-10-17 code/models.py

@@ -1,0 +1,513 @@
+from experiment_imports import *
+
+"""
+Model class I: No criterion is learned
+- perceptual drift (y/n)
+- criterial drift (NA)
+- delay-sensitive updating (y/n)
+
+Model class II: Criterion is learned
+- perceptual drift (y/n)
+- criterial drift (y/n)
+- delay-sensitive updating (y/n)
+"""
+
+
+def simulate_class_I(params, args):
+
+    # Unpack params into sigma_perceptual_noise, alpha_actor, and alpha_critic
+    sigma_perceptual_noise = params[0]
+    alpha_actor = params[1]
+    alpha_critic = params[2]
+    eta_perceptual_drift = params[3]
+    delay_sensitive_update = params[4]
+
+    # Unpack args into xc_true, t_delay, t_iti, and stim
+    xc_true = args[0]
+    t_delay = args[1]
+    t_iti = args[2]
+    stim = args[3]
+
+    if delay_sensitive_update:
+        t_delay_actor = t_delay
+        t_delay_critic = t_delay
+    else:
+        t_delay_actor = 1
+        t_delay_critic = 1
+
+    num_consecutive_correct = 0
+    t2c = 0
+
+    # static params
+    dim = 100
+    w_base = 0.5
+    w_noise = 0.01
+    vis_amp = 1.0
+
+    # buffers
+    vis_act = np.zeros(dim)
+    w_A = np.zeros(dim)
+    w_B = np.zeros(dim)
+
+    # init
+    v = 0.5
+
+    # init weights: Note that we assume fresh weights every problem
+    w_A = np.random.normal(w_base, w_noise, dim)
+    w_B = np.random.normal(w_base, w_noise, dim)
+
+    for x in stim:
+
+        # Determine true category label
+        if x < xc_true:
+            cat = 1
+        else:
+            cat = 2
+
+        vis_act = np.zeros(dim)
+
+        # compute input activation via radial basis functions
+        vis_dist_x = 0.0
+        for i in range(0, dim):
+            vis_dist_x = x - i
+            vis_act[i] = vis_amp * math.exp(-(vis_dist_x**2) / sigma_perceptual_noise)
+
+        # Compute A and B unit activations via dot product
+        act_A = 0.0
+        act_B = 0.0
+        act_A = np.inner(vis_act, w_A)
+        act_B = np.inner(vis_act, w_B)
+
+        # Compute resp via max
+        discrim = act_A - act_B
[…]

### 2025-10-21 write/cover_letter_revision.tex
bug fix

\documentclass[12pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\setstretch{1.0}
\signature{Matthew J. Crossley \\ on behalf of all authors}
\date{\today}

\begin{document}

\begin{letter}{Editorial Office \\ 
\textit{Journal of Experimental Psychology: Learning, Memory, and Cognition}}

\opening{Dear Colleagues,}

Please find enclosed our detailed response to the reviewers’
and action editor’s comments on our manuscript entitled
\textit{“Is there a Criterion in Criterial Learning?
Insights from Studying Feedback Delays”} (Manuscript
\#XLM-2024-0436).

We have carefully considered all feedback and made extensive
revisions throughout the paper. These include a substantial
expansion of the model space, clearer presentation of
simulation results, reorganization of manuscript sections,
and several textual clarifications. We believe these changes
have significantly strengthened the manuscript and improved
its clarity and theoretical contribution.

We sincerely thank you and the reviewers for the insightful
comments and constructive suggestions that guided these
improvements.

\closing{Sincerely,}

\end{letter}

\end{document}

### 2025-10-21 write/main_revision.tex
bug fix

@@
 the Long ITI condition ($t(21.60) = -2.14$, $p < .05$,
-$-.80$).  Performance in the Control and Long-ITI conditions
-did not differ significantly from each other ($t(25.97) =
--1.50$, $p=.15$, $d=-.53$).
+$d=-.80$).  Performance in the Control and Long-ITI
+conditions did not differ significantly from each other
+($t(25.97) = -1.50$, $p=.15$, $d=-.53$).
 
@@
 Figure \ref{fig_exp_2_t2c}B. An independent-samples t-test
-revealed no significant effect of condition ($t(32) =
+revealed no significant effect of condition ($t(32.0) =
 -1.99$, $p = .06$, $d = -.68$).

### 2025-10-21 code/inspect_results_behave.py

@@ -1,6 +1,6 @@
 from util_func import *
 from imports import *
 
-report_exp_1()
-# report_exp_2()
+# report_exp_1()
+report_exp_2()
 # report_exp_3()

### 2025-10-21 code/util_exgauss.py

@@ -456,10 +456,10 @@ def plot_exg_mixture(model, X, bins=30, title=None, ax=None):
                        color=colors[k % len(colors)],
                        label="")
 
-    ax.set_xlabel("Trials to criterion", fontsize=12)
-    ax.set_ylabel("Density", fontsize=12)
+    ax.set_xlabel("Trials to criterion", fontsize=14)
+    ax.set_ylabel("Density", fontsize=14)
     if title:
-        ax.set_title(title)
+        ax.set_title(title, fontsize=14)
     yl = ax.get_ylim()
     ax.set_ylim(min(yl[0], -0.01), yl[1])
     ax.legend(loc="best")

### 2025-10-21 code/inspect_results_behave.py
tiny error

@@ -1,6 +1,6 @@
 from util_func import *
 from imports import *
 
-# report_exp_1()
+report_exp_1()
 report_exp_2()
 # report_exp_3()

### 2025-10-21 code/inspect_results_model.py

@@ -437,141 +437,3 @@ def inspect_class_II():
 if __name__ == "__main__":
     inspect_class_I()
     inspect_class_II()
-
-# # NOTE: old figure
-# fig = plt.figure(figsize=(10, 8))
-# 
-# xywhA = [0.1, 0.6, 0.375, 0.3]
-# xywhC = [0.1, 0.15, 0.375, 0.3]
-# xywhB = [0.55, 0.55, 0.4, 0.45]
-# xywhD = [0.55, 0.1, 0.4, 0.45]
-# 
-# ax1 = fig.add_axes(xywhA)
-# ax2 = fig.add_axes(xywhC)
-# ax3 = fig.add_axes(xywhB, projection='3d')
-# ax4 = fig.add_axes(xywhD, projection='3d')
-# 
-# ax_labels = ['A', 'C', 'B', 'D']
-# xyA = (0.0, 0.95)
-# xyC = (0.0, 0.5)
-# xyB = (0.5, 0.95)
-# xyD = (0.5, 0.5)
-# xy = [xyA, xyC, xyB, xyD]
-# for i, ax in enumerate([ax1, ax2, ax3, ax4]):
-#     ax.annotate(ax_labels[i],
-#                 xy=xy[i],
-#                 xycoords='figure fraction',
-#                 horizontalalignment='left',
-#                 verticalalignment='top',
-#                 fontsize=20)
-# 
-# # Panel A: Proportion of parameter space
-# ax = ax1
-# sns.countplot(x='psp_class',
-#               data=dp,
-#               hue='psp_class',
-#               stat='proportion',
-#               legend=False,
-#               ax=ax)
-# ax.spines['top'].set_visible(False)
-# ax.spines['right'].set_visible(False)
-# ax.spines['left'].set_linewidth(1.5)
-# ax.spines['bottom'].set_linewidth(1.5)
-# ax.set_xlabel('')
-# ax.set_ylabel('Proportion of parameter space', fontsize=12)
-# ax.set_xticks(np.arange(0, 4, 1))
-# labels = [
-#     tick.get_text().replace('Impaired', '\nImpaired')
-#     for tick in ax.get_xticklabels()
-# ]
-# ax.set_xticklabels(labels, fontsize=10)
-# 
-# # Panel C: parameter values
-# ax = ax2
-# pr = dp.copy()
-# pr = pr.melt(id_vars='psp_class',
-#              value_vars=[
[…]
